### 2011-03-03 templates/Zusammenfassung.tex
Templates mit etwas Beispielen für Latex

 \documentclass[a4paper, twocolumn]{article}

\usepackage{amsmath, amsthm, amssymb, amsfonts} 
%\usepackage{german}
\usepackage[utf8]{inputenc}
\usepackage{multicol}  
\usepackage{multirow}
%\usepackage{dsfont} 
%\usepackage[rflt]{floatflt}  
\usepackage{graphics}
\usepackage{epsfig} 
\usepackage[pdftex]{hyperref}
\usepackage{moreverb}

% ZuFa Template laden
% Layout by Thorben Bocheneck
\usepackage{tbabstract}

\title{Templates}
\author{Patrick Bänziger}

%Kommandos

\definecolor{lightgrey}{rgb}{0.96,0.96,0.96}
\newcommand{\keyword}[1]{\textbf{\color{black}{#1}}}
\newcommand{\example}[1]{\textbf{\medskip\colorbox{lightgrey}{\color{black}{Beispiel: #1}}}}
\newcommand{\typecomb}{$\blacktriangleright$}

\begin{document}
\maketitle

\section{Examples follow below}

\subsection{Motivation}
As studies have shown, the vast majority ($\approx$ 84\%) of all software projects can be considered unsuccessful. They either are too late, over budget, don't contain all promised features or get outright cancelled. \\

There are two major challenges which need to be addressed in successful software engineering:

\subsubsection*{Complexity}
Complexity in software engineering is introduced by three main causes, namely the:
\begin{itemize}
	\item complexity of the problem domain
	\item complexity of the development process
	\item required flexibility of software
\end{itemize}
Among the aspects of complexity are team fluctuation, multi-version software and conflicting objectives (quality versus price).
imation methods, depending on the situation:\\

$\mathbf{CEAC_1}$: If the current variances (e.g. lower staff performance, higher complexity) will now remain at the same level.
$$CEAC_1 = \frac{BAC}{CPI}$$

$\mathbf{CEAC_2}$: If the past variances will not occur in the future anymore. (We take actual cost for the past and remain at estimates for the future.)
$$CEAC_2 = AC + BAC - EV$$

$\mathbf{CEAC_3}$: If the old estimate was just fundamentally wrong.
$$CEAC_3 = AC + newEstimate$$

\subsubsection{Golden rules}
What remains to be said? There is the question how to calculate the value of an activity that has started.

\begin{enumerate}
	\item Every chunk of earned value should be verified by \textbf{physically examining the product} of the activity.
	\item For unfinished activities, earned value estimates are just a guess! One of the following rules should be applied consistently:
	\begin{itemize}
		\item \textbf{50/50 Rule:} After a tasks starts, it is considered 50\% complete until it is finished.
		\item \textbf{20/80:} After a task starts, it is considered 20\% complete. This is realistic, as the timeconsuming or difficult parts are towards the end of a task.
		\item \textbf{0/100:} The task does not generate any earned value until completion. Very conservative.
	\end{itemize}
\end{enumerate}

\end{document}

### 2011-06-06 project/EvaluateInpainting.m
Remove Evaluation Script

@@ -1,55 +1,0 @@
-% Measure approximation error and compression ratio for several images.
-%
-% NOTE Images must be have .png ending and reside in the same folder.
-
-global U;
-
-file_list = dir(); 
-k = 1;
-
-Errors = []; % mean squared errors for each image would be stored here
-
-
-% UPDATE:
-% If available load the dictionary.mat file with dictionary U
-try 
-  temp = load('dictionary.mat');
-  U = temp.U;
-catch
-end
-
-for i = 3:length(dir) % running through the folder
-    
-    file_name = file_list(i).name; % get current filename
-    
-    % Only keep the images in the loop
-    if (length(file_name) < 5)
-        continue;
-    elseif ( max(file_name(end-4:end) ~= '2.png'))
-        continue;
-    end
-    mask_name = [file_name(1:end-4) '_mask.png'];
-        
-    % Read image, convert to double precision and map to [0,1] interval
-    I = imread(file_name); 
-    I = double(I) / 255; 
-    
-    % Read the respective binary mask
-    % EVALUATION IS DONE WITH A FIXED MASK
-    mask = imread(mask_name);
-    
-    I_mask = I;
-    I_mask(~mask) = 0;
-          
-    % Call the main inPainting function
-    I_rec = inPainting(I_mask, mask);
-    
-    % Measure approximation error
-    Errors(k) = mean(mean(mean( ((I - I_rec) ).^2)));
-    
-    k = k+1;
-end
-
-Result(1) = mean(Errors);
-
-disp(['Average quadratic error: ' num2str(Result(1))])

### 2011-06-06 project/gaussInterpolate.m
+ Added gaussInterpolate

@@ -1,0 +1,12 @@
+function I_out = gaussInterpolate(I, mask)
+
+mask = double(mask);
+
+sigma = 0.8;
+gauss_size = 5;
+
+h = fspecial('gaussian', gauss_size, sigma);
+I_masked = I.*mask;
+
+I_filtered = imfilter(I_masked, h, 'replicate')./imfilter(mask, h, 'replicate');
+I_out = (I_filtered.*(1-mask))+(I_masked.*(mask));

### 2011-06-06 project/linearInPainting.m

@@ -1,0 +1,139 @@
+function I_rec = linearInPainting(I, mask)
+
+% Perform the actual inpainting of the image 
+
+% INPUT
+% I: (n x n) masked image
+% mask: (n x n) the mask hidding image information
+%
+% OUTPUT
+% I_rec = Reconstructed image 
+
+global U;
+
+% Parameters
+rc_min = 0.01; % rc_min: minimal residual correlation before stopping
+neib = 32; % neib: The patch sizes used in the decomposition of the image
+sigma = 0.01; % sigma: residual error stopping criterion, normalized by signal norm
+
+patchsize=16; neighbourhood=8; t=88;
+
+
+% Get patches of size neib x neib from the image and the mask and
+% convert each patch to 1D signal
+
+[w, h] = size(I);
+
+for x=1:w
+  for y=1:h
+    if mask(x, y) == 0
+      values = 0;
+      s = 0;
+
+      for (i=1:10)
+        try
+          if mask(x+i, y) ~= 0
+            values = values + I(x+i, y)/i;
+            s = s + 1/i;
+            break;
+          end
+        catch
+          break;
+        end
+      end
+
+      for (i=1:10)
+        try
+          if mask(x, y+i) ~= 0
+            values = values + I(x, y+i)/i;
+            s = s + 1/i;
+            break;
+          end
+        catch
+          break;
+        end
+      end
+
+      for (i=1:10)
+        try
+          if mask(x, y-i) ~= 0
+            values = values + I(x, y-i)/i;
+            s = s + 1/i;
+            break;
+          end
+        catch
+          break;
+        end
+      end
+
+      for (i=1:10)
+        try
+          if mask(x-i, y) ~= 0
+            values = values + I(x-i, y)/i;
+            s = s + 1/i;
+            break;
+          end
+        catch
+          break;
+        end
+      end
+
+      I(x, y) = values / s;
+    end
+  end
+end
+
+% Create overlapping patches
+numBlocks=size(I,1)/patchsize;
+
+if(mod(size(I,1),patchsize)~=0)
+  throw('Imagesize not multiple of pactchsize');
+end
+
+tic;
+
+IG = zeros(size(I,1)+2*neighbourhood, size(I,1)+2*neighbourhood);
[…]

### 2011-06-06 project/inPainting.m
+ Finished fft-gauss with one iteration

@@ -1,4 +1,4 @@
-function I_rec = linearInPainting(I, mask)
+function I_rec = InPainting(I, mask)
 
 % Perform the actual inpainting of the image 
 
@@ -9,8 +9,6 @@ function I_rec = linearInPainting(I, mask)
 % OUTPUT
 % I_rec = Reconstructed image 
 
-global U;
-
 % Parameters
 rc_min = 0.01; % rc_min: minimal residual correlation before stopping
 neib = 32; % neib: The patch sizes used in the decomposition of the image
@@ -18,70 +16,11 @@ sigma = 0.01; % sigma: residual error stopping criterion, normalized by signal n
 
 patchsize=16; neighbourhood=8; t=88;
 
-
 % Get patches of size neib x neib from the image and the mask and
 % convert each patch to 1D signal
-
-[w, h] = size(I);
-
-for x=1:w
-  for y=1:h
-    if mask(x, y) == 0
-      values = 0;
-      s = 0;
-
-      for (i=1:10)
-        try
-          if mask(x+i, y) ~= 0
-            values = values + I(x+i, y)/i;
-            s = s + 1/i;
-            break;
-          end
-        catch
-          break;
-        end
-      end
-
-      for (i=1:10)
-        try
-          if mask(x, y+i) ~= 0
-            values = values + I(x, y+i)/i;
-            s = s + 1/i;
-            break;
-          end
-        catch
-          break;
-        end
-      end
-
-      for (i=1:10)
-        try
-          if mask(x, y-i) ~= 0
-            values = values + I(x, y-i)/i;
-            s = s + 1/i;
-            break;
-          end
-        catch
-          break;
-        end
-      end
-
-      for (i=1:10)
-        try
-          if mask(x-i, y) ~= 0
-            values = values + I(x-i, y)/i;
-            s = s + 1/i;
-            break;
-          end
-        catch
-          break;
-        end
-      end
-
-      I(x, y) = values / s;
-    end
-  end
-end
+I = gaussInterpolate(I, mask);
+%I_rec = I;
+%return;
 
 % Create overlapping patches
 numBlocks=size(I,1)/patchsize;
@@ -111,7 +50,9 @@ for i=1:numBlocks
         block = IG( patchsize*(i-1)+1 : patchsize*i + 2*neighbourhood , patchsize*(j-1)+1:patchsize*j+2*neighbourhood);
         F = fft2(block);
         T = abs(F);
[…]

### 2011-06-06 project/Paper/paper.tex

\documentclass[10pt,conference,compsocconf]{IEEEtran}

%\usepackage{times}
%\usepackage{balance}
\usepackage{url}
\usepackage{graphicx}	% For figure environment
\usepackage[pdftex]{hyperref}
\usepackage[utf8]{inputenc} % To make umlauts and other symbols correct

\begin{document}
\title{Image restoration using Fourier transforms and linear interpolation}

\author{
  Patrick Bänziger, Kaspar Etter, Jan Rüegg \\
  Department of Computer Science, ETH Zurich, Switzerland
}

\maketitle

\begin{abstract}
(( TODO ))
  A critical part of scientific discovery is the
  communication of research findings to peers or the general public.
  Mastery of the process of scientific communication improves the
  visibility and impact of research. While this guide is a necessary
  tool for learning how to write in a manner suitable for publication
  at a scientific venue, it is by no means sufficient, on its own, to
  make its reader an accomplished writer. We also describe the rules
  for submission in the computational intelligence laboratory.
  This guide should be a
  starting point for further development of writing skills.
\end{abstract}

\section{Introduction}
(( TODO ))
The aim of writing a paper is to infect the mind of your reader with
the brilliance of your idea \cite{jones08}. 
The hope is that after reading your
paper, the audience will be convinced to try out your idea. In other
words, it is the medium to transport the idea from your head to your
reader's head. 
In the following
section, we show a common structure of scientific papers and briefly
outline some tips for writing good papers in
Section \ref{sec:tips-writing}.

At that
point, it is important that the reader is able to reproduce your
work~\cite{schwab00,wavelab,gentleman05}. This is why it is also
important that if the work has a computational component, the software
associated with producing the results are also made available in a
useful form. Several guidelines for making your user's experience with
your software as painless as possible is given in
Section \ref{sec:tips-software}.

Finally, we describe the rules for submission in the computational
intelligence laboratory in Section~\ref{sec:cil}. 
This brief guide is by no means sufficient, on its own, to
make its reader an accomplished writer. The reader is urged to use the
references to further improve his or her writing skills.

\section{The Structure of a Paper}
\label{sec:structure-paper}

Scientific papers usually begin with the description of the problem,
justifying why the problem is interesting. Most importantly, it argues
that the problem is still unsolved, or that the current solutions are
unsatisfactory. This leads to the main gist of the paper, which is
``the idea''. The authors then show evidence, using derivations or
experiments, that the idea works. Since science does not occur in a
vacuum, a proper comparison to the current state of the art is often
part of the results. Following these ideas, papers usually have the
following structure:
\begin{description}
\item[Abstract] \ \\
  Short description of the whole paper, to help the
  reader decide whether to read it.
\item[Introduction] \ \\
  Describe your problem and state your
  contributions.
\item[Methods] \ \\
  Describe your idea and how it was implemented to solve
  the problem. Survey the related work, giving credit where credit is
  due.
\item[Results] \ \\
  Show evidence to support your claims made in the
  introduction.
\item[Discussion] \ \\
  Discuss the strengths and weaknesses of your
  approach, based on the results. Point out the implications of your
  novel idea on the application concerned.
\item[Summary] \ \\
  Summarise your contributions in light of the new
  results.
\end{description}

\section{Tips for Good Writing}
\label{sec:tips-writing}

The ideas for good writing have come
from~\cite{editor10,jones08,anderson04}.

\subsection{Getting Help}
One should try to get a draft read by as many friendly people as
possible. And remember to treat your test readers with respect. If
they are unable to understand something in your paper, then it is
highly likely that your reviewers will not understand it
either. Therefore, do not be defensive about the criticisms you get,
but use it as an opportunity to improve the paper. Before your submit
your friends to the pain of reading your draft, please \emph{use a
  spell checker}.

\subsection{Abstract}
The abstract should really be written last, along with the title of
the paper. The four points that should be covered~\cite{jones08}:
\begin{enumerate}
\item State the problem.
\item Say why it is an interesting problem.
\item Say what your solution achieves.
\item Say what follows from your solution.
\end{enumerate}

\subsection{Figures and Tables}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{images/denoised_signal_1d}
  \caption{Signal compression and denoising using the Fourier basis.}
  \label{fig:denoise-fourier}
\end{figure}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\columnwidth]{images/local_wdenoised_1d}
  \caption{Signal compression and denoising using the Daubechies wavelet basis.}
  \label{fig:denoise-wavelet}
\end{figure}

Use examples and illustrations to clarify ideas and results. For
example, by comparing Figure~\ref{fig:denoise-fourier} and
Figure~\ref{fig:denoise-wavelet}, we can see the two different
situations where Fourier and wavelet basis perform well.

\subsection{Methods}

The style in this section should read as if you were verbally
describing the conduct of the experiment. 
The methods section is not a step-by-step, directive,
protocol as you might see in your lab manual, but detailed enough such
that an interested reader can reproduce your
work~\cite{anderson04,wavelab}.
The methods
section should describe what was
done to answer the research question, describe how it was done,
justify the experimental design, and
explain how the results were analyzed.

The methods section of a research paper provides the information by
which a study's validity is judged.
Therefore, it requires a clear and precise description of how an
experiment was done, and the rationale
for why specific experimental procedures were chosen.
It is usually helpful to
structure the methods section by~\cite{kallet04methods}:
\begin{enumerate}
\item Describing the algorithms used in the study, briefly including
  details such as hyperparameter values (e.g. thresholds), and
  preprocessing steps (e.g. normalizing the data to have mean value of
  zero).
\item Explaining how the materials were prepared, for example the
  images used and their resolution.
\item Describing the research protocol, for example which examples
  were used for estimating the parameters (training) and which were
  used for computing performance.
\item Explaining how measurements were made and what
  calculations were performed. Do not reproduce the full source code in
  the paper, but explain the key steps.
\end{enumerate}

\subsection{Results}

Organize the results section based on the sequence of table and
figures you include. Prepare the tables and figures as soon as all
the data are analyzed and arrange them in the sequence that best
presents your findings in a logical way. A good strategy is to note,
on a draft of each table or figure, the one or two key results you
want to addess in the text portion of the results.
The information from the figures is
summarized in Table~\ref{tab:fourier-wavelet}.

\begin{table*}[htbp]
  \centering
  \begin{tabular}[c]{|l||l|l|l|}
    \hline
    Basis&Support&Suitable signals&Unsuitable signals\\
    \hline
    Fourier&global&sine like&localized\\
    wavelet&local&localized&sine like\\
    \hline
  \end{tabular}
  \caption{Characteristics of Fourier and wavelet basis.}
  \label{tab:fourier-wavelet}
\end{table*}

When reporting computational or measurement results, always
report the mean (average value) along with a measure of variablility
(standard deviation(s) or standard error of the mean).

\section{Tips for Good Software}
\label{sec:tips-software}

There is a lot of literature (for example~\cite{hunt99pragmatic} and
\cite{spolsky04software}) on how to write software. It is not the
intention of this section to replace software engineering
courses. However, in the interests of reproducible
research~\cite{schwab00}, there are a few guidelines to make your
reader happy:
\begin{itemize}
\item Have a \texttt{README} file that (at least) describes what your
  software does, and which commands to run to obtain results. Also
  mention anything special that needs to be set up, such as
  toolboxes\footnote{For those who are
  particularly interested, other common structures can be found at
  \url{http://en.wikipedia.org/wiki/README} and
  \url{http://www.gnu.org/software/womb/gnits/}.}.
\item A list of authors and contributors can be included in a file
  called \texttt{AUTHORS}, acknowledging any help that you may have
  obtained. For small projects, this information is often also
  included in the \texttt{README}.
\item Use meaningful filenames, and not \texttt{temp1.m},
  \texttt{temp2.m}. The code should also unzip into a subdirectory.
\item Document your code. Each file should at least have a short
  description about its reason for existence. Non obvious steps in the
  code should be commented.
\item Describe how the results presented in your paper can potentially
  be reproduced.
\end{itemize}

\section{Computational Intelligence Laboratory Requirements}
\label{sec:cil}

Your semester project is a group effort. It consists of four parts:
\begin{enumerate}
\item The programming assignments you solve during the semester.
\item Developing a novel solution for one of the assignments, e.g. by
  combining methods from previous programming assignments into a novel
  solution.
\item Comparing your novel solution to previous assignments.
\item Writing up your findings in a short scientific paper.
\end{enumerate}

\subsection{Developing a Novel Solution}

As your final programming assignment, you develop a novel solution to
one of the four application problems. You are free to exploit any idea
you have, provided it is not identical to any other group submission
or existing Matlab implementation of an algorithm on the
internet\footnote{\url{http://www.ethz.ch/students/semester/plagiarism_s_en.pdf}}.

Two examples for developing a novel solution:
\begin{itemize}
\item You implemented a collaborative filtering algorithm based on
  dimension reduction as part of an assignment. Now you apply
  dimension reduction to inpainting.
\item You implemented both a clustering and a sparse coding algorithm
  for image compression. Now you combine both techniques into a novel
  compression method.
\end{itemize}

\subsection{Comparison to Baselines}

You compare your novel algorithm to \emph{at least two baseline
  algorithms}. For the baselines, you can use the implementations you
developed as part of the programming assignments.

\subsection{Write Up}

The submission must be in PDF form, using the \LaTeX{} template
corresponding to the IEEE style of publication. Refer to
Section~\ref{sec:latex-primer} for more information about preparing
your document. The document should be a maximum of {\bf 4 pages}.

\subsection{\LaTeX{} Primer}
\label{sec:latex-primer}

\LaTeX{} is one of the most commonly used document preparation systems
for scientific journals and conferences. It is based on the idea
that authors should be able to focus on the content of what they are
writing without being distracted by its visual presentation.
The source of this file can be used as a starting point for how to use
the different commands in \LaTeX{}. We are using an IEEE style for
this course.

\subsubsection{Installation}

There are various different packages available for processing \LaTeX{}
documents.
On Windows, use the Mik\TeX{} package (\url{http://miktex.org/}), and
on OSX use MacTeX
(\url{http://www.tug.org/mactex/2009/}). Alternatively, on OSX, you
can install the \texttt{tetex} package via
Fink\footnote{\url{http://www.finkproject.org/}} or 
Macports\footnote{\url{http://www.macports.org/}}.

\subsubsection{Compiling \LaTeX{}}
Your directory should contain at least 4 files, in addition to image
files. Images should be in \texttt{.png}, \texttt{.jpg} or
\texttt{.pdf} format.
\begin{itemize}
\item IEEEtran.cls
\item IEEEtran.bst
\item groupXX-submission.tex
\item groupXX-literature.bib
\end{itemize}
Note that you should replace groupXX with your chosen group name.
Then, from the command line, type:
\begin{verbatim}
$ pdflatex groupXX-submission
$ bibtex groupXX-literature
$ pdflatex groupXX-submission
$ pdflatex groupXX-submission
\end{verbatim}
This should give you a PDF document \texttt{groupXX-submission.pdf}.

\subsubsection{Equations}

There are three types of equations available: inline equations, for
example $y=mx + c$, which appear in the text, unnumbered equations
$$y=mx + c,$$
which are presented on a line on its own, and numbered equations
\begin{equation}
  \label{eq:linear}
  y = mx + c
\end{equation}
which you can refer to at a later point (Equation~(\ref{eq:linear})).

\subsubsection{Tables and Figures}

Tables and figures are ``floating'' objects, which means that the text
can flow around it.
Note
that \texttt{figure*} and \texttt{table*} cause the corresponding
figure or table to span both columns.

\subsection{Grading}

There are two different types of grading criteria applied to your
project, with the corresponding weights shown in brackets.
\begin{description}
\item[Competitive] \ \\
  The following criteria is scored based on your rank
  in comparison with the rest of the class.
  \begin{itemize}
  \item time taken for computation (10\%)
  \item average rank for all other criteria relevant to the task, for
    example reconstruction error and sparsity (20\%)
  \end{itemize}
  The ranks will then be converted on a linear scale into a grade
  between 4 and 6.
\item[Non-competitive] \ \\
  The following criteria is scored based on an
  evaluation by the teaching assistants.
  \begin{itemize}
  \item quality of paper (30\%)
  \item quality of implementation (20\%)
  \item creativity of solution (20\%)
  \end{itemize}
\end{description}

\subsection{Submission System}

The deadline for submitting your project report is Friday, 25 June
2010.
You need to submit:
\begin{itemize}
\item PDF of paper.
\item Archive (\texttt{.tar.gz} or \texttt{.zip}) of software. Please
  do not forget to include author information in the source archive.
\end{itemize}

\section{Summary}

The aim of a scientific paper is to convey the idea or discovery of
the researcher to the minds of the readers. The associated software
package provides the relevant details, which are often only briefly
explained in the paper, such that the research can be reproduced.
To write good papers, identify your key idea, make your contributions
explicit, and use examples and illustrations to describe the problems
and solutions.

\section*{Acknowledgements}
The author thanks Christian Sigg for his careful reading and helpful
suggestions.

\bibliographystyle{IEEEtran}
\bibliography{biblio}
\end{document}

### 2011-06-06 project/EvaluateInpainting.m
+ readded gaussInterpolate

@@ -13,6 +13,7 @@ Errors = []; % mean squared errors for each image would be stored here
 % UPDATE:
 % If available load the dictionary.mat file with dictionary U
 try 
+  global U;
   temp = load('dictionary.mat');
   U = temp.U;
 catch

### 2011-06-06 project/Paper/paper.tex
Removed lots of blabla, now should have only relevant stuff left. Ready to start!

@@
 \begin{document}
-\title{Image restoration using Fourier transforms and linear interpolation}
+\title{Image restoration using dimension reduction in the Fourier domain}
 
@@
 \begin{abstract}
-(( TODO ))
-  A critical part of scientific discovery is the
-  communication of research findings to peers or the general public.
-  Mastery of the process of scientific communication improves the
-  visibility and impact of research. While this guide is a necessary
-  tool for learning how to write in a manner suitable for publication
-  at a scientific venue, it is by no means sufficient, on its own, to
-  make its reader an accomplished writer. We also describe the rules
-  for submission in the computational intelligence laboratory.
-  This guide should be a
-  starting point for further development of writing skills.
+  This paper suggests a novel approach to restore noisy images.
+  ((( TODO )))
+  The abstract should really be written last, along with the title of
+the paper. The four points that should be covered~\cite{jones08}:
+\begin{enumerate}
+\item State the problem.
+\item Say why it is an interesting problem.
+\item Say what your solution achieves. (CLAIM!)
+\item Say what follows from your solution.
+\end{enumerate}
 \end{abstract}
@@
 \section{Introduction}
-(( TODO ))
+In image inpainting, the aim is to restore an image from a disturbed source.
+These distortions may, for example, result from a transmission over a noisy channel.
+We distinguish inpainting from image 
+((( TODO )))
 The aim of writing a paper is to infect the mind of your reader with
@@
 reader's head.
-In the following
-section, we show a common structure of scientific papers and briefly
-outline some tips for writing good papers in
-Section \ref{sec:tips-writing}.
 
-At that
-point, it is important that the reader is able to reproduce your
-work~\cite{schwab00,wavelab,gentleman05}. This is why it is also
-important that if the work has a computational component, the software
-associated with producing the results are also made available in a
-useful form. Several guidelines for making your user's experience with
-your software as painless as possible is given in
-Section \ref{sec:tips-software}.
+\section{Methods}
+\subsection{Related work}
+((( TODO )))
+\subsection{New approach}
+((( TODO )))
+Describe your idea and how it was implemented to solve
+  the problem. Survey the related work, giving credit where credit is
+  due.
 
-Finally, we describe the rules for submission in the computational
-intelligence laboratory in Section~\ref{sec:cil}. 
-This brief guide is by no means sufficient, on its own, to
-make its reader an accomplished writer. The reader is urged to use the
-references to further improve his or her writing skills.
+\section{Results}
+ ((( TODO )))
+  Show evidence to support your claims made in the
+  introduction.
+
+    The style in this section should read as if you were verbally
+describing the conduct of the experiment. 
+The methods section is not a step-by-step, directive,
+protocol as you might see in your lab manual, but detailed enough such
+that an interested reader can reproduce your
+work~\cite{anderson04,wavelab}.
+The methods
+section should describe what was
+done to answer the research question, describe how it was done,
+justify the experimental design, and
+explain how the results were analyzed.
+
+The methods section of a research paper provides the information by
+which a study's validity is judged.
+Therefore, it requires a clear and precise description of how an
+experiment was done, and the rationale
+for why specific experimental procedures were chosen.
+It is usually helpful to
+structure the methods section by~\cite{kallet04methods}:
+\begin{enumerate}
+\item Describing the algorithms used in the study, briefly including
+  details such as hyperparameter values (e.g. thresholds), and
+  preprocessing steps (e.g. normalizing the data to have mean value of
+  zero).
+\item Explaining how the materials were prepared, for example the
+  images used and their resolution.
+\item Describing the research protocol, for example which examples
+  were used for estimating the parameters (training) and which were
+  used for computing performance.
+\item Explaining how measurements were made and what
+  calculations were performed. Do not reproduce the full source code in
+  the paper, but explain the key steps.
+\end{enumerate}
+
+Organize the results section based on the sequence of table and
+figures you include. Prepare the tables and figures as soon as all
+the data are analyzed and arrange them in the sequence that best
+presents your findings in a logical way. A good strategy is to note,
+on a draft of each table or figure, the one or two key results you
+want to address in the text portion of the results.
+The information from the figures is
+summarized in Table~\ref{tab:fourier-wavelet}.
+
+\begin{table*}[htbp]
+  \centering
+  \begin{tabular}[c]{|l||l|l|l|}
+    \hline
+    Basis&Support&Suitable signals&Unsuitable signals\\
+    \hline
+    Fourier&global&sine like&localized\\
+    wavelet&local&localized&sine like\\
+    \hline
+  \end{tabular}
+  \caption{Characteristics of Fourier and wavelet basis.}
+  \label{tab:fourier-wavelet}
+\end{table*}
+
+When reporting computational or measurement results, always
+report the mean (average value) along with a measure of variablility
+(standard deviation(s) or standard error of the mean).
+
+You compare your novel algorithm to \emph{at least two baseline
+  algorithms}. For the baselines, you can use the implementations you
+developed as part of the programming assignments.
+
+\section{Discussion}
+  Discuss the strengths and weaknesses of your
+  approach, based on the results. Point out the implications of your
+  novel idea on the application concerned. (((( TODO ))))
 
@@
 
-\section{Tips for Good Writing}
-\label{sec:tips-writing}
-
-The ideas for good writing have come
-from~\cite{editor10,jones08,anderson04}.
-
-\subsection{Getting Help}
-One should try to get a draft read by as many friendly people as
-possible. And remember to treat your test readers with respect. If
-they are unable to understand something in your paper, then it is
-highly likely that your reviewers will not understand it
-either. Therefore, do not be defensive about the criticisms you get,
-but use it as an opportunity to improve the paper. Before your submit
-your friends to the pain of reading your draft, please \emph{use a
-  spell checker}.
-
-\subsection{Abstract}
-The abstract should really be written last, along with the title of
-the paper. The four points that should be covered~\cite{jones08}:
-\begin{enumerate}
-\item State the problem.
-\item Say why it is an interesting problem.
-\item Say what your solution achieves.
-\item Say what follows from your solution.
-\end{enumerate}
-
 \subsection{Figures and Tables}
@@
 
-\subsection{Methods}
-
-The style in this section should read as if you were verbally
-describing the conduct of the experiment. 
-The methods section is not a step-by-step, directive,
-protocol as you might see in your lab manual, but detailed enough such
-that an interested reader can reproduce your
-work~\cite{anderson04,wavelab}.
-The methods
-section should describe what was
-done to answer the research question, describe how it was done,
-justify the experimental design, and
-explain how the results were analyzed.
-
-The methods section of a research paper provides the information by
-which a study's validity is judged.
-Therefore, it requires a clear and precise description of how an
-experiment was done, and the rationale
-for why specific experimental procedures were chosen.
-It is usually helpful to
-structure the methods section by~\cite{kallet04methods}:
-\begin{enumerate}
-\item Describing the algorithms used in the study, briefly including
-  details such as hyperparameter values (e.g. thresholds), and
-  preprocessing steps (e.g. normalizing the data to have mean value of
-  zero).
-\item Explaining how the materials were prepared, for example the
-  images used and their resolution.
-\item Describing the research protocol, for example which examples
-  were used for estimating the parameters (training) and which were
-  used for computing performance.
-\item Explaining how measurements were made and what
-  calculations were performed. Do not reproduce the full source code in
-  the paper, but explain the key steps.
-\end{enumerate}
-
-\subsection{Results}
-
-Organize the results section based on the sequence of table and
-figures you include. Prepare the tables and figures as soon as all
-the data are analyzed and arrange them in the sequence that best
-presents your findings in a logical way. A good strategy is to note,
-on a draft of each table or figure, the one or two key results you
-want to addess in the text portion of the results.
-The information from the figures is
-summarized in Table~\ref{tab:fourier-wavelet}.
-
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}[c]{|l||l|l|l|}
-    \hline
-    Basis&Support&Suitable signals&Unsuitable signals\\
-    \hline
-    Fourier&global&sine like&localized\\
-    wavelet&local&localized&sine like\\
-    \hline
-  \end{tabular}
-  \caption{Characteristics of Fourier and wavelet basis.}
-  \label{tab:fourier-wavelet}
-\end{table*}
-
-When reporting computational or measurement results, always
-report the mean (average value) along with a measure of variablility
-(standard deviation(s) or standard error of the mean).
-
-\section{Tips for Good Software}
-\label{sec:tips-software}
-
-There is a lot of literature (for example~\cite{hunt99pragmatic} and
-\cite{spolsky04software}) on how to write software. It is not the
-intention of this section to replace software engineering
-courses. However, in the interests of reproducible
-research~\cite{schwab00}, there are a few guidelines to make your
-reader happy:
-\begin{itemize}
-\item Have a \texttt{README} file that (at least) describes what your
-  software does, and which commands to run to obtain results. Also
-  mention anything special that needs to be set up, such as
-  toolboxes\footnote{For those who are
-  particularly interested, other common structures can be found at
-  \url{http://en.wikipedia.org/wiki/README} and
-  \url{http://www.gnu.org/software/womb/gnits/}.}.
-\item A list of authors and contributors can be included in a file
-  called \texttt{AUTHORS}, acknowledging any help that you may have
-  obtained. For small projects, this information is often also
-  included in the \texttt{README}.
-\item Use meaningful filenames, and not \texttt{temp1.m},
-  \texttt{temp2.m}. The code should also unzip into a subdirectory.
-\item Document your code. Each file should at least have a short
-  description about its reason for existence. Non obvious steps in the
-  code should be commented.
-\item Describe how the results presented in your paper can potentially
-  be reproduced.
-\end{itemize}
-
-\section{Computational Intelligence Laboratory Requirements}
-\label{sec:cil}
-
-Your semester project is a group effort. It consists of four parts:
-\begin{enumerate}
-\item The programming assignments you solve during the semester.
-\item Developing a novel solution for one of the assignments, e.g. by
-  combining methods from previous programming assignments into a novel
-  solution.
-\item Comparing your novel solution to previous assignments.
-\item Writing up your findings in a short scientific paper.
-\end{enumerate}
-
-\subsection{Developing a Novel Solution}
-
-As your final programming assignment, you develop a novel solution to
-one of the four application problems. You are free to exploit any idea
-you have, provided it is not identical to any other group submission
-or existing Matlab implementation of an algorithm on the
-internet\footnote{\url{http://www.ethz.ch/students/semester/plagiarism_s_en.pdf}}.
-
-Two examples for developing a novel solution:
-\begin{itemize}
-\item You implemented a collaborative filtering algorithm based on
-  dimension reduction as part of an assignment. Now you apply
-  dimension reduction to inpainting.
-\item You implemented both a clustering and a sparse coding algorithm
-  for image compression. Now you combine both techniques into a novel
-  compression method.
-\end{itemize}
-
-\subsection{Comparison to Baselines}
-
-You compare your novel algorithm to \emph{at least two baseline
-  algorithms}. For the baselines, you can use the implementations you
-developed as part of the programming assignments.
-
-\subsection{Write Up}
-
-The submission must be in PDF form, using the \LaTeX{} template
-corresponding to the IEEE style of publication. Refer to
-Section~\ref{sec:latex-primer} for more information about preparing
-your document. The document should be a maximum of {\bf 4 pages}.
-
-\subsection{\LaTeX{} Primer}
-\label{sec:latex-primer}
-
-\LaTeX{} is one of the most commonly used document preparation systems
-for scientific journals and conferences. It is based on the idea
-that authors should be able to focus on the content of what they are
-writing without being distracted by its visual presentation.
-The source of this file can be used as a starting point for how to use
-the different commands in \LaTeX{}. We are using an IEEE style for
-this course.
-
-\subsubsection{Installation}
-
-There are various different packages available for processing \LaTeX{}
-documents.
-On Windows, use the Mik\TeX{} package (\url{http://miktex.org/}), and
-on OSX use MacTeX
-(\url{http://www.tug.org/mactex/2009/}). Alternatively, on OSX, you
-can install the \texttt{tetex} package via
-Fink\footnote{\url{http://www.finkproject.org/}} or 
-Macports\footnote{\url{http://www.macports.org/}}.
-
-\subsubsection{Compiling \LaTeX{}}
-Your directory should contain at least 4 files, in addition to image
-files. Images should be in \texttt{.png}, \texttt{.jpg} or
-\texttt{.pdf} format.
-\begin{itemize}
-\item IEEEtran.cls
-\item IEEEtran.bst
-\item groupXX-submission.tex
-\item groupXX-literature.bib
-\end{itemize}
-Note that you should replace groupXX with your chosen group name.
-Then, from the command line, type:
-\begin{verbatim}
-$ pdflatex groupXX-submission
-$ bibtex groupXX-literature
-$ pdflatex groupXX-submission
-$ pdflatex groupXX-submission
-\end{verbatim}
-This should give you a PDF document \texttt{groupXX-submission.pdf}.
-
-\subsubsection{Equations}
-
-There are three types of equations available: inline equations, for
-example $y=mx + c$, which appear in the text, unnumbered equations
-$$y=mx + c,$$
-which are presented on a line on its own, and numbered equations
-\begin{equation}
-  \label{eq:linear}
-  y = mx + c
-\end{equation}
-which you can refer to at a later point (Equation~(\ref{eq:linear})).
-
-\subsubsection{Tables and Figures}
-
-Tables and figures are ``floating'' objects, which means that the text
-can flow around it.
-Note
-that \texttt{figure*} and \texttt{table*} cause the corresponding
-figure or table to span both columns.
-
-\subsection{Grading}
-
-There are two different types of grading criteria applied to your
-project, with the corresponding weights shown in brackets.
-\begin{description}
-\item[Competitive] \ \\
-  The following criteria is scored based on your rank
-  in comparison with the rest of the class.
-  \begin{itemize}
-  \item time taken for computation (10\%)
-  \item average rank for all other criteria relevant to the task, for
-    example reconstruction error and sparsity (20\%)
-  \end{itemize}
-  The ranks will then be converted on a linear scale into a grade
-  between 4 and 6.
-\item[Non-competitive] \ \\
-  The following criteria is scored based on an
-  evaluation by the teaching assistants.
-  \begin{itemize}
-  \item quality of paper (30\%)
-  \item quality of implementation (20\%)
-  \item creativity of solution (20\%)
-  \end{itemize}
-\end{description}
-
-\subsection{Submission System}
-
-The deadline for submitting your project report is Friday, 25 June
-2010.
-You need to submit:
-\begin{itemize}
-\item PDF of paper.
-\item Archive (\texttt{.tar.gz} or \texttt{.zip}) of software. Please
-  do not forget to include author information in the source archive.
-\end{itemize}
-
 \section{Summary}
-
-The aim of a scientific paper is to convey the idea or discovery of
-the researcher to the minds of the readers. The associated software
-package provides the relevant details, which are often only briefly
-explained in the paper, such that the research can be reproduced.
-To write good papers, identify your key idea, make your contributions
-explicit, and use examples and illustrations to describe the problems
-and solutions.
+(((( TODO ))))
+  Summarise your contributions in light of the new
+  results.
 
 \section*{Acknowledgements}
-The author thanks Christian Sigg for his careful reading and helpful
-suggestions.
+(((( TODO ))))
 

### 2011-06-06 project/gaussInterpolate.m
Parameterized Inpainting with lots of code refactoring.

@@ -1,12 +1,9 @@
-function I_out = gaussInterpolate(I, mask)
+function I_out = gaussInterpolate(I, mask, parameters)
 
 mask = double(mask);
 
-sigma = 0.8;
-gauss_size = 5;
+h = fspecial('gaussian', parameters.gauss_size, parameters.gauss_sigma);
+I_masked = I .* mask;
 
-h = fspecial('gaussian', gauss_size, sigma);
-I_masked = I.*mask;
-
-I_filtered = imfilter(I_masked, h, 'replicate')./imfilter(mask, h, 'replicate');
-I_out = (I_filtered.*(1-mask))+(I_masked.*(mask));
+I_filtered = imfilter(I_masked, h, 'replicate') ./ imfilter(mask, h, 'replicate');
+I_out = (I_filtered .* (1-mask)) + (I_masked .* (mask));

### 2011-06-06 project/inPainting.m

@@ -1,79 +1,11 @@
-function I_rec = InPainting(I, mask)
+function I_rec = inPainting(I, mask)
 
-% Perform the actual inpainting of the image 
+parameters.gauss_size = 5;
+parameters.gauss_sigma = 0.8;
 
-% INPUT
-% I: (n x n) masked image
-% mask: (n x n) the mask hidding image information
-%
-% OUTPUT
-% I_rec = Reconstructed image 
+parameters.patch_size = 16;
+parameters.patch_frame = 8;
 
-% Parameters
-rc_min = 0.01; % rc_min: minimal residual correlation before stopping
-neib = 32; % neib: The patch sizes used in the decomposition of the image
-sigma = 0.01; % sigma: residual error stopping criterion, normalized by signal norm
+parameters.iterations = 4;
 
-patchsize=16; neighbourhood=8; t=88;
-
-% Get patches of size neib x neib from the image and the mask and
-% convert each patch to 1D signal
-I = gaussInterpolate(I, mask);
-%I_rec = I;
-%return;
-
-% Create overlapping patches
-numBlocks=size(I,1)/patchsize;
-
-if(mod(size(I,1),patchsize)~=0)
-  throw('Imagesize not multiple of pactchsize');
-end
-
-tic;
-
-IG = zeros(size(I,1)+2*neighbourhood, size(I,1)+2*neighbourhood);
-IG(neighbourhood+1 : neighbourhood+size(I,1), neighbourhood+1 : neighbourhood+size(I,1)) = I;
-
-%Fill borders
-for i=1:neighbourhood
-  IG(:,i) = IG(:,neighbourhood+1);
-  IG(:,end-i+1) = IG(:,end-neighbourhood);
-  IG(i,:) = IG(neighbourhood+1,:);
-  IG(end-i+1,:) = IG(end-neighbourhood,:);
-end
-
-% Create inverted mask (0=keep, 1=replace)
-mask = logical(1-mask);
-for i=1:numBlocks
-    for j=1:numBlocks
-      % Get block from neighbourhooded picture
-        block = IG( patchsize*(i-1)+1 : patchsize*i + 2*neighbourhood , patchsize*(j-1)+1:patchsize*j+2*neighbourhood);
-        F = fft2(block);
-        T = abs(F);
-
-        % perc = prctile(reshape(T,1,[]),t);
-        perc = 1;
-        F(abs(F) < perc) = 0;
-        % Set reconstructed part to be inverse of FFT
-        Whole = abs(ifft2(F));
-        
[…]

### 2011-06-06 project/inPaintingParameterized.m

@@ -1,0 +1,61 @@
+function I_rec = inPaintingParameterized(I, mask, parameters)
+
+% Perform the actual inpainting of the image.
+
+% INPUT
+% I: Masked image (n x n)
+% mask: Mask itself (n x n)
+% parameters: Parameters (as an object)
+%
+% OUTPUT
+% I_rec: Reconstructed image 
+
+% Interpolate the missing pixels with Gaussian weighting.
+I_rec = gaussInterpolate(I, mask, parameters);
+
+% Frame the original image by duplicating the first and last row respectively column.
+if mod(size(I, 1), parameters.patch_size) ~= 0
+	throw('The patch size does not divide the image size!'):
+end
+
+num_patches = size(I, 1) / parameters.patch_size;
+num_pixels = size(I, 1) + 2 * parameters.frame;
+
+I_framed = zeros(num_pixels, num_pixels);
+I_framed(parameters.frame + 1 : parameters.frame + size(I, 1), parameters.frame + 1 : parameters.frame + size(I, 1)) = I;
+
+for i = 1 : parameters.frame
+	I_framed(:, i) = I_framed(:, parameters.frame + 1);
+	I_framed(:, end - i + 1) = I_framed(:, end - parameters.frame);
+	I_framed(i, :) = I_framed(parameters.frame + 1, :);
+	I_framed(end - i + 1, :) = I_framed(end - parameters.frame, :);
+end
+
+% Invert the mask (0 = keep, 1 = replace)
+mask = logical(1 - mask);
+
+% Repeat the dimension reduction in the Fourier domain k times.
+for k = 1 : parameters.iterations
+	for i = 1 : num_patches
+		for j = 1 : num_patches
+			P_framed = I_framed(parameters.patch_size * (i - 1) + 1 : parameters.patch_size * i + 2 * parameters.frame, ...
+					parameters.patch_size * (j - 1) + 1 : parameters.patch_size * j + 2 * parameters.frame);
+			P_fft = fft2(P_framed);
+			P_abs = abs(P_fft);
+			
+			% perc = prctile(reshape(P_abs, 1, []), parameters.percentile);
+			perc = 1;
+			P_fft(abs(P_fft) < perc) = 0;
+			P_ifft = abs(ifft2(P_fft));
+			
+			P_rec = P_ifft(parameters.frame + 1 : end - parameters.frame, parameters.frame + 1 : end - parameters.frame);
+			
[…]

### 2011-06-06 project/inPaintingParameterized.m
Fix.

@@ -15,7 +15,7 @@ I_rec = gaussInterpolate(I, mask, parameters);
 
 % Frame the original image by duplicating the first and last row respectively column.
 if mod(size(I, 1), parameters.patch_size) ~= 0
-	throw('The patch size does not divide the image size!'):
+	throw('The patch size does not divide the image size!');
 end
 
 num_patches = size(I, 1) / parameters.patch_size;

### 2011-06-06 project/inPaintingParameterized.m
Fix 2.

@@ -19,16 +19,16 @@ if mod(size(I, 1), parameters.patch_size) ~= 0
 end
 
 num_patches = size(I, 1) / parameters.patch_size;
-num_pixels = size(I, 1) + 2 * parameters.frame;
+num_pixels = size(I, 1) + 2 * parameters.patch_frame;
 
 I_framed = zeros(num_pixels, num_pixels);
-I_framed(parameters.frame + 1 : parameters.frame + size(I, 1), parameters.frame + 1 : parameters.frame + size(I, 1)) = I;
+I_framed(parameters.patch_frame + 1 : parameters.patch_frame + size(I, 1), parameters.patch_frame + 1 : parameters.patch_frame + size(I, 1)) = I;
 
-for i = 1 : parameters.frame
-	I_framed(:, i) = I_framed(:, parameters.frame + 1);
-	I_framed(:, end - i + 1) = I_framed(:, end - parameters.frame);
-	I_framed(i, :) = I_framed(parameters.frame + 1, :);
-	I_framed(end - i + 1, :) = I_framed(end - parameters.frame, :);
+for i = 1 : parameters.patch_frame
+	I_framed(:, i) = I_framed(:, parameters.patch_frame + 1);
+	I_framed(:, end - i + 1) = I_framed(:, end - parameters.patch_frame);
+	I_framed(i, :) = I_framed(parameters.patch_frame + 1, :);
+	I_framed(end - i + 1, :) = I_framed(end - parameters.patch_frame, :);
 end
 
 % Invert the mask (0 = keep, 1 = replace)
@@ -38,8 +38,8 @@ mask = logical(1 - mask);
 for k = 1 : parameters.iterations
 	for i = 1 : num_patches
 		for j = 1 : num_patches
-			P_framed = I_framed(parameters.patch_size * (i - 1) + 1 : parameters.patch_size * i + 2 * parameters.frame, ...
-					parameters.patch_size * (j - 1) + 1 : parameters.patch_size * j + 2 * parameters.frame);
+			P_framed = I_framed(parameters.patch_size * (i - 1) + 1 : parameters.patch_size * i + 2 * parameters.patch_frame, ...
+					parameters.patch_size * (j - 1) + 1 : parameters.patch_size * j + 2 * parameters.patch_frame);
 			P_fft = fft2(P_framed);
 			P_abs = abs(P_fft);
 			
@@ -48,7 +48,7 @@ for k = 1 : parameters.iterations
[…]

### 2011-06-06 project/inPaintingParameterized.m
Fix 3.

@@ -55,7 +55,7 @@ for k = 1 : parameters.iterations
 			
 			mask_P = logical(zeros(size(mask)));
 			mask_P(range_i, range_j) = mask(range_i, range_j);
-			I_rec(mask_P) = P(mask(range_i, range_j));
+			I_rec(mask_P) = P_rec(mask(range_i, range_j));
 		end
 	end
 end

### 2011-06-06 project/Paper/paper.tex
Added rough draft of algorithm

@@
 \begin{document}
-\title{Image restoration using dimension reduction in the Fourier domain}
+\title{Image restoration using dimensionality reduction in the Fourier domain}
 
@@
 In image inpainting, the aim is to restore an image from a disturbed source.
-These distortions may, for example, result from a transmission over a noisy channel.
-We distinguish inpainting from image 
-((( TODO )))
+These disturbances may, for example, result from a transmission over a noisy channel or from a degraded physical copy of the original image.
+Another application is object removal where a part of the image to be removed and the hole to be filled using inpainting.\\
+In our approach, we use an iterative scheme with dimensionality reduction in the Fourier domain. 
+((( TODO )))\\
+
 The aim of writing a paper is to infect the mind of your reader with
@@
 ((( TODO )))
+
 \subsection{New approach}
+(((REWRITE)))
+We combine several ideas in our approach to obtain a inpainting algorithm that provides good results and fast performance.
+Our algorithm requires as input the image which shall be restored and a mask which indicates the pixels that shall be filled.
+Such a mask can be created by the user or automatically (((?
+checksum))).\\
+The main inspiration behind our algorithm is that we try to interpolate the pixels from its neighboring values.
+By this we add some noise compared to the original, undamaged image.
+We then try to use dimensionality reduction to filter out this generated noise to reconstruct the image.\\
+
+We use an iterative scheme to determine the right parameters used in our algorithm.
+For this, we use cross-validation: We randomly remove a percentage of pixels from the image as a test set.
+We then use this test set to test the accuracy of our parameters and determine the best set of parameters for this image.
+
+We then proceed to the main steps of the algorithm:
+\begin{enumerate}
+\label{algo}
+\item \label{algo:step1}
+We use a Gaussian to interpolate the missing pixels indicated by the mask, using a convolution operation. Both $\sigma$ and the size of the Gaussian matrix are determined by the cross-validation.
+This step is required by the Fourier transform which cannot ignore samples. (((Humm)))
+
+\item \label{algo:step2}
+We transform the image to the Fourier domain.
+In this domain we set the magnitudes of all frequencies to zero which are below the $n$-th percentile threshold.
+Again, $n$ is determined by cross-validation.
+(((EXPERIMENTAL: Validate, Rewrite))) We then repeat this step a certain number of times...
+
+\item \label{algo:step3}
+In this last step, we invert the Fourier transform and combine the reconstructed image with the undamaged part of the image, yielding our result.
+
+\end{enumerate}
+
 ((( TODO )))

### 2011-06-06 project/gaussInterpolate.m
+ Added online cross-validation

@@ -3,7 +3,7 @@ function I_out = gaussInterpolate(I, mask)
 mask = double(mask);
 
 sigma = 0.8;
-gauss_size = 5;
+gauss_size = 10;
 
 h = fspecial('gaussian', gauss_size, sigma);
 I_masked = I.*mask;

### 2011-06-06 project/inPainting.m

@@ -16,9 +16,25 @@ sigma = 0.01; % sigma: residual error stopping criterion, normalized by signal n
 
 patchsize=16; neighbourhood=8; t=88;
 
+validation_percentage = 0.5;
+
+% split into validation and test set
+good_indices = find(mask);
+num_good = length(good_indices);
+perm = randperm(num_good);
+val_indices = good_indices(perm(1:round(validation_percentage*num_good)));
+val_mask = zeros(size(mask));
+val_mask(val_indices) = 1;
+
+I_copy = I;
+I_copy(val_indices) = 0;
+
+mask_copy = mask;
+mask_copy(val_indices) = 0;
+
 % Get patches of size neib x neib from the image and the mask and
 % convert each patch to 1D signal
-I = gaussInterpolate(I, mask);
+I_copy = gaussInterpolate(I_copy, mask_copy);
 %I_rec = I;
 %return;
 
@@ -31,8 +47,8 @@ end
 
 tic;
 
-IG = zeros(size(I,1)+2*neighbourhood, size(I,1)+2*neighbourhood);
-IG(neighbourhood+1 : neighbourhood+size(I,1), neighbourhood+1 : neighbourhood+size(I,1)) = I;
+IG = zeros(size(I_copy,1)+2*neighbourhood, size(I_copy,1)+2*neighbourhood);
+IG(neighbourhood+1 : neighbourhood+size(I,1), neighbourhood+1 : neighbourhood+size(I,1)) = I_copy;
 
 %Fill borders
 for i=1:neighbourhood
@@ -43,28 +59,59 @@ for i=1:neighbourhood
 end
 
 % Create inverted mask (0=keep, 1=replace)
-mask = logical(1-mask);
+%mask = logical(1-mask);
 for i=1:numBlocks
     for j=1:numBlocks
       % Get block from neighbourhooded picture
         block = IG( patchsize*(i-1)+1 : patchsize*i + 2*neighbourhood , patchsize*(j-1)+1:patchsize*j+2*neighbourhood);
-        F = fft2(block);
-        T = abs(F);
+        
+        val_mask_block = val_mask(patchsize*(i-1)+1 : patchsize*i, patchsize*(j-1)+1:patchsize*j);
+        ground_truth = I(patchsize*(i-1)+1 : patchsize*i, patchsize*(j-1)+1:patchsize*j).*val_mask_block;
 
-        % perc = prctile(reshape(T,1,[]),t);
-        perc = 1;
[…]

### 2011-06-06 project/inPainting.m
+ Finished online cross-validation

@@ -35,6 +35,7 @@ mask_copy(val_indices) = 0;
 % Get patches of size neib x neib from the image and the mask and
 % convert each patch to 1D signal
 I_copy = gaussInterpolate(I_copy, mask_copy);
+I_orig_gauss = gaussInterpolate(I, mask);
 %I_rec = I;
 %return;
 
@@ -50,12 +51,21 @@ tic;
 IG = zeros(size(I_copy,1)+2*neighbourhood, size(I_copy,1)+2*neighbourhood);
 IG(neighbourhood+1 : neighbourhood+size(I,1), neighbourhood+1 : neighbourhood+size(I,1)) = I_copy;
 
+IG_orig = zeros(size(I,1)+2*neighbourhood, size(I,1)+2*neighbourhood);
+IG_orig(neighbourhood+1 : neighbourhood+size(I,1), neighbourhood+1 : neighbourhood+size(I,1)) = I_orig_gauss;
+
 %Fill borders
 for i=1:neighbourhood
   IG(:,i) = IG(:,neighbourhood+1);
   IG(:,end-i+1) = IG(:,end-neighbourhood);
   IG(i,:) = IG(neighbourhood+1,:);
   IG(end-i+1,:) = IG(end-neighbourhood,:);
+
+
+  IG_orig(:,i) = IG_orig(:,neighbourhood+1);
+  IG_orig(:,end-i+1) = IG_orig(:,end-neighbourhood);
+  IG_orig(i,:) = IG_orig(neighbourhood+1,:);
+  IG_orig(end-i+1,:) = IG_orig(end-neighbourhood,:);
 end
 
 % Create inverted mask (0=keep, 1=replace)
@@ -76,16 +86,10 @@ for i=1:numBlocks
 
         for perc=1:10
           F = fft2(block);
-
-          %before = length(find(F))
           F(abs(F) < perc) = 0;
-          %after = length(find(F))
 
           % Set reconstructed part to be inverse of FFT
           new_block = abs(ifft2(F));
-          %figure(2);
-          %imshow(new_block);
-
           new_block_small = new_block(neighbourhood+1:patchsize+neighbourhood, neighbourhood+1:patchsize+neighbourhood);
 
           diff = (val_mask_block.*new_block_small - ground_truth);
@@ -94,7 +98,7 @@ for i=1:numBlocks
 
           if (err < best_error)
             best_error = err;
-            best_block = new_block_small;
+            best_threshold = perc;
           end
[…]

### 2011-06-08 project/Paper/paper.tex
Small commit so I can switch branches

@@
 
+In our implementation, we use several tweaks to improve performance or quality:
+\begin{itemize}
+\item We use an extended neighborhood when ((( TODO )))
+\end{itemize}
+
+We use a 
 ((( TODO )))
@@
   due.
-
-\section{Results}
- ((( TODO )))
-  Show evidence to support your claims made in the
-  introduction.
 
@@
 \end{enumerate}
+
+\section{Results}
+ ((( TODO )))
+  Show evidence to support your claims made in the
+  introduction.
 

### 2011-06-08 project/EvaluateInpaintingParameterized.m
Not fully merged version.

@@ -1,0 +1,42 @@
+% Measure approximation error for several images.
+
+function error = EvaluateInpaintingParameterized(parameters)
+
+file_list = dir(); 
+k = 1;
+
+Errors = []; % mean squared errors for each image
+
+for i = 3 : length(dir) % running through the folder
+	
+	file_name = file_list(i).name; % get current filename
+	
+	% Only keep the images in the loop
+	if (length(file_name) < 5)
+		continue;
+	elseif ( max(file_name(end-4:end) ~= '2.png'))
+		continue;
+	end
+	mask_name = [file_name(1:end-4) '_mask.png'];
+		
+	% Read image, convert to double precision and map to [0,1] interval
+	I = imread(file_name);
+	I = double(I) / 255;
+	
+	% Read the respective binary mask
+	% EVALUATION IS DONE WITH A FIXED MASK
+	mask = imread(mask_name);
+	
+	I_mask = I;
+	I_mask(~mask) = 0;
+		  
+	% Call the main inPainting function
+	I_rec = inPainting(I_mask, mask);
+	
+	% Measure approximation error
+	Errors(k) = mean(mean(mean( ((I - I_rec) ).^2)));
+	
+	k = k+1;
+end
+
+error = mean(Errors);

### 2011-06-08 project/OptimizeInpainting.m

@@ -1,0 +1,30 @@
+% Optimize the parameters of inPainting.
+
+parameters.gauss_size = 5;
+parameters.gauss_sigma = 0.8;
+
+parameters.patch_size = 16;
+parameters.patch_frame = 8;
+
+parameters.iterations = 1;
+
+error_old = 1;
+error = EvaluateInpaintingParameterized(parameters);
+
+while abs(error_old - error) / error_old > 0.05
+	
+	parameters.gauss_size = 5;
+	parameters.gauss_sigma = 0.8;
+	
+	parameters.patch_size = 16;
+	parameters.patch_frame = 8;
+	
+	parameters.iterations = 1;
+	
+	error_old = error;
+	error = EvaluateInpaintingParameterized(parameters);
+	
+	disp(['Average quadratic error: ' error]);
+	disp(['Used parameters: ' parameters]);
+	
+end

### 2011-06-08 project/inPainting.m

@@ -6,6 +6,7 @@ parameters.gauss_sigma = 0.8;
 parameters.patch_size = 16;
 parameters.patch_frame = 8;
 
+parameters.validation = 0.5; % Fraction of pixels used for validating the threshold
 parameters.iterations = 4;
 
 I_rec = inPaintingParameterized(I, mask, parameters);

### 2011-06-08 project/inPaintingParameterized.m

@@ -10,32 +10,58 @@ function I_rec = inPaintingParameterized(I, mask, parameters)
 % OUTPUT
 % I_rec: Reconstructed image 
 
+% Split into validation and training set
+good_indices = find(mask);
+num_good = length(good_indices);
+perm = randperm(num_good);
+val_indices = good_indices(perm(1 : round(parameters.validation * num_good)));
+val_mask = zeros(size(mask));
+val_mask(val_indices) = 1;
+
+I_training = I;
+I_training(val_indices) = 0;
+
+mask_training = mask;
+mask_training(val_indices) = 0;
+
 % Interpolate the missing pixels with Gaussian weighting.
+I_training = gaussInterpolate(I_training, mask_training, parameters);
 I_rec = gaussInterpolate(I, mask, parameters);
 
-% Frame the original image by duplicating the first and last row respectively column.
-if mod(size(I, 1), parameters.patch_size) ~= 0
+image_size = size(I, 1);
+
+if mod(image_size, parameters.patch_size) ~= 0
 	throw('The patch size does not divide the image size!');
 end
 
-num_patches = size(I, 1) / parameters.patch_size;
-num_pixels = size(I, 1) + 2 * parameters.patch_frame;
+num_patches = image_size / parameters.patch_size;
+num_pixels = image_size + 2 * parameters.patch_frame;
 
-I_framed = zeros(num_pixels, num_pixels);
-I_framed(parameters.patch_frame + 1 : parameters.patch_frame + size(I, 1), parameters.patch_frame + 1 : parameters.patch_frame + size(I, 1)) = I;
-
-for i = 1 : parameters.patch_frame
-	I_framed(:, i) = I_framed(:, parameters.patch_frame + 1);
-	I_framed(:, end - i + 1) = I_framed(:, end - parameters.patch_frame);
-	I_framed(i, :) = I_framed(parameters.patch_frame + 1, :);
-	I_framed(end - i + 1, :) = I_framed(end - parameters.patch_frame, :);
-end
+I_rec_framed = zeros(num_pixels, num_pixels);
+I_training_framed = zeros(num_pixels, num_pixels);
 
 % Invert the mask (0 = keep, 1 = replace)
-mask = logical(1 - mask);
+imask = logical(1 - mask);
 
 % Repeat the dimension reduction in the Fourier domain k times.
 for k = 1 : parameters.iterations
+	
[…]

### 2011-06-09 project/Paper/paper.tex
Remove fill stuff from template

@@
 Our algorithm requires as input the image which shall be restored and a mask which indicates the pixels that shall be filled.
-Such a mask can be created by the user or automatically (((?
-checksum))).\\
+Such a mask can be created by the user or automatically (e.g.
+when a checksum algorithm detects faulty pixels).\\
 The main inspiration behind our algorithm is that we try to interpolate the pixels from its neighboring values.
@@
 We use an iterative scheme to determine the right parameters used in our algorithm.
-For this, we use cross-validation: We randomly remove a percentage of pixels from the image as a test set.
-We then use this test set to test the accuracy of our parameters and determine the best set of parameters for this image.
+For this, we use cross-validation: We randomly remove a percentage of pixels from the image as a validation set.
+(((STILL VALID?))) We then use this set to test the accuracy of our parameters and determine the best set of parameters for this image.
 
@@
 In this domain we set the magnitudes of all frequencies to zero which are below the $n$-th percentile threshold.
-Again, $n$ is determined by cross-validation.
-(((EXPERIMENTAL: Validate, Rewrite))) We then repeat this step a certain number of times...
+Again, our ideal cutoff $n$ is determined by cross-validation using a loop.
 
@@
 \begin{itemize}
-\item We use an extended neighborhood when ((( TODO )))
-\end{itemize}
-
-We use a 
+\item We perform the Fourier transform and thresholding on image patches to allow an individual threshold per patch.
+\item To avoid discontinuity at the boundary of the patches, we use a frame around the actual patch to capture more information.
+This frame is not filled in by the reconstructed patch from the Fourier step.
+(((IMAGE))) The frames of adjacent patches are overlapping, if the frame size is set to a value larger than zero.\end{itemize}
 ((( TODO )))
@@
 \section{Results}
+We evaluated two versions of our algorithm: An iterative variant and a non-iterative variant.
+Our iterative variant repeats 
  ((( TODO )))
@@
 
-\section{The Structure of a Paper}
-\label{sec:structure-paper}
-
-Scientific papers usually begin with the description of the problem,
-justifying why the problem is interesting. Most importantly, it argues
-that the problem is still unsolved, or that the current solutions are
-unsatisfactory. This leads to the main gist of the paper, which is
-``the idea''. The authors then show evidence, using derivations or
-experiments, that the idea works. Since science does not occur in a
-vacuum, a proper comparison to the current state of the art is often
-part of the results. Following these ideas, papers usually have the
-following structure:
-\begin{description}
-\item[Abstract] \ \\
-  Short description of the whole paper, to help the
-  reader decide whether to read it.
-\item[Introduction] \ \\
-  Describe your problem and state your
-  contributions.
-\item[Methods] \ \\
-  Describe your idea and how it was implemented to solve
-  the problem. Survey the related work, giving credit where credit is
-  due.
-\item[Results] \ \\
-  Show evidence to support your claims made in the
-  introduction.
-\item[Discussion] \ \\
-  Discuss the strengths and weaknesses of your
-  approach, based on the results. Point out the implications of your
-  novel idea on the application concerned.
-\item[Summary] \ \\
-  Summarise your contributions in light of the new
-  results.
-\end{description}
-
 \subsection{Figures and Tables}

### 2011-06-15 project/inPainting.m
Added abort criterion based on error improvement (iterative only)

@@ -7,7 +7,11 @@ parameters.patch_size = 16;
 parameters.patch_frame_size = 8;
 
 parameters.validation = 0.2; % Fraction of pixels used for validating the threshold
-parameters.iterative = true;
-parameters.max_iterations = 5;
+
+parameters.iterative = true;         % Enable or disable iterative scheme
+
+%Parameters for iterative scheme  
+parameters.max_iterations = 10;      % (Hard) abort criterion for iterations
+parameters.abortbelow_change = 0.05; % Minimal percentage change in error such that iterations continue. Tradeoff Speed-Accuracy
 
 I_rec = inPaintingParameterized(I, mask, parameters);

### 2011-06-15 project/inPaintingParameterized.m

@@ -37,12 +37,18 @@ mask_training_framed = createFrame(mask_training, parameters);
 
 if (parameters.iterative)
   % TODO add stopping criterion based on validation set
+  previous_error = +Inf;
   for i = 1:parameters.max_iterations
     [T, I_trained] = determineThresholds(I_training_framed, val_mask, I, parameters);
     I_training_framed = I_training_framed.*mask_training_framed + ...
                         (1-mask_training_framed).*I_trained;
     diff = (removeFrame(I_training_framed, parameters) - I).*val_mask;
     err = sum(sum(diff.*diff))
+    if(1/previous_error*err > 1-parameters.abortbelow_change)
+      sprintf('No significant improvement anymore. Ending iteration');
+      break;
+    end
+    previous_error=err;
   end
   I_final = I_training_framed;
 else

### 2011-06-15 project/Paper/paper.tex
Added image. Small stuff in Paper draft.

@@
 These disturbances may, for example, result from a transmission over a noisy channel or from a degraded physical copy of the original image.
-Another application is object removal where a part of the image to be removed and the hole to be filled using inpainting.\\
-In our approach, we use an iterative scheme with dimensionality reduction in the Fourier domain. 
+Another application is object removal where a part of the image to be removed and the hole to be filled using inpainting.
+(((EXPAND)))\\
+Our approach focuses on the first inpainting application.
+We use an iterative algorithm with dimensionality reduction in the Fourier domain paired with several adaptations to improve speed and reduce artifacts in the output.
 ((( TODO )))\\
@@
 \item \label{algo:step1}
-We use a Gaussian to interpolate the missing pixels indicated by the mask, using a convolution operation. Both $\sigma$ and the size of the Gaussian matrix are determined by the cross-validation.
+We use a Gaussian to interpolate the missing pixels indicated by the mask, using a convolution operation.
+Both $\sigma$ and the size of the Gaussian matrix are determined by the cross-validation.
+(((TODO?!)))
 This step is required by the Fourier transform which cannot ignore samples. (((Humm)))
@@
 \item To avoid discontinuity at the boundary of the patches, we use a frame around the actual patch to capture more information.
-This frame is not filled in by the reconstructed patch from the Fourier step.
-(((IMAGE))) The frames of adjacent patches are overlapping, if the frame size is set to a value larger than zero.\end{itemize}
+(Image \ref{fig:framing}) This frame is not filled in by the reconstructed patch from the Fourier step.
+The frames of adjacent patches are overlapping if the frame size is set to a value larger than zero.
+At the image boundaries, we replicate the outermost pixels to provide a sufficient frame.\end{itemize}
 ((( TODO )))
@@
 Our iterative variant repeats 
+ ((( TODO )))
+
+ We compare our algorithm with the following two baseline algorithms:
+ \begin{enumerate}
+ \item A simple inpainting algorithm using (((TODO )))
+ \item A matching pursuit algorithm with a dictionary consisting of (((TODO)))
+ \end{enumerate}
+
+ We tested our algorithm and the two baselines on a set of natural images and further on a set of artificially created ones. (((TODO More than 1)))
+
  ((( TODO )))
@@
   \centering
-  \includegraphics[width=\columnwidth]{images/denoised_signal_1d}
-  \caption{Signal compression and denoising using the Fourier basis.}
-  \label{fig:denoise-fourier}
+  \includegraphics[width=\columnwidth]{images/imageFraming}
+  \caption{Only the inner patch (marked red) is reconstructed in one step while the frame only provides additional signal. (Undistorted image shown) }
+  \label{fig:framing}
 \end{figure}
@@
 Use examples and illustrations to clarify ideas and results. For
-example, by comparing Figure~\ref{fig:denoise-fourier} and
+example, by comparing Figure~\ref{fig:framing} and
 Figure~\ref{fig:denoise-wavelet}, we can see the two different
@@
 \section*{Acknowledgements}
+We would like to recognize and thank the following individuals or groups for their contribution to this work:\\
+Our implementation is based on the inpainting and evaluation framework provided by the course "Computer Intelligence Lab" held by Prof. Joachim Buhmann at ETH Zürich.
 (((( TODO ))))

### 2011-06-16 project/OptimizeInpainting.m
+ Only use iterative parameters when needed

@@ -37,8 +37,10 @@ function optimizeInpainting()
     final_parameters.td_middle              = gradientDescent(7, @getNextTDMiddle, parameters, cost);
     final_parameters.validation             = gradientDescent(8, @getNextValidation, parameters, cost);
     %Dont iterate over bool 'iterative'
-    final_parameters.max_iterations         = gradientDescent(10, @getNextMaxIterations, parameters, cost);
-    final_parameters.abortbelow_change      = gradientDescent(11, @getNextAbortBelowChange, parameters, cost);
+    if parameters.iterative
+      final_parameters.max_iterations         = gradientDescent(10, @getNextMaxIterations, parameters, cost);
+      final_parameters.abortbelow_change      = gradientDescent(11, @getNextAbortBelowChange, parameters, cost);
+    end
   end
 
 end

### 2011-06-16 project/OptimizeInpainting.m
+ Uncomented keep old setting

@@ -77,17 +77,17 @@ function new_value = gradientDescent(index, getNext, parameters, cost);
   fprintf('cost %g\n', new_cost_minus);
 
   param_cell = struct2cell(parameters);
-  if(new_cost_plus > 0 && new_cost_minus > 0)
+%  if(new_cost_plus > 0 && new_cost_minus > 0)
       %Keep the old setting
-      fprintf('%s: %g --\n\n', fields{index}, param_cell{index})
+%      fprintf('%s: %g --\n\n', fields{index}, param_cell{index})
 
-      new_value = param_cell{index};
-  else
+%      new_value = param_cell{index};
+%  else
       stepsize = -1*(new_cost_plus-new_cost_minus)/2*learning_rate;
       new_value = getNext(param_cell{index}, stepsize);
       
       fprintf('%s: %g => %g (step %g)\n\n', fields{index}, param_cell{index}, new_value, stepsize)
-  end
+%  end
   %pause;
 end
 

### 2011-06-16 project/OptimizeInpainting.m
+ Better optimization starting values

@@ -14,8 +14,8 @@ function optimizeInpainting()
   parameters.td_abortbelow_stepsize = 2;
   parameters.td_middle = 10;
   parameters.validation = 0.2; 
-  parameters.iterative = true; 
-  parameters.max_iterations = 1; 
+  parameters.iterative = false; 
+  parameters.max_iterations = 5;
   parameters.abortbelow_change = 0.15; 
   
   % Result set

### 2011-06-16 project/Paper/paper.tex
Added new picture

@@
 \end{figure}
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\columnwidth]{images/local_wdenoised_1d}
-  \caption{Signal compression and denoising using the Daubechies wavelet basis.}
-  \label{fig:denoise-wavelet}
-\end{figure}
 
@@
 example, by comparing Figure~\ref{fig:framing} and
-Figure~\ref{fig:denoise-wavelet}, we can see the two different
+Figure~\ref{fig:flowchart}, we can see the two different
 situations where Fourier and wavelet basis perform well.
@@
 \bibliography{biblio}
+
+\clearpage
+\section{Appendix}
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=\textwidth]{images/flowchart}
+  \caption{The algorithm as a flow chart}
+  \label{fig:flowchart}
+\end{figure}
 \end{document}

### 2011-06-17 project/EvaluateInpaintingParameterized.m
Own parameters and cost function.

@@ -41,15 +41,10 @@ function cost = EvaluateInpaintingParameterized(parameters)
     k = k+1;
   end
   
-  sprintf('error: %f', mean(mean(Errors)));
-  sprintf('times: %f', mean(mean(Times)));
-
-  timemax = 90;
-  errormax = 0.001;
   global mincost
   global best_params
 
-  cost = exp(mean(Errors)/errormax) + exp(mean(Times)/timemax);
+  cost = 10000 * mean(Errors) + 0.01 * mean(Times);
   if (length(mincost) == 0 || cost < mincost)
     fprintf('\n-------------------------------\nFound new best params with cost %g', cost);
     mincost = cost;

### 2011-06-17 project/OptimizeInpainting.m

@@ -6,18 +6,17 @@ function optimizeInpainting()
 
   % Initial parameters
   parameters = struct;
-  parameters.gauss_size = 7.2;
-  parameters.gauss_sigma = 0.8;
-  parameters.patch_size = 7.5;
-  parameters.patch_frame_size = 7.5;
-  parameters.td_abortbelow_stdev = 2.2;
-  parameters.td_abortbelow_stepsize = 1.2;
-  parameters.td_middle = 18.5;
-  parameters.validation = 0.2; 
+  parameters.gauss_size = 9.3;
+  parameters.gauss_sigma = 0.7;
+  parameters.patch_size = 7.9;
+  parameters.patch_frame_size = 8.9;
+  parameters.td_abortbelow_stdev = 3.8;
+  parameters.td_abortbelow_stepsize = 2.3;
+  parameters.td_middle = 8.1;
+  parameters.validation = 0.05; 
   parameters.iterative = true; 
-  parameters.max_iterations = 5;
-  parameters.abortbelow_change = 0.15; 
-
+  parameters.max_iterations = 4.9; 
+  parameters.abortbelow_change = 0.18; 
   old = zeros(11, 1);
   
   % Result set
@@ -48,7 +47,7 @@ function optimizeInpainting()
 end
 
 function [new_value, next_old] = gradientDescent(index, getNext, parameters, old, cost);
-  learning_rate = 30;
+  learning_rate = 5;
   alpha = 0.9;
 
   fields = {'gauss_size', ...

### 2011-06-17 project/Paper/paper.tex
+ Added subsections
Added modified flowchart dia. Insert flowchart into paper

@@
 \item We perform the Fourier transform and thresholding on image patches to allow an individual threshold per patch.
-\item To avoid discontinuity at the boundary of the patches, we use a frame around the actual patch to capture more information.
+\item To avoid discontinuity (Image \ref{fig:boundaryArtifacts}) at the boundary of the patches, we use a frame around the actual patch to capture more information.
 (Image \ref{fig:framing}) This frame is not filled in by the reconstructed patch from the Fourier step.
@@
   \centering
+  \includegraphics[width=\columnwidth]{images/boundaryArtifact_noframe}
+  \caption{Artifacts that occur without a frame (frame size = 0). The patches are recognizable in the reconstruction}
+  \label{fig:boundaryArtifacts}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
   \includegraphics[width=\columnwidth]{images/imageFraming}
@@
 \end{figure}
+
+((( TODO: Plot missingPixel\% vs. accuracy (vs. speed?) )))
 

### 2011-06-17 project/OptimizeInpainting.m
Modify to create mask at random. Modify to allow for arbitrary number of missing pixels in mask

@@ -1,22 +1,24 @@
 % Optimize the parameters of inPainting.
 function optimizeInpainting()
 
+  missing_pixels = 0.6;
+
   % Set random seet to get reproducable results
   rand('seed', 12345);
 
   % Initial parameters
   parameters = struct;
-  parameters.gauss_size = 9.3;
-  parameters.gauss_sigma = 0.7;
-  parameters.patch_size = 7.9;
-  parameters.patch_frame_size = 8.9;
-  parameters.td_abortbelow_stdev = 3.8;
-  parameters.td_abortbelow_stepsize = 2.3;
-  parameters.td_middle = 8.1;
-  parameters.validation = 0.05; 
+  parameters.gauss_size = 5;
+  parameters.gauss_sigma = 2;
+  parameters.patch_size = 5;
+  parameters.patch_frame_size = 4;
+  parameters.td_abortbelow_stdev = 5;
+  parameters.td_abortbelow_stepsize = 5;
+  parameters.td_middle = 4;
+  parameters.validation = 0.5; 
   parameters.iterative = true; 
-  parameters.max_iterations = 4.9; 
-  parameters.abortbelow_change = 0.18; 
+  parameters.max_iterations = 8; 
+  parameters.abortbelow_change = 0.32; 
   old = zeros(11, 1);
   
   % Result set
@@ -25,7 +27,7 @@ function optimizeInpainting()
   while(true)
     fprintf('========== Starting new round ===========\n')
     parameters = final_parameters
-    cost = EvaluateInpaintingParameterized(parameters);
+    cost = EvaluateInpaintingParameterized(parameters, missing_pixels);
     fprintf('=> cost %g\n\n', cost);
 
     % gauss_size
@@ -67,7 +69,7 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   param_cell{index} = getNext(param_cell{index}, 1);
   new_parameters = cell2struct(param_cell, fields, 1);
   fprintf('%s: %g ... ', fields{index}, param_cell{index})
-  new_cost_plus = EvaluateInpaintingParameterized(new_parameters) - cost;
+  new_cost_plus = EvaluateInpaintingParameterized(new_parameters, missing_pixels) - cost;
   fprintf('cost %g\n', new_cost_plus);
 
   % Case -1
@@ -75,7 +77,7 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
[…]

### 2011-06-17 project/EvaluateInpaintingParameterized.m
+ Loading and saving of best parameters

@@ -40,15 +40,16 @@ function cost = EvaluateInpaintingParameterized(parameters)
 
     k = k+1;
   end
-  
-  global mincost
-  global best_params
+ 
+  global global_best_cost 
 
   cost = 10000 * mean(Errors) + 0.01 * mean(Times);
-  if (length(mincost) == 0 || cost < mincost)
-    fprintf('\n-------------------------------\nFound new best params with cost %g', cost);
-    mincost = cost;
-    best_params = parameters
-    fprintf('-------------------------------\n\n')
+  if (cost < global_best_cost)
+    fprintf('\n-------------------------------\nFound new best params with cost %g\n', cost);
+    global_best_cost = cost;
+    best_cost = cost;
+    parameters
+    save('params.mat', 'best_cost', 'parameters');
+    fprintf('-------------------------------\n')
   end
 end

### 2011-06-17 project/OptimizeInpainting.m

@@ -4,19 +4,28 @@ function optimizeInpainting()
   % Set random seet to get reproducable results
   rand('seed', 12345);
 
+  % Load parameters and best_cost from (binary) file
+  load('params.mat');
+
+  % Store best value in global variable, for usage in Evaluation script
+  global global_best_cost;
+  global_best_cost = best_cost;
+  fprintf('\nStarting optimization, current best cost is %g\n\n', best_cost);
+
   % Initial parameters
-  parameters = struct;
-  parameters.gauss_size = 9.3;
-  parameters.gauss_sigma = 0.7;
-  parameters.patch_size = 7.9;
-  parameters.patch_frame_size = 8.9;
-  parameters.td_abortbelow_stdev = 3.8;
-  parameters.td_abortbelow_stepsize = 2.3;
-  parameters.td_middle = 8.1;
-  parameters.validation = 0.05; 
-  parameters.iterative = true; 
-  parameters.max_iterations = 4.9; 
-  parameters.abortbelow_change = 0.18; 
+  % parameters = struct;
+  % parameters.gauss_size = 9.3;
+  % parameters.gauss_sigma = 0.7;
+  % parameters.patch_size = 7.9;
+  % parameters.patch_frame_size = 8.9;
+  % parameters.td_abortbelow_stdev = 3.8;
+  % parameters.td_abortbelow_stepsize = 2.3;
+  % parameters.td_middle = 8.1;
+  % parameters.validation = 0.05; 
+  % parameters.iterative = true; 
+  % parameters.max_iterations = 4.9; 
+  % parameters.abortbelow_change = 0.18; 
+
   old = zeros(11, 1);
   
   % Result set

### 2011-06-17 project/inPainting.m

@@ -1,26 +1,7 @@
 function I_rec = inPainting(I, mask)
+  % Load parameters from file
+  load('params.mat');
 
-parameters.gauss_size = 8.7392;
-parameters.gauss_sigma = 0.6526;
-
-% Is the exponent of the power of two
-% due to the optimize script
-parameters.patch_size = 7.4459;
-parameters.patch_frame_size = 8.7963;
-
-% Parameters for determining thresholds in Fourier
-% Is the negative exponent of the power of 10
-% due to the optimize script
-parameters.td_abortbelow_stdev = 0.2479;
-parameters.td_abortbelow_stepsize = 2.6688;
-
-parameters.td_middle = 6.7928;
-parameters.validation = 0.0518; % Fraction of pixels used for validating the threshold
-
-parameters.iterative = true;         % Enable or disable iterative scheme
-
-% Parameters for iterative scheme  
-parameters.max_iterations = 4.1118;      % (Hard) abort criterion for iterations
-parameters.abortbelow_change = 0.1948; % Minimal percentage change in error such that iterations continue. Tradeoff Speed-Accuracy
-
-I_rec = inPaintingParameterized(I, mask, parameters);
+  % Start parametrized inpainting
+  I_rec = inPaintingParameterized(I, mask, parameters);
+end

### 2011-06-17 project/EvaluateInpaintingParameterized.m
Missing pixels and error again output

@@ -1,6 +1,6 @@
 % Measure approximation error for several images.
 
-function cost = EvaluateInpaintingParameterized(parameters, missing_pixels_fract)  
+function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_pixels_fract)  
 
   file_list = dir(); 
   k = 1;
@@ -43,6 +43,8 @@ function cost = EvaluateInpaintingParameterized(parameters, missing_pixels_fract
 
     k = k+1;
   end
+  
+  avgQErr = mean(Errors);
   
   global mincost
   global best_params

### 2011-06-17 project/OptimizeInpainting.m

@@ -27,7 +27,7 @@ function optimizeInpainting()
   while(true)
     fprintf('========== Starting new round ===========\n')
     parameters = final_parameters
-    cost = EvaluateInpaintingParameterized(parameters, missing_pixels);
+    [cost, unused] = EvaluateInpaintingParameterized(parameters, missing_pixels);
     fprintf('=> cost %g\n\n', cost);
 
     % gauss_size
@@ -69,7 +69,8 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   param_cell{index} = getNext(param_cell{index}, 1);
   new_parameters = cell2struct(param_cell, fields, 1);
   fprintf('%s: %g ... ', fields{index}, param_cell{index})
-  new_cost_plus = EvaluateInpaintingParameterized(new_parameters, missing_pixels) - cost;
+  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels);
+  new_cost_plus = new_cost - cost;
   fprintf('cost %g\n', new_cost_plus);
 
   % Case -1
@@ -77,7 +78,8 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   param_cell{index} = getNext(param_cell{index}, -1);
   new_parameters = cell2struct(param_cell, fields, 1);
   fprintf('%s: %g ... ', fields{index}, param_cell{index})
-  new_cost_minus = EvaluateInpaintingParameterized(new_parameters, missing_pixels) - cost;
+  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels);
+  new_cost_minus = new_cost - cost;
   fprintf('cost %g\n', new_cost_minus);
 
   param_cell = struct2cell(parameters);

### 2011-06-17 project/OptimizeInpainting.m
Fix compile error & Warning in the function name

@@ -1,5 +1,5 @@
 % Optimize the parameters of inPainting.
-function optimizeInpainting()
+function OptimizeInpainting()
 
   missing_pixels = 0.6;
 
@@ -39,24 +39,24 @@ function optimizeInpainting()
     fprintf('=> cost %g\n\n', cost);
 
     % gauss_size
-    [final_parameters.gauss_size, old]             = gradientDescent(1, @getNextGaussSize, parameters, old, cost);
-    [final_parameters.gauss_sigma, old]            = gradientDescent(2, @getNextGaussSigma, parameters, old, cost);
-    [final_parameters.patch_size, old]             = gradientDescent(3, @getNextPatchSize, parameters, old, cost);
-    [final_parameters.patch_frame_size, old]       = gradientDescent(4, @getNextFrameSize, parameters, old, cost);
-    [final_parameters.td_abortbelow_stdev, old]    = gradientDescent(5, @getNextTDAbortBelowStdev, parameters, old, cost);
-    [final_parameters.td_abortbelow_stepsize, old] = gradientDescent(6, @getNextTDAbortBelowStep, parameters, old, cost);
-    [final_parameters.td_middle, old]              = gradientDescent(7, @getNextTDMiddle, parameters, old, cost);
-    [final_parameters.validation, old]             = gradientDescent(8, @getNextValidation, parameters, old, cost);
+    [final_parameters.gauss_size, old]             = gradientDescent(1, @getNextGaussSize, parameters, old, cost, missing_pixels);
+    [final_parameters.gauss_sigma, old]            = gradientDescent(2, @getNextGaussSigma, parameters, old, cost, missing_pixels);
+    [final_parameters.patch_size, old]             = gradientDescent(3, @getNextPatchSize, parameters, old, cost, missing_pixels);
+    [final_parameters.patch_frame_size, old]       = gradientDescent(4, @getNextFrameSize, parameters, old, cost, missing_pixels);
+    [final_parameters.td_abortbelow_stdev, old]    = gradientDescent(5, @getNextTDAbortBelowStdev, parameters, old, cost, missing_pixels);
+    [final_parameters.td_abortbelow_stepsize, old] = gradientDescent(6, @getNextTDAbortBelowStep, parameters, old, cost, missing_pixels);
+    [final_parameters.td_middle, old]              = gradientDescent(7, @getNextTDMiddle, parameters, old, cost, missing_pixels);
[…]

### 2011-06-17 project/EvaluateInpaintingParameterized.m
Fix funny bug

@@ -26,10 +26,11 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
     % Generate mask at random, with 60% missing pixels
     mask = ones(size(I));
     numpixels = size(I,1)*size(I,2);
-    holes = randperm(floor(missing_pixels_fract*numpixels));
+    holes = randperm(numpixels);
+    holes = holes(1:floor(missing_pixels_fract*numpixels));
     mask(holes) = 0;
     
-    
+
     I_mask = I;
     I_mask(~mask) = 0;
 

### 2011-06-17 project/OptimizeInpainting.m
+ When old_stepsize was zero, do not use momentum!

@@ -99,7 +99,11 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
 %  else
       stepsize = -1*(new_cost_plus-new_cost_minus)/2*learning_rate;
       old_stepsize = old(index);
-      new_stepsize = old_stepsize*alpha + (1-alpha)*stepsize;
+      if (old_stepsize ~= 0)
+        new_stepsize = old_stepsize*alpha + (1-alpha)*stepsize;
+      else
+        new_stepsize = stepsize;
+      end
       fprintf('steps(old/new/together): %g/%g/%g\n', old_stepsize, stepsize, new_stepsize)
       old(index) = new_stepsize;
       new_value = getNext(param_cell{index}, new_stepsize);

### 2011-06-17 project/CreatePaperPlots.m
Added script to generate plots

@@ -1,0 +1,28 @@
+% Create Error vs. MissingPixels Plot
+% Requires parameters saved in a parameters.mat file
+
+load('params.mat');
+
+  stepsize = 2;
+  no_steps = floor(100/stepsize);
+  error = zeros(no_steps, 1);
+
+  %Use parallel computation for this
+  parfor k=0:no_steps
+
+    [unused, e] = EvaluateInpaintingParameterized(parameters,stepsize*k/100);
+    % Normalized
+    error(k+1) = e/(k*stepsize/100);
+    
+  end
+  
+  handle = figure;
+  hold on;
+  xlabel('Missing pixels (%)');
+  ylabel('Average squared error per missing pixel');
+  plot(0:stepsize:100, error);
+  saveas(handle, 'plots/missingpixelsVsError.png');
+  hold off;
+  
+  
+  

### 2011-06-17 project/OptimizeInpainting.m
Best parameters with repeated evaluation.

@@ -58,7 +58,7 @@ end
 
 function [new_value, next_old] = gradientDescent(index, getNext, parameters, old, cost, missing_pixels);
   learning_rate = 5;
-  alpha = 0.9;
+  alpha = 0.5;
 
   fields = {'gauss_size', ...
             'gauss_sigma', ...

### 2011-06-17 project/Paper/paper.tex
Introduction of paper.

--- 2011-06-15 project/Paper/paper.tex
@@
 \usepackage{url}
-\usepackage{graphicx}	% For figure environment
+\usepackage{graphicx}
 \usepackage[pdftex]{hyperref}
-\usepackage[utf8]{inputenc} % To make umlauts and other symbols correct
+\usepackage[utf8]{inputenc}
 
 \begin{document}
-\title{Image restoration using dimensionality reduction in the Fourier domain}
+\title{Inpainting with dimensionality reduction in the Fourier domain}
 
 \author{
-  Patrick Bänziger, Kaspar Etter, Jan Rüegg \\
+  Patrick Bänziger, Kaspar Etter, Jan Rüegg\\
   Department of Computer Science, ETH Zurich, Switzerland
@@
 \section{Introduction}
-In image inpainting, the aim is to restore an image from a disturbed source.
-These disturbances may, for example, result from a transmission over a noisy channel or from a degraded physical copy of the original image.
-Another application is object removal where a part of the image to be removed and the hole to be filled using inpainting.
-(((EXPAND)))\\
-Our approach focuses on the first inpainting application.
+Inpainting is about reconstructing missing spots of an image by guessing the correct values.
+The quality of such a restoration method can be evaluated by masking an image and comparing the result with the original.\\
+Our approach is based on the assumption that (natural) images have some underlying characteristic that can be distinguished from noise.
+The computed pixels can be viewed as a disturbed version of the genuine ones, hence the goal is to get rid of this noise.
+We do so by analyzing image patches in the Fourier space, removing the weak frequencies that capture the noise and thereby recovering the intrinsic signal.\\
 We use an iterative algorithm with dimensionality reduction in the Fourier domain paired with several adaptations to improve speed and reduce artifacts in the output.
-((( TODO )))\\
-
-The aim of writing a paper is to infect the mind of your reader with
-the brilliance of your idea \cite{jones08}. 
-The hope is that after reading your
-paper, the audience will be convinced to try out your idea. In other
-words, it is the medium to transport the idea from your head to your
-reader's head.
 
@@
 \end{enumerate}
+
+\subsection{Online threshold estimation}
+In the algorithm as described above, we have many different parameters that are required to be tuned
+
+\subsection{Offline parameter estimation}
 

### 2011-06-17 project/Paper/paper.tex

@@
 \end{figure}
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\columnwidth]{images/local_wdenoised_1d}
-  \caption{Signal compression and denoising using the Daubechies wavelet basis.}
-  \label{fig:denoise-wavelet}
-\end{figure}
+
+((( TODO: Plot missingPixel\% vs. accuracy (vs. speed?) )))
 

### 2011-06-17 project/Paper/paper.tex
+ Started work on gradient descent description

@@
 \subsection{Offline parameter estimation}
+As shown in table \ref{parameters}, there are a total of 10 parameters in our algorithm that can be set.
+For many of them, it is quite clear what values should be used. For example, the size of the gauss kernel
+should be a bit larger than the biggest assumed hole in the image. For others, like the maximum number of
+iterations, it is not so clear. And yet others have a clear speed / accuracy tradeoff, like the stopping
+conditions of the online threshold estimation.
+
+Because we thought it not feasable to tune all the parameters by hand, we wrote a program to
+do the offline parameter estimation. The idea is to do this once for a few representative images, and get
+parameters that can model as closely as possible many of the images we are interested in.
+Our proposed method combines a gradient descent method with ideas from cross-validation, and tunes our
+10-dimensional parameter-space, maximising time and accuracy.
+
+\subsubsection{Error function}
+
+\subsubsection{Gradient descent}
 

### 2011-06-17 project/EvaluateInpaintingParameterized.m
Parallelized Evaluation loop

@@ -3,13 +3,17 @@
 function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_pixels_fract)  
 
   file_list = dir(); 
-  k = 1;
+  rep=3;
+  
+  dir_length = length(dir);
 
-  Errors = []; % mean squared errors for each image
-  Times = [];
+  Errors_final = [];
+  Times_final = [];
 
-  for r = 1 : 3
-     for i = 3 : length(dir) % running through the folder
+  parfor r = 1 : rep
+     Errors = []; % mean squared errors for each image
+     Times = [];
+     for i = 3 : dir_length  % running through the folder
 
         file_name = file_list(i).name; % get current filename
 
@@ -38,20 +42,20 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
         % Call the main inPainting function
         tic;
         I_rec = inPaintingParameterized(I_mask, mask, parameters);
-        Times(k) = toc;
+        Times = [Times toc];
 
         % Measure approximation error
-        Errors(k) = mean(mean(mean( ((I - I_rec) ).^2)));
-
-        k = k+1;
+        Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
      end
+     Errors_final = [Errors_final Errors];
+     Times_final = [Times_final Times];
   end
   
-  avgQErr = mean(Errors);
+  avgQErr = mean(Errors_final);
  
   global global_best_cost 
 
-  cost = 10000 * mean(Errors) + 0.01 * mean(Times);
+  cost = 10000 * mean(Errors_final) + 0.001 * mean(Times_final);
   if (cost < global_best_cost)
     fprintf('\n-------------------------------\nFound new best params with cost %g\n', cost);
     global_best_cost = cost;

### 2011-06-17 project/Paper/paper.tex
Introduction of the paper.

--- 2011-06-17 project/Paper/paper.tex
@@
 The computed pixels can be viewed as a disturbed version of the genuine ones, hence the goal is to get rid of this noise.
-We do so by analyzing image patches in the Fourier space, removing the weak frequencies that capture the noise and thereby recovering the intrinsic signal.\\
-We use an iterative algorithm with dimensionality reduction in the Fourier domain paired with several adaptations to improve speed and reduce artifacts in the output.
+We do so by analyzing image patches in the Fourier space, removing the weak frequencies that (hopefully) capture the noise and thereby recovering the intrinsic signal.\\
+We repeat this procedure with the newly determined values until the error of a randomly chosen validation set converges or the maximum number of iterations is reached.
+For a faster start, we initialize our reconstructed image with what we call Gaussian interpolation: Using MATLAB's filtering function, we compute a weighted sum over all the neighboring pixels that are known.
+In order to determine the best threshold for suppressing weaker frequencies, we randomly remove even more pixels from the image and use them as a validation set.
 

### 2011-06-20 project/EvaluateInpaintingParameterized.m
Use final versions of error and times

@@ -55,13 +55,13 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
  
   global global_best_cost 
 
-  if (mean(Times) > 60)
-    times_error = exp((mean(Times)/60)-1)
+  if (mean(Times_final) > 60)
+    times_error = exp((mean(Times_final)/60)-1);
   else
     times_error = 0;
   end
 
-  cost = 10000 * mean(Errors) + times_error;
+  cost = 10000 * mean(Errors_final) + times_error;
   if (cost < global_best_cost)
     fprintf('\n-------------------------------\nFound new best params with cost %g\n', cost);
     global_best_cost = cost;

### 2011-06-20 project/Paper/paper.tex
Paper: File extensions.
+ Work on paper

@@
 %\usepackage{balance}
+\usepackage{amsmath}
 \usepackage{url}
@@
 
-Because we thought it not feasable to tune all the parameters by hand, we wrote a program to
+Because we thought it not feasable to tune all the parameters by hand, we wrote an algorithm to
 do the offline parameter estimation. The idea is to do this once for a few representative images, and get
@@
 \subsubsection{Error function}
+When doing a minimization of any kind, an error function is needed. Of course we already had the error function
+of the evaluation script, that gives a mean over the squared differences of the reconstructed to the original
+image. However, using only an error measure, the time would potentially grow to infinity to get the best accuracy possible during tuning.
+
+Therefore we tried different error function including both, the image error $E$, and the time $T$ spent. The image error here is the average
+squared error of a reconstructed pixel to the original pixel, scaled with a factor (usually about 10000), while the time is measured in seconds.
+
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|p{5cm}|p{5cm}|}
+    \hline
+    Function&Advantages&Disadvantages\\
+    \hline
+    $\exp(E)+\exp(T)$ & Increases cost fast to avoid $E$ or $T$ over a certain threshold & Not enough gradient for very low $E$, $T$ values\\
+    \hline
+    $(1-\exp(E))*(1-\exp(T))$ & Makes it possible to have large times for very low errors & Works also the other way around, and optimization only minimizes time\\
+    \hline
+    $E + T$ & Everywhere a good gradient & Time from 60 to 59 seconds is same \textit{improvement} as time from 2 to 1 second\\
+    \hline
+    $\begin{cases}
+  \exp(E)  & \text{if }T<60\\
+  \exp(E)+\exp(T/60 - 1) - 1 & \text{else}
+\end{cases}$ & Optimizes for error only when time is low enough. Function still continous, penalizes high times. & None so far, this is the function used by us\\
+    \hline
+  \end{tabular}
+  \caption{Different combined error functions}
+  \label{error_function_table}
+\end{table*}
 
@@
 
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}[c]{|l||l|l|l|}
-    \hline
-    Basis&Support&Suitable signals&Unsuitable signals\\
-    \hline
-    Fourier&global&sine like&localized\\
-    wavelet&local&localized&sine like\\
-    \hline
-  \end{tabular}
-  \caption{Characteristics of Fourier and wavelet basis.}
-  \label{tab:fourier-wavelet}
-\end{table*}
-
 When reporting computational or measurement results, always
@@
   \centering
-  \includegraphics[width=\columnwidth]{images/boundaryArtifact_noframe}
+  \includegraphics[width=\columnwidth]{images/boundaryArtifact_noframe.png}
   \caption{Artifacts that occur without a frame (frame size = 0). The patches are recognizable in the reconstruction}
@@
   \centering
-  \includegraphics[width=\columnwidth]{images/imageFraming}
+  \includegraphics[width=\columnwidth]{images/imageFraming.png}
   \caption{Only the inner patch (marked red) is reconstructed in one step while the frame only provides additional signal. (Undistorted image shown) }

### 2011-06-20 project/EvaluateInpaintingParameterized.m
! Fixed cost function

@@ -56,7 +56,7 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
   global global_best_cost 
 
   if (mean(Times) > 60)
-    times_error = exp((mean(Times)/60)-1)
+    times_error = exp((mean(Times)/60)-1) - 1
   else
     times_error = 0;
   end

### 2011-06-20 project/Paper/paper.tex
Small stuff to allow pull again
+ Added gradient descent to paper

@@
 
-Therefore we tried different error function including both, the image error $E$, and the time $T$ spent. The image error here is the average
+Therefore we tried different error function including both, the image error $E$, and the time $T$ spent (see table \ref{error_function_table}). The image error here is the average
 squared error of a reconstructed pixel to the original pixel, scaled with a factor (usually about 10000), while the time is measured in seconds.
@@
 
+% TODO maybe add 3d plot for some of the error functions
+
 \subsubsection{Gradient descent}
+The idea behind our gradient descent is to optimize all the parameters at once, instead of fine-tuning each of them separately. For that, we
+calculate the derivative of the error function depending on one of our 10 parameters, and then move the parameter into the direction where the
+error decreases fastest. We also consider the slope, and make big changes if the derivative is big, and small changes if the derivative is small.
+Iteratively, we calculate all of the 10 derivatives, change all the parameters and start over again.
+
+Since the error and time are measured, we cannot algebraically determine the gradient of the error function. Instead, we do a discrete estimation,
+measuring the error with finite differences. The error for parameter values plus and minus a small delta, divided by two, gives us our estimated
+derivative.
+
+In figure \ref{}
 

### 2011-06-20 project/Paper/paper.tex
Paper: New Approach.

--- 2011-06-17 project/Paper/paper.tex
@@
 \begin{document}
-\title{Inpainting with dimensionality reduction in the Fourier domain}
+\title{Inpainting with controlled dimensionality reduction in the Fourier domain}
 
@@
 \section{Introduction}
-Inpainting is about reconstructing missing spots of an image by guessing the correct values.
+Inpainting is about reconstructing the missing spots of an image by guessing the correct values.
 The quality of such a restoration method can be evaluated by masking an image and comparing the result with the original.\\
@@
 For a faster start, we initialize our reconstructed image with what we call Gaussian interpolation: Using MATLAB's filtering function, we compute a weighted sum over all the neighboring pixels that are known.
-In order to determine the best threshold for suppressing weaker frequencies, we randomly remove even more pixels from the image and use them as a validation set.
+In order to determine the best threshold for suppressing weaker frequencies, we randomly remove even more pixels from the image and use them as a validation set (for which we know the correct values).
 
@@
 \subsection{Related work}
-((( TODO )))
+Without actually knowing the broad field of research about inpainting (also known as image interpolation), we could not find any papers covering similar concepts as ours.
+Most methods are based on partial differential equations or some form of texture synthesis.
 
 \subsection{New approach}
-(((REWRITE)))
-We combine several ideas in our approach to obtain a inpainting algorithm that provides good results and fast performance.
-Our algorithm requires as input the image which shall be restored and a mask which indicates the pixels that shall be filled.
-Such a mask can be created by the user or automatically (e.g.
-when a checksum algorithm detects faulty pixels).\\
-The main inspiration behind our algorithm is that we try to interpolate the pixels from its neighboring values.
-By this we add some noise compared to the original, undamaged image.
-We then try to use dimensionality reduction to filter out this generated noise to reconstruct the image.\\
+As outlined in the introduction, we combined several ideas for obtaining an inpainting algorithm that provides both good results and fast performance.
+The input of our algorithm is the image to be restored together with a mask that indicates its unknown spots.\\
+Before entering the iterative part of our algorithm, the following three steps are performed (as seen in figure ??):
+\begin{enumerate}
+\item In order to guide the elimination of frequencies by finding the optimal threshold (as described below), we first remove additional pixels from the image and use them as the validation set.
+The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section ??).
+\item Since our reduction in the Fourier space works best with framed patches, we frame the complete image by mirroring its margin. (The size of the frame is another optimizable global parameter.)
+\item ...
+\end{enumerate}
+
+---------------------------\\
 

### 2011-06-20 project/Paper/paper.tex

--- 2011-06-20 project/Paper/paper.tex
@@
 
+\paragraph{Note}
+All plots in this paper were created with MATLAB$\textsuperscript{\textregistered}$ and can be reproduced using the script 'CreatePaperPosts.m' in our submission.\\
+
 \begin{figure}[tbp]
@@
 
-((( TODO: Plot missingPixel\% vs. accuracy (vs. speed?) )))
+((( TODO: Insert plot MissingPixels vs. Accuracy )))
 
@@
 
-\clearpage
-\section{Appendix}
-\begin{figure}[htbp]
+\begin{figure*}[b]
   \centering
-  \includegraphics[width=\textwidth]{images/flowchart}
+  \includegraphics[width= 400px]{images/flowchart}
   \caption{The algorithm as a flow chart}
   \label{fig:flowchart}
-\end{figure}
+\end{figure*}
 \end{document}

### 2011-06-20 project/CreatePaperPlots.m
Fixed bug in CreatePaperPlots

@@ -25,7 +25,7 @@ load('params.mat');
   cd('baseline1/')
   parfor k=0:no_steps
     %Baseline 1
-    error_base1(k+1) = feval('EvaluateInpaintingParametrized', stepsize*k/100);
+    error_base1(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
   end
   cd(main_path)
 

### 2011-06-20 project/CreatePaperPlots.m
Added baseline 2

@@ -7,6 +7,7 @@ load('params.mat');
   no_steps = floor(100/stepsize);
   error_algo = zeros(no_steps, 1);
   error_base1 = zeros(no_steps, 1);
+  error_base2 = zeros(no_steps, 1);
   
   main_path = pwd;
   
@@ -28,15 +29,24 @@ load('params.mat');
     error_base1(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
   end
   cd(main_path)
+  
+  fprintf('1C) Baseline 1\n')
+  cd('baseline2/')
+  parfor k=0:no_steps
+    %Baseline 1
+    error_base2(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
+  end
+  cd(main_path)
 
+  fprintf('1) Plotting\n')
   
   handle = figure;
   range = 0:stepsize:100;
   hold on;
   xlabel('Missing pixels (%)');
   ylabel('Average squared error per missing pixel');
-  plot(range, error_algo, 'r', range, error_base1,'b');
-  legend('New algorithm', 'Baseline 1');
+  plot(range, error_algo, 'r', range, error_base1,'b', range, error_base2, 'g');
+  legend('New algorithm', 'Baseline 1 (linear)', 'Baseline 2 (Matching Pursuit)');
   saveas(handle, 'plots/missingpixelsVsError.png');
   hold off;
   

### 2011-06-20 project/baseline2/EvaluateInpaintingParameterized.m

@@ -1,0 +1,57 @@
+% Measure approximation error for several images.
+
+function [avgQErr] = EvaluateInpaintingParameterized(missing_pixels_fract)  
+
+  file_list = dir('..'); 
+  rep=3;
+  
+  dir_length = length(dir('..'));
+
+  Errors_final = [];
+  LTimes_final = [];
+
+  parfor r = 1 : rep
+     Errors = []; % mean squared errors for each image
+     LTimes = [];
+     for i = 3 : dir_length  % running through the folder
+
+        file_name = file_list(i).name; % get current filename
+
+        % Only keep the images in the loop
+        if (length(file_name) < 5)
+          continue;
+        elseif ( max(file_name(end-4:end) ~= '2.png'))
+          continue;
+        end
+
+        % Read image, convert to double precision and map to [0,1] interval
+        I = imread(strcat('../',file_name));
+        I = double(I) / 255;
+
+        % Generate mask at random, with 60% missing pixels
+        mask = ones(size(I));
+        numpixels = size(I,1)*size(I,2);
+        holes = randperm(numpixels);
+        holes = holes(1:floor(missing_pixels_fract*numpixels));
+        mask(holes) = 0;
+
+
+        I_mask = I;
+        I_mask(~mask) = 0;
+
+        % Call the main inPainting function
+        tic;
+        I_rec = inPainting(I_mask, mask);
+        LTimes = [LTimes toc];
+
+        % Measure approximation error
+        Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
+     end
+     Errors_final = [Errors_final Errors];
+     LTimes_final = [LTimes_final LTimes];
+  end
+  
+  avgQErr = mean(Errors_final);
+
+
+end

### 2011-06-20 project/baseline2/inPainting.m

@@ -1,0 +1,35 @@
+function I_rec = inPainting(I, mask)
+
+% Perform the actual inpainting of the image 
+
+% INPUT
+% I: (n x n) masked image
+% mask: (n x n) the mask hidding image information
+%
+% OUTPUT
+% I_rec = Reconstructed image 
+
+global U;
+
+% Parameters
+rc_min = 0.01; % rc_min: minimal residual correlation before stopping
+neib = 16; % neib: The patch sizes used in the decomposition of the image
+sigma = 0.01; % sigma: residual error stopping criterion, normalized by signal norm
+
+% Get patches of size neib x neib from the image and the mask and
+% convert each patch to 1D signal
+X = my_im2col(I, neib);  
+M = my_im2col(mask, neib);  
+    
+% Construct your dictionary
+% If you load your own dictionary U calculated offline you don't have to 
+% add anything here
+U = buildDictionary(neib*neib);
+    
+% Do the sparse coding with modified Matching Pursuit
+Z = sparseCoding(U, X, M, sigma, rc_min);
+
+% You need to do the image reconstruction using the known image information
+% and for the missing pixels use the reconstruction from the sparse coding.
+% The mask will help you to distinguish between these two parts.
+I_rec = my_col2im(Z, neib, size(I, 1));

### 2011-06-20 project/baseline2/sparseCoding.m

@@ -1,0 +1,69 @@
+function Z = sparseCoding(U, X, M_orig, sigma, rc_min)
+%
+% Perform sparse coding using a modified matching pursuit tailored to the
+% inpainting problem with residual stopping criterion.
+%
+% INPUT
+% U: (d x l) unit norm atoms
+% X: (d x n) observations
+% M: (d x n) mask denoting which observations are unknown
+% sigma: residual error stopping criterion, normalized by signal norm
+% rc_min: minimal residual correlation before stopping
+%
+% OUTPUT
+% Z: MP coding
+%
+
+l = size(U,2);
+n = size(X,2);
+
+Z = zeros(size(X));
+
+% Loop over all observations in the columns of X
+for nn = 1:n
+    %[nn, n]
+
+    % Initialize the residual with the observation x
+    % For the modification with masking make sure that you only take into
+    % account the known observations defined by the mask M
+    R = X(:, nn);
+    M = M_orig(:, nn);
+
+    % Remove the missing pixels and corresponding ??? of Z
+    masked_R = R;
+    % d_e_l_e_t_e those
+    masked_R(M==0) = [];
+
+    %Adapt U to fit size of masked_R
+    masked_U = U;
+    masked_U(M==0,:) = [];
+
+    % Initialize z to zero
+    z = zeros(l, 1);
+
+    threshold = sigma*norm(X(:, nn));
+    rc_max = +inf;
+
+    while (norm(masked_R) > threshold && (abs(rc_max) > rc_min))
+      % Select atom with maximum absolute correlation to the residual
+      % Update the maximum absolute correlation
+      rc_max = 0;
+      best_atom = 0;
+
+      results = masked_R'*masked_U;
+      [tilde, best_atom] = max(abs(results));
+      
+      rc_max = results(best_atom);
+
+      % Update coefficient vector z and residual z
+      masked_R = masked_R - rc_max*masked_U(:, best_atom);
+      z(best_atom) = z(best_atom) + rc_max;
+
+      % For the inpainting modification make sure that you only consider
+      % the known observations defined by the mask M
+      %masked_R = M*R;
+    end
+    
+    % Add the calculated coefficient vector z to the overall matrix Z
+    Z(:,nn) = M.*X(:, nn) + (1-M).*(U*z);
+end

### 2011-06-20 project/gaussInterpolate.m
+ Added some asserts

@@ -11,4 +11,10 @@ I_out = (I_filtered .* (1-mask)) + (I_masked .* (mask));
 
 %Make sure we eliminate NaNs that may occur,
 %when gauss is too small and we have a large hole
-I_out(isnan(I_out)) = mean(mean(I(logical(mask))));
+%I_out(isnan(I_out)) = mean(mean(I(logical(mask))));
+
+s = sum(sum(isnan(I_out)));
+if (s > 0)
+  error(sprintf('Have %g nans!', s));
+end
+

### 2011-06-20 project/inPaintingParameterized.m

@@ -10,6 +10,16 @@ function I_rec = inPaintingParameterized(I, mask, parameters)
 % OUTPUT
 % I_rec: Reconstructed image 
 
+s = size(I, 1);
+if s ~= 512
+  error(sprintf('Wrong image size "%g"', s));
+end
+
+percent_unknown = sum(sum(~mask)) / (s*s);
+if percent_unknown > 0.65
+  error(sprintf('Too many unknowns "%g"', percent_unknown));
+end
+
   % Make sure that parameters do have possible values
   % Round floating point values that come from optimizing
   parameters.gauss_size = round(parameters.gauss_size);

### 2011-06-20 project/CreatePaperPlots.m
Save as fig instead of png

@@ -1,7 +1,7 @@
 % Create Error vs. MissingPixels Plot
 % Requires parameters saved in a parameters.mat file
 
-load('params.mat');
+  load('params.mat');
 
   stepsize = 2;
   no_steps = floor(100/stepsize);
@@ -47,7 +47,7 @@ load('params.mat');
   ylabel('Average squared error per missing pixel');
   plot(range, error_algo, 'r', range, error_base1,'b', range, error_base2, 'g');
   legend('New algorithm', 'Baseline 1 (linear)', 'Baseline 2 (Matching Pursuit)');
-  saveas(handle, 'plots/missingpixelsVsError.png');
+  saveas(handle, 'plots/missingpixelsVsError.fig');
   hold off;
   
   

### 2011-06-20 project/Paper/paper.tex
Updated .bib file. Removed unused references

@@
   The abstract should really be written last, along with the title of
-the paper. The four points that should be covered~\cite{jones08}:
+the paper. The four points that should be covered:
 \begin{enumerate}
@@
 that an interested reader can reproduce your
-work~\cite{anderson04,wavelab}.
+work.
 The methods
@@
 It is usually helpful to
-structure the methods section by~\cite{kallet04methods}:
+structure the methods section by:
 \begin{enumerate}
@@
  \begin{enumerate}
- \item A simple inpainting algorithm using (((TODO )))
- \item A matching pursuit algorithm with a dictionary consisting of (((TODO)))
+	 \item A simple linear inpainting algorithm using the nearest neighbor pixels.
+	 \item A matching pursuit algorithm \cite{matchingpursuit93} with an overcomplete dictionary consisting of Discrete Cosine transforms.
  \end{enumerate}
@@
 \paragraph{Note}
-All plots in this paper were created with MATLAB$\textsuperscript{\textregistered}$ and can be reproduced using the script 'CreatePaperPosts.m' in our submission.\\
+All plots in this paper were created with MATLAB$\textsuperscript{\textregistered}$ and can be reproduced using the script 'CreatePaperPlots.m' in our submission.\\
 

### 2011-06-20 project/Paper/paper.tex
Removed assertion again

@@
 \section{Results}
-We evaluated two versions of our algorithm: An iterative variant and a non-iterative variant.
-Our iterative variant repeats 
+
  ((( TODO )))
@@
 	 \item A simple linear inpainting algorithm using the nearest neighbor pixels.
-	 \item A matching pursuit algorithm \cite{matchingpursuit93} with an overcomplete dictionary consisting of Discrete Cosine transforms.
+	 \item A matching pursuit algorithm \cite{matchingpursuit93} with an overcomplete dictionary built with Discrete Cosine transforms.
  \end{enumerate}
+
+  Our algorithm is resistant to high percentages of missing pixels, the average error per missing pixel rises only slowly until ((( TODO level))).
 
@@
 report the mean (average value) along with a measure of variablility
-(standard deviation(s) or standard error of the mean).
+(standard deviation(s) or standard error of the mean). ((( TODO !!! )))
 

### 2011-06-20 project/gaussInterpolate.m

@@ -11,10 +11,5 @@ I_out = (I_filtered .* (1-mask)) + (I_masked .* (mask));
 
 %Make sure we eliminate NaNs that may occur,
 %when gauss is too small and we have a large hole
-%I_out(isnan(I_out)) = mean(mean(I(logical(mask))));
-
-s = sum(sum(isnan(I_out)));
-if (s > 0)
-  error(sprintf('Have %g nans!', s));
-end
+I_out(isnan(I_out)) = mean(mean(I(logical(mask))));
 

### 2011-06-20 project/Paper/paper.tex
Paper: New approach completed.

--- 2011-06-20 project/Paper/paper.tex
@@
 \begin{abstract}
-  This paper suggests a novel approach to restore noisy images.
-  ((( TODO )))
 The abstract should really be written last, along with the title of
@@
 \item Since our reduction in the Fourier space works best with framed patches, we frame the complete image by mirroring its margin. (The size of the frame is another optimizable global parameter.)
-\item ...
+\item For obtaining an initial reconstruction of the image, we compute a weighted sum over all the neighboring pixels that are known.
+By setting all unknown pixels in the image to zero and encoding known values with 1s in the mask, elementwise division of the filtered image by the similarly filtered mask results in a normalized sum weighted by the filtering kernel.
+(This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.)
 \end{enumerate}
 
----------------------------\\
+After these preparing steps, we start to iteratively improve the reconstructed image with the following steps:
+\begin{enumerate}
+\item Using the abovementioned validation set, we evaluate several thresholds for every patch and keep for each the best one: After calculating a fast Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
+Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to perform as few threshold evaluations as possible.
+For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and decreasing the offset.
+For this optimization as well as for the whole iteration, we have both a number- and a convergence-based termination criterion.
+\item Having stored the reconstruction of every patch with its best threshold, all that remains to do is copying the newly estimated missing pixels (including those of the validation set) back into the reconstructed image.
+\end{enumerate}
+Finally, the frame added in the beginning is removed again and the pixels of the validation set are overwritten by the original ones.
+A variant of the presented method is to do it non-iteratively and revert to the original data after determining the thresholds, such that the Gaussian interpolation and the thresholding in the Fourier space is performed without the validation set.
 
-We use an iterative scheme to determine the right parameters used in our algorithm.
-For this, we use cross-validation: We randomly remove a percentage of pixels from the image as a validation set.
-(((STILL VALID?))) We then use this set to test the accuracy of our parameters and determine the best set of parameters for this image.
-
-We then proceed to the main steps of the algorithm:
-\begin{enumerate}
-\label{algo}
-\item \label{algo:step1}
-We use a Gaussian to interpolate the missing pixels indicated by the mask, using a convolution operation.
-Both $\sigma$ and the size of the Gaussian matrix are determined by the cross-validation.
-(((TODO?!)))
-This step is required by the Fourier transform which cannot ignore samples. (((Humm)))
-
-\item \label{algo:step2}
-We transform the image to the Fourier domain.
-In this domain we set the magnitudes of all frequencies to zero which are below the $n$-th percentile threshold.
-Again, our ideal cutoff $n$ is determined by cross-validation using a loop.
-
-\item \label{algo:step3}
-In this last step, we invert the Fourier transform and combine the reconstructed image with the undamaged part of the image, yielding our result.
-
-\end{enumerate}
-
-In our implementation, we use several tweaks to improve performance or quality:
-\begin{itemize}
-\item We perform the Fourier transform and thresholding on image patches to allow an individual threshold per patch.
-\item To avoid discontinuity (Image \ref{fig:boundaryArtifacts}) at the boundary of the patches, we use a frame around the actual patch to capture more information.
-(Image \ref{fig:framing}) This frame is not filled in by the reconstructed patch from the Fourier step.
-The frames of adjacent patches are overlapping if the frame size is set to a value larger than zero.
-At the image boundaries, we replicate the outermost pixels to provide a sufficient frame.\end{itemize}
-((( TODO )))
-Describe your idea and how it was implemented to solve
-  the problem. Survey the related work, giving credit where credit is
-  due.
-
-  The style in this section should read as if you were verbally
-describing the conduct of the experiment. 
-The methods section is not a step-by-step, directive,
-protocol as you might see in your lab manual, but detailed enough such
-that an interested reader can reproduce your
-work~\cite{anderson04,wavelab}.
-The methods
-section should describe what was
-done to answer the research question, describe how it was done,
-justify the experimental design, and
-explain how the results were analyzed.
-
-The methods section of a research paper provides the information by
-which a study's validity is judged.
-Therefore, it requires a clear and precise description of how an
-experiment was done, and the rationale
-for why specific experimental procedures were chosen.
-It is usually helpful to
-structure the methods section by~\cite{kallet04methods}:
-\begin{enumerate}
-\item Describing the algorithms used in the study, briefly including
-  details such as hyperparameter values (e.g. thresholds), and
-  preprocessing steps (e.g. normalizing the data to have mean value of
-  zero).
-\item Explaining how the materials were prepared, for example the
-  images used and their resolution.
-\item Describing the research protocol, for example which examples
-  were used for estimating the parameters (training) and which were
-  used for computing performance.
-\item Explaining how measurements were made and what
-  calculations were performed. Do not reproduce the full source code in
-  the paper, but explain the key steps.
-\end{enumerate}
-
-\subsection{Online threshold estimation}
-In the algorithm as described above, we have many different parameters that are required to be tuned
-
-\subsection{Offline parameter estimation}
+\subsection{Global parameter estimation}
 As shown in table \ref{parameters}, there are a total of 10 parameters in our algorithm that can be set.
@@
 \paragraph{Note}
-All plots in this paper were created with MATLAB$\textsuperscript{\textregistered}$ and can be reproduced using the script 'CreatePaperPosts.m' in our submission.\\
+All plots in this paper were created with MATLAB and can be reproduced using the script 'CreatePaperPosts.m' in our submission.\\
 

### 2011-06-20 project/Paper/paper.tex

@@
  \begin{enumerate}
- \item A simple inpainting algorithm using (((TODO )))
- \item A matching pursuit algorithm with a dictionary consisting of (((TODO)))
+	 \item A simple linear inpainting algorithm using the nearest neighbor pixels.
+	 \item A matching pursuit algorithm \cite{matchingpursuit93} with an overcomplete dictionary built with Discrete Cosine transforms.
  \end{enumerate}
+
+  Our algorithm is resistant to high percentages of missing pixels, the average error per missing pixel rises only slowly until ((( TODO level))).
 

### 2011-06-20 project/Paper/paper.tex
Minor, for pulling
Added parameter table

@@
   \exp(E)+\exp(T/60 - 1) - 1 & \text{else}
-\end{cases}$ & Optimizes for error only when time is low enough. Function still continous, penalizes high times. & None so far, this is the function used by us\\
+\end{cases}$ & Optimizes for error only when time is low enough. Function still continuous, penalizes high times. & None so far, this is the function used by us\\
     \hline
@@
 
- We tested our algorithm and the two baselines on a set of natural images and further on a set of artificially created ones. (((TODO More than 1)))
+We tested our algorithm and the two baselines on a set of natural images and further on a set of artificially created ones. 
+For evaluation, we tested these algorithms with different levels of missing pixels from 0 to 100\% in 5\% steps.
 
@@
 \subsection{Figures and Tables}
+
+\begin{table*}[p]
+\begin{center}
+\begin{tabular}{|c||c|c|c|}
+\hline
+Parameter & Meaning & Initial value & Optimized value\\
+\hline\hline 
+gauss\_size & Size of the gaussian & ? & ? \\
+gauss\_sigma & Sigma of the gaussian & ? & ? \\
+patch\_size & Patch size & ? & ? \\
+patch\_frame\_size & Frame size & ? & ? \\
+td\_abortbelow\_stdev & Minimal standard deviation for threshold determination& ? & ? \\
+td\_abortbelow\_stepsize & Minimal step size for threshold determination & ? & ? \\
+td\_middle &  Initial threshold & ? & ? \\
+validation &  Size of the the validation set & ? & ? \\
+max\_iterations & Maximum number of iterations & ? & ? \\
+abortbelow\_change & Minimal change to perform another iteration & ? & ? \\
+
+\hline
+\end{tabular}
+\end{center}
+\caption{Parameters, their initial and optimized values}
+\label{parameter_table}
+\end{table*}%
 

### 2011-06-20 project/dimensionReduction.m
Removed duplicate calls Added comments

@@ -1,4 +1,5 @@
 function I_reduced = dimensionReduction(I, T, parameters)
+% DIMENSIONREDUCTION Reduces dimension in Fourier space with threshold T
 
 pfs = parameters.patch_frame_size;
 ps = parameters.patch_size;
@@ -7,21 +8,24 @@ image_size = size(I, 1);
 num_patches = (image_size-2*pfs) / ps;
 I_reduced = I;
 
+% Iterate over all patches
 for i = 1 : num_patches
   for j = 1 : num_patches
+    % Extract patch
     P_framed = I(ps * (i - 1) + 1 : ...
                  ps * i + 2 * pfs, ...
 		 ps * (j - 1) + 1 : ...
                  ps * j + 2 * pfs);
 
+    % To fourier domain
     P_fft = fft2(P_framed);
-
+  
+    % Thresholding, then back to spatial domain
     P_fft(abs(P_fft) < T(i, j)) = 0;
     P_ifft = ifft2(P_fft);
 
     % Make sure we do not fall outside the value range of an image
     P_rec = removeFrame(P_ifft, parameters);
-    P_rec = boundImageValues(P_rec);
 
     range_i = pfs + ps*(i-1) + 1 : pfs + ps * i;
     range_j = pfs + ps*(j-1) + 1 : pfs + ps * j;

### 2011-06-20 project/Paper/paper.tex
+ Added nice flowchart

--- 2011-06-20 project/Paper/paper.tex
@@
 \usepackage[utf8]{inputenc}
+
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
 
@@
 \subsection{New approach}
+
+% Define block styles
+\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
+    text width=4.5em, text badly centered, inner sep=0pt]
+
+\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
+    text width=7em, text centered, rounded corners, minimum height=4em]
+
+\tikzstyle{line} = [draw, -latex']
+
+\tikzstyle{cloud} = [draw, ellipse,fill=red!20,
+    minimum height=2em]
+
+\begin{tikzpicture}[node distance = 2cm, auto]
+  % Place nodes
+  \node [cloud] (node1) {Start};
+  \node [block, right of=node1, node distance = 3.1cm] (node2) {Remove validation set};
+  \node [block, below of=node2] (node3) {Frame image};
+  \node [block, below of=node3] (node4) {Gaussian interpolation};
+  \node [decision, below of=node4] (node5) {Iterative?};
+  \node [block, right of=node5, node distance = 3.1cm, fill=green!20] (node6) {Find best threshold};
+  \node [block, below of=node6, fill=green!20] (node7) {Reset known values except validation set};
+
+  \node [block, left of=node5, node distance = 3.1cm] (node8) {Find best threshold};
+  \node [block, below of=node8] (node9) {Gaussian interpolation on original data};
+  \node [block, below of=node9] (node10) {Apply best threshold};
+
+  \node [block, right of=node10, node distance = 3.1cm] (node11) {Unframe image};
+  \node [block, below of=node11] (node12) {Reset known values};
+  \node [cloud, right of=node12, node distance = 3.1cm] (node13) {End};
+
+  % Draw edges
+  \path [line] (node1) -- (node2);
+  \path [line] (node2) -- (node3);
+  \path [line] (node3) -- (node4);
+  \path [line] (node4) -- (node5);
+  \path [line] (node5) -- node[above]{yes} (node6);
+  \path [line] (node5) -- node[above]{no} (node8);
+
+  \draw [->] (node6) to [bend left=20] (node7);
+  \draw [->] (node7) to [bend left=20] (node6);
+
+  \path [line] (node8) -- (node9);
+  \path [line] (node9) -- (node10);
+  \path [line] (node10) -- (node11);
+  \path [line] (node11) -- (node12);
+  \path [line] (node12) -- (node13);
+
+  \path [line] (node7) |- (node11);
+\end{tikzpicture}
+
 As outlined in the introduction, we combined several ideas for obtaining an inpainting algorithm that provides both good results and fast performance.

### 2011-06-20 project/Paper/paper.tex
Changes.

--- 2011-06-20 project/Paper/paper.tex
@@
 
-\subsection{Global parameter estimation}
+\subsection{Global parameter optimization}
 As shown in table \ref{parameters}, there are a total of 10 parameters in our algorithm that can be set.
-For many of them, it is quite clear what values should be used. For example, the size of the gauss kernel
-should be a bit larger than the biggest assumed hole in the image. For others, like the maximum number of
-iterations, it is not so clear. And yet others have a clear speed / accuracy tradeoff, like the stopping
-conditions of the online threshold estimation.
-
-Because we thought it not feasable to tune all the parameters by hand, we wrote an algorithm to
-do the offline parameter estimation. The idea is to do this once for a few representative images, and get
-parameters that can model as closely as possible many of the images we are interested in.
-Our proposed method combines a gradient descent method with ideas from cross-validation, and tunes our
-10-dimensional parameter-space, maximising time and accuracy.
+For many of them, it is quite clear what values should be used.
+For example, the size of the gauss kernel should be a bit larger than the biggest assumed hole in the image.
+For others, like the maximum number of iterations, it is not so clear.
+And yet others have a clear speed / accuracy tradeoff, like the stopping conditions of the threshold optimization.\\
+Because we thought it not feasable to tune all the parameters by hand, we wrote an algorithm to do the global parameter optimization.
+The idea is to do this once for a few representative images, and get parameters that can model as closely as possible many of the images we are interested in.
+Our proposed method combines a gradient descent method with ideas from cross-validation, and tunes our 10-dimensional parameter-space, minimizing time and maximizing accuracy.
 
 \subsubsection{Error function}
-When doing a minimization of any kind, an error function is needed. Of course we already had the error function
-of the evaluation script, that gives a mean over the squared differences of the reconstructed to the original
-image. However, using only an error measure, the time would potentially grow to infinity to get the best accuracy possible during tuning.
-
-Therefore we tried different error function including both, the image error $E$, and the time $T$ spent (see table \ref{error_function_table}). The image error here is the average
-squared error of a reconstructed pixel to the original pixel, scaled with a factor (usually about 10000), while the time is measured in seconds.
+When doing a minimization of any kind, an error function is needed.
+Of course we already had the error function of the evaluation script, that gives a mean over the squared differences of the reconstructed to the original image.
+However, using only an error measure, the time would potentially grow to infinity to get the best accuracy possible during tuning.\\
+Therefore we tried different error function including both, the image error $E$, and the time $T$ spent (see table \ref{error_function_table}).
+The image error here is the average squared error of a reconstructed pixel to the original pixel, scaled with a factor (usually about 10000), while the time is measured in seconds.
 
@@
     $\begin{cases}
-  \exp(E)  & \text{if }T<60\\
-  \exp(E)+\exp(T/60 - 1) - 1 & \text{else}
+  E  & \text{if }T<60\\
+  E+\exp(T/60 - 1) - 1 & \text{otherwise}
 \end{cases}$ & Optimizes for error only when time is low enough. Function still continous, penalizes high times. & None so far, this is the function used by us\\

### 2011-06-20 project/Paper/paper.tex

--- 2011-06-20 project/Paper/paper.tex
@@
 
+\begin{table*}[p]
+\begin{center}
+\begin{tabular}{|c||c|c|c|}
+\hline
+Parameter & Meaning & Initial value & Optimized value\\
+\hline\hline 
+gauss\_size & Size of the gaussian & ? & ? \\
+gauss\_sigma & Sigma of the gaussian & ? & ? \\
+patch\_size & Patch size & ? & ? \\
+patch\_frame\_size & Frame size & ? & ? \\
+td\_abortbelow\_stdev & Minimal standard deviation for threshold determination& ? & ? \\
+td\_abortbelow\_stepsize & Minimal step size for threshold determination & ? & ? \\
+td\_middle &  Initial threshold & ? & ? \\
+validation &  Size of the the validation set & ? & ? \\
+max\_iterations & Maximum number of iterations & ? & ? \\
+abortbelow\_change & Minimal change to perform another iteration & ? & ? \\
+
+\hline
+\end{tabular}
+\end{center}
+\caption{Parameters, their initial and optimized values}
+\label{parameter_table}
+\end{table*}%
+
 \paragraph{Note}

### 2011-06-20 project/inPaintingParameterized.m
Documentation / Comments

@@ -36,24 +36,32 @@ val_mask(val_indices) = 1;
 % ----------------------------------
 
 I_training = I;
+% Delete pixels that we do not know
 I_training(val_indices) = 0;
 
 mask_training = mask;
 mask_training(val_indices) = 0;
 
+% Do Gauss interpolation
 I_training = gaussInterpolate(I_training, mask_training, parameters);
 
+% Create a frame around the image
 I_training_framed = createFrame(I_training, parameters);
+% Dito: Mask
 mask_training_framed = createFrame(mask_training, parameters);
 
 if (parameters.iterative)
+  %Iterative version
   previous_error = +Inf;
   for i = 1:parameters.max_iterations
     % TODO Abort if gauss is fine anyway?
+    % Find the thresholds
     [T, I_trained] = determineThresholds(I_training_framed, val_mask, I, parameters);
     
+    % Composite the image with known stuff
     I_training_framed = I_training_framed.*mask_training_framed + ...
                         (1-mask_training_framed).*I_trained;
+    % Calculate error
     diff = (removeFrame(I_training_framed, parameters) - I).*val_mask;
     err = sum(sum(diff.*diff));
     if(1/previous_error*err > 1-parameters.abortbelow_change)
@@ -62,8 +70,10 @@ if (parameters.iterative)
     end
     previous_error=err;
   end
+  % Write back the image
   I_final = I_training_framed;
 else
+  % Non-iterative variant
   I_framed = createFrame(I, parameters);
   mask_framed = createFrame(mask, parameters);
   I_framed = gaussInterpolate(I_framed, mask_framed, parameters);
@@ -71,5 +81,5 @@ else
   [T, I_trained] = determineThresholds(I_training_framed, val_mask, I, parameters);
   I_final = dimensionReduction(I_framed, T, parameters);
 end
-
+% Composite the image back together and return it
 I_rec = I.*mask + (1-mask).*removeFrame(I_final, parameters);

### 2011-06-20 project/CreatePaperPlots.m
Fix output...

@@ -1,3 +1,5 @@
+% Creates plots and figures used in the paper. (CAVEAT: Takes long!)
+
 % Create Error vs. MissingPixels Plot
 % Requires parameters saved in a parameters.mat file
 
@@ -5,6 +7,7 @@
 
   stepsize = 2;
   no_steps = floor(100/stepsize);
+  % Init arrays
   error_algo = zeros(no_steps, 1);
   error_base1 = zeros(no_steps, 1);
   error_base2 = zeros(no_steps, 1);
@@ -13,9 +16,8 @@
   
   fprintf('1) Calculating error vs. missing pixels\n')
   fprintf('1A) New algorithm\n')
-  %Use parallel computation for this
+  %Use parallel computation for this, if available
   parfor k=0:no_steps
-
     % Our algorithm
     [unused, e] = EvaluateInpaintingParameterized(parameters,stepsize*k/100);
     error_algo(k+1) = e/(k*stepsize/100);% Normalized
@@ -30,7 +32,7 @@
   end
   cd(main_path)
   
-  fprintf('1C) Baseline 1\n')
+  fprintf('1C) Baseline 2\n')
   cd('baseline2/')
   parfor k=0:no_steps
     %Baseline 1

### 2011-06-20 project/Paper/paper.tex
Paper: Legacy figures removed, caption, minor improvements.

--- 2011-06-20 project/Paper/paper.tex
@@
 \author{
+  Group: NoobzRockz\\
   Patrick Bänziger, Kaspar Etter, Jan Rüegg\\
@@
 
+\begin{figure}
 \begin{tikzpicture}[node distance = 2cm, auto]
@@
 \end{tikzpicture}
+\caption{Our newly proposed algorithm as a flowchart.}
+\end{figure}
 
@@
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\columnwidth]{images/imageFraming.png}
-  \caption{Only the inner patch (marked red) is reconstructed in one step while the frame only provides additional signal. (Undistorted image shown) }
-  \label{fig:framing}
-\end{figure}
-
 ((( TODO: Insert plot MissingPixels vs. Accuracy )))
@@
 
-\begin{figure*}[b]
-  \centering
-  \includegraphics[width= 400px]{images/flowchart}
-  \caption{The algorithm as a flow chart}
-  \label{fig:flowchart}
-\end{figure*}
 \end{document}

### 2011-06-21 project/CreatePaperPlots.m
c Cleanup, comments

@@ -1,89 +1,89 @@
-% Creates plots and figures used in the paper. (CAVEAT: Takes long!)
+% CreatePaperPlots
+% ================
+% Script to create plots and figures used in the paper. (CAVEAT: Takes long!)
 
 % Create Error vs. MissingPixels Plot
 % Requires parameters saved in a parameters.mat file
 
-  load('params.mat');
+load('params.mat');
 
-  stepsize = 2;
-  no_steps = floor(100/stepsize);
-  % Init arrays
-  error_algo = zeros(no_steps, 1);
-  error_base1 = zeros(no_steps, 1);
-  error_base2 = zeros(no_steps, 1);
-  error_base3 = zeros(no_steps, 1);
-  
-  main_path = pwd;
-  
-  fprintf('1) Calculating error vs. missing pixels\n')
-  fprintf('1A) New algorithm\n')
-  
-  if(~exist('plots/error_A.mat', 'file'))
-    %Use parallel computation for this, if available
-    parfor k=0:no_steps
-      % Our algorithm
-      [unused, e] = EvaluateInpaintingParameterized(parameters,stepsize*k/100);
-      error_algo(k+1) = e/(k*stepsize/100);% Normalized
-    end
-    save('plots/error_A.mat', 'error_algo');
-  else
-    fprintf('Found saved data file. Loading...')
-    load('plots/error_A.mat');
-  end
-  
-  fprintf('1B) Baseline 1\n')
-  if(~exist('plots/error_B1.mat', 'file'))    
-    cd('baseline1/')
-    parfor k=0:no_steps
-      %Baseline 1
-      error_base1(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
-    end
-    cd(main_path)
-    save('plots/error_B1.mat', 'error_base1');
-  else
-    fprintf('Found saved data file. Loading...')
-    load('plots/error_B1.mat');
-  end
-  
-  fprintf('1C) Baseline 2\n')
-  if(~exist('plots/error_B2.mat', 'file'))    
-    cd('baseline2/')
-    parfor k=0:no_steps
-      %Baseline 2
-      error_base2(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
-    end
-    cd(main_path)
-    save('plots/error_B2.mat', 'error_base2');
-  else
-    fprintf('Found saved data file. Loading...')
[…]

### 2011-06-21 project/EvaluateInpaintingParameterized.m

@@ -1,10 +1,9 @@
+% EvaluateInpaintingParameterized
+% ===============================
 % Measure approximation error for several images.
-
+% Adjusted function of the framework thatrepeats the evaluation step to compensate
+% for randomness introduced by the random mask (fixed mask in the initially provided implementation)
 function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_pixels_fract)  
-  %EvaluateInpaintingParameterized Adjusted function of the framework that
-  %repeats the evaluation step to compensate for randomness introduced by
-  %the random mask (fixed mask in the initially provided implementation)
-  
   file_list = dir(); 
   rep=3;
   
@@ -13,48 +12,49 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
   Errors_final = [];
   Times_final = [];
 
-  % Repreat rep times to compensate for noise due to the random mask
+  % Repeat rep times to compensate for noise due to the random mask
   % Parfor will do so in parallel, if possible
   parfor r = 1 : rep
-     %Reduction variables (parfor)
-     Errors = []; % mean squared errors for each image
-     Times = [];
+    % Reduction variables (parfor)
+    Errors = []; % mean squared errors for each image
+    Times = [];
      
-     %Search for pictures
+    % Search for pictures
      for i = 3 : dir_length  % running through the folder
+       file_name = file_list(i).name; % get current filename
 
-        file_name = file_list(i).name; % get current filename
+       % Only keep the images in the loop
+       if (length(file_name) < 5)
+         continue;
+       elseif ( max(file_name(end-4:end) ~= '2.png'))
+         continue;
+       end
 
-        % Only keep the images in the loop
-        if (length(file_name) < 5)
-          continue;
-        elseif ( max(file_name(end-4:end) ~= '2.png'))
-          continue;
-        end
+       % Read image, convert to double precision and map to [0,1] interval
+       I = imread(file_name);
+       I = double(I) / 255;
 
-        % Read image, convert to double precision and map to [0,1] interval
-        I = imread(file_name);
-        I = double(I) / 255;
-
-        % Generate mask at random, with adjustable number missing pixels
-        mask = ones(size(I));
-        numpixels = size(I,1)*size(I,2);
-        holes = randperm(numpixels);
[…]

### 2011-06-21 project/createFrame.m

@@ -1,16 +1,19 @@
+% createFrame
+% ===========
+% Create a frame around the image.
+% Size of the frame depending on the parameters. The frame is a mirrored
+% version of the outermost rows/columns.
+
 function I_framed = createFrame(I, parameters)
-  %CREATEFRAME Create a frame around the image
-  %Size of the frame depending on the parameters. The frame is a mirrored
-  %version of the outermost rows/columns.
   image_size = size(I, 1);
 
+  % Create a bigger image, with zeros as borders
   I_framed = zeros(image_size + 2*parameters.patch_frame_size);
-
   I_framed(parameters.patch_frame_size + 1 : parameters.patch_frame_size + image_size, ...
            parameters.patch_frame_size + 1 : parameters.patch_frame_size + image_size) = I;
   
+  % Fill the borders with the mirrored pixels
   for i = 1 : parameters.patch_frame_size
-    % TODO optimize, use repmat
     I_framed(:, i) = I_framed(:, 2*parameters.patch_frame_size - i+1);
     I_framed(:, end - i + 1) = I_framed(:, end - 2*parameters.patch_frame_size + i);
     I_framed(i, :) = I_framed(2*parameters.patch_frame_size - i+1, :);

### 2011-06-21 project/determineThresholds.m

@@ -1,23 +1,28 @@
-function [T, I_trained] = determineThresholds(I_training_framed, val_mask, ...
-                                              val_I, parameters)
-  %DETERMINETHRESHOLDS Determines the best thresholds, reconstruction for each patch
-  
+% determineThresholds
+% ===================
+% Determines the best thresholds in the fourier domain for each patch.
+% The function takes a framed training image, a validation mask and a validation image (ground truth),
+% together with a set of paramateters. It then calculates the optimal thresholds in the fourier domain
+% for each of the patches
+
+function [T, I_trained] = determineThresholds(I_training_framed, val_mask, val_I, parameters)
   image_size = size(val_I, 1);
   num_patches = image_size / parameters.patch_size;
   pfs = parameters.patch_frame_size;
   ps = parameters.patch_size;
 
+  % Initialize training image, best thresholds
   T = zeros(num_patches);
   I_trained = I_training_framed;
 
-  %Iterate over all patches
+  % Iterate over all patches
   for i = 1 : num_patches
     for j = 1 : num_patches
       best_t = 1;
       best_error = +Inf;
       best_P = 0;
 
-      %Extract patches from image / mask / validation mask
+      % Extract patches from image / mask / validation mask
       P_framed = I_training_framed( ...
                             ps * (i - 1) + 1 : ...
                             ps * i + 2 * pfs, ...
@@ -34,22 +39,22 @@ function [T, I_trained] = determineThresholds(I_training_framed, val_mask, ...
                          ps * (j - 1) + 1 : ...
                          ps * j).*P_mask;
 
-      %Prepare iteration
+      % Prepare iteration
       dev = +Inf;
       middle = parameters.td_middle;
       stepsize = middle;
       errors = [Inf, Inf, Inf];
-      %Ps: Reconstructed patches.
+
+      % Ps: Reconstructed patches.
       Ps = {};
       % Calculate middle value once before loop.
       [errors(2), Ps{2}] = determineError(P_framed, middle, P_mask, P_validate, parameters);
       
-      %DEBUG / Visualization
+      % DEBUG / Visualization
       middles = [middle];
       
-      %Find best threshold iteratively
+      % Find best threshold iteratively
       while(stepsize > parameters.td_abortbelow_stepsize && dev > parameters.td_abortbelow_stdev)
-        
         % Right value
[…]

### 2011-06-21 project/dimensionReduction.m

@@ -1,36 +1,38 @@
+% dimensionReduction
+% ==================
+% Reduces dimension in Fourier space with threshold T
+% Takes an input image and a threshold matrix, together with the
+% desired parameters struct
 function I_reduced = dimensionReduction(I, T, parameters)
-% DIMENSIONREDUCTION Reduces dimension in Fourier space with threshold T
+  pfs = parameters.patch_frame_size;
+  ps = parameters.patch_size;
 
-pfs = parameters.patch_frame_size;
-ps = parameters.patch_size;
+  image_size = size(I, 1);
+  num_patches = (image_size-2*pfs) / ps;
+  I_reduced = I;
 
-image_size = size(I, 1);
-num_patches = (image_size-2*pfs) / ps;
-I_reduced = I;
+  % Iterate over all patches
+  for i = 1 : num_patches
+    for j = 1 : num_patches
+      % Extract patch
+      P_framed = I(ps * (i - 1) + 1 : ...
+                   ps * i + 2 * pfs, ...
+                   ps * (j - 1) + 1 : ...
+                   ps * j + 2 * pfs);
 
-% Iterate over all patches
-for i = 1 : num_patches
-  for j = 1 : num_patches
-    % Extract patch
-    P_framed = I(ps * (i - 1) + 1 : ...
-                 ps * i + 2 * pfs, ...
-		 ps * (j - 1) + 1 : ...
-                 ps * j + 2 * pfs);
+      % To fourier domain
+      P_fft = fft2(P_framed);
+    
+      % Thresholding, then back to spatial domain
+      P_fft(abs(P_fft) < T(i, j)) = 0;
+      P_ifft = ifft2(P_fft);
 
-    % To fourier domain
-    P_fft = fft2(P_framed);
-  
-    % Thresholding, then back to spatial domain
-    P_fft(abs(P_fft) < T(i, j)) = 0;
-    P_ifft = ifft2(P_fft);
+      % Make sure we do not fall outside the value range of an image
+      P_rec = removeFrame(P_ifft, parameters);
 
-    % Make sure we do not fall outside the value range of an image
-    P_rec = removeFrame(P_ifft, parameters);
-
-    range_i = pfs + ps*(i-1) + 1 : pfs + ps * i;
-    range_j = pfs + ps*(j-1) + 1 : pfs + ps * j;
-    I_reduced(range_i, range_j) = P_rec;
[…]

### 2011-06-21 project/inPaintingParameterized.m

@@ -1,15 +1,13 @@
-function I_rec = inPaintingParameterized(I, mask, parameters)
-
-% Perform the actual inpainting of the image.
-
+% inPaintingParametrized
+% ======================
 % INPUT
 % I: Masked image (n x n)
 % mask: Mask itself (n x n)
 % parameters: Parameters (as an object)
 %
 % OUTPUT
-% I_rec: Reconstructed image 
-
+% I_rec: Reconstructed image
+function I_rec = inPaintingParameterized(I, mask, parameters)
   % Make sure that parameters do have possible values
   % Round floating point values that come from optimizing
   parameters.gauss_size = round(parameters.gauss_size);
@@ -20,66 +18,69 @@ function I_rec = inPaintingParameterized(I, mask, parameters)
   parameters.td_abortbelow_stepsize = 10^-(parameters.td_abortbelow_stepsize);
   parameters.abortbelow_change = 0.15; 
 
+  % Split into validation and training set
+  good_indices = find(mask);
+  num_good = length(good_indices);
+  perm = randperm(num_good);
+  val_indices = good_indices(perm(1:round(parameters.validation * num_good)));
+  val_mask = zeros(size(mask));
+  val_mask(val_indices) = 1;
 
-% Split into validation and training set
-good_indices = find(mask);
-num_good = length(good_indices);
-perm = randperm(num_good);
-val_indices = good_indices(perm(1:round(parameters.validation * num_good)));
-val_mask = zeros(size(mask));
-val_mask(val_indices) = 1;
+  % ----------------------------------
+  % | known                 | masked | = Input
+  % ----------------------------------
+  % | training | validation | masked | = Own Partitioning
+  % ----------------------------------
 
-% ----------------------------------
-% | known                 | masked | = Input
-% ----------------------------------
-% | training | validation | masked | = Own Partitioning
-% ----------------------------------
+  I_training = I;
 
-I_training = I;
-% Delete pixels that we do not know
-I_training(val_indices) = 0;
+  % Delete pixels that we do not know in the training set
+  I_training(val_indices) = 0;
+  mask_training = mask;
+  mask_training(val_indices) = 0;
 
-mask_training = mask;
-mask_training(val_indices) = 0;
+  % Do Gauss interpolation
+  I_training = gaussInterpolate(I_training, mask_training, parameters);
 
-% Do Gauss interpolation
-I_training = gaussInterpolate(I_training, mask_training, parameters);
[…]

### 2011-06-21 project/EvaluateInpaintingParameterized.m
! Set repetition to 2

@@ -5,7 +5,7 @@
 % for randomness introduced by the random mask (fixed mask in the initially provided implementation)
 function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_pixels_fract)  
   file_list = dir(); 
-  rep=3;
+  rep=2;
   
   dir_length = length(dir);
 
@@ -40,7 +40,6 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
        holes = randperm(numpixels);
        holes = holes(1:floor(missing_pixels_fract*numpixels));
        mask(holes) = 0;
-
 
        I_mask = I;
        I_mask(~mask) = 0;

### 2011-06-21 project/EvaluateInpaintingParameterized.m
+ Added computer speed measurement
Signed-off-by: Jan Rüegg <rggjan@gmail.com>

@@ -63,10 +63,16 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
   % Following part is used for the optimization
   % Will store best found parameters into a file
   global global_best_cost 
+  global computer_speed
 
-  % Do not use time in cost function when below 60 seconds
-  if (mean(Times_final) > 60)
-    times_error = exp((mean(Times_final)/60)-1) - 1;
+  % Calculate speed of computer
+  if (length(computer_speed) == 0)
+    computer_speed = mean(Times_final)
+  end
+
+  % Do not use time in cost function when below 5 times starting speed seconds
+  if (mean(Times_final) > computer_speed*5)
+    times_error = exp((mean(Times_final)/(computer_speed*5))-1) - 1;
   else
     times_error = 0;
   end

### 2011-06-21 project/OptimizeInpainting.m

@@ -13,7 +13,10 @@ function OptimizeInpainting(rounds)
   load('params.mat');
 
   % Store best value in global variable, for usage in Evaluation script
+  % Also reset computer_speed
   global global_best_cost;
+  global computer_speed;
+  computer_speed = [];
   global_best_cost = best_cost;
   fprintf('\nStarting optimization, current best cost is %g\n\n', best_cost);
 

### 2011-06-21 project/OptimizeInpainting.m
! Fix minimum value

@@ -168,6 +168,8 @@ function new = getNextPatchSize(Value, Stepsize)
   % Here we work with the exponent of the power of two
   if(new > 9)
     new = 9;
+  if(new < 2)
+    new = 2;
   end
 end
 

### 2011-06-21 project/EvaluateInpaintingParameterized.m
c Indenting

@@ -20,42 +20,42 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
     Times = [];
      
     % Search for pictures
-     for i = 3 : dir_length  % running through the folder
-       file_name = file_list(i).name; % get current filename
+    for i = 3 : dir_length  % running through the folder
+      file_name = file_list(i).name; % get current filename
 
-       % Only keep the images in the loop
-       if (length(file_name) < 5)
-         continue;
-       elseif ( max(file_name(end-4:end) ~= '2.png'))
-         continue;
-       end
+      % Only keep the images in the loop
+      if (length(file_name) < 5)
+       continue;
+      elseif ( max(file_name(end-4:end) ~= '2.png'))
+       continue;
+      end
 
-       % Read image, convert to double precision and map to [0,1] interval
-       I = imread(file_name);
-       I = double(I) / 255;
+      % Read image, convert to double precision and map to [0,1] interval
+      I = imread(file_name);
+      I = double(I) / 255;
 
-       % Generate mask at random, with adjustable number missing pixels
-       mask = ones(size(I));
-       numpixels = size(I,1)*size(I,2);
-       holes = randperm(numpixels);
-       holes = holes(1:floor(missing_pixels_fract*numpixels));
-       mask(holes) = 0;
+      % Generate mask at random, with adjustable number missing pixels
+      mask = ones(size(I));
+      numpixels = size(I,1)*size(I,2);
+      holes = randperm(numpixels);
+      holes = holes(1:floor(missing_pixels_fract*numpixels));
+      mask(holes) = 0;
 
-       I_mask = I;
-       I_mask(~mask) = 0;
+      I_mask = I;
+      I_mask(~mask) = 0;
 
-       % Call the main inPainting function
-       tic;
-       I_rec = inPaintingParameterized(I_mask, mask, parameters);
-       Times = [Times toc];
+      % Call the main inPainting function
+      tic;
+      I_rec = inPaintingParameterized(I_mask, mask, parameters);
+      Times = [Times toc];
 
-       % Measure approximation error
-       Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
-     end
+      % Measure approximation error
+      Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
+    end
 
-     % Collect the errors and the times
-     Errors_final = [Errors_final Errors];
-     Times_final = [Times_final Times];
[…]

### 2011-06-21 project/OptimizeInpainting.m
! Fix

@@ -168,6 +168,7 @@ function new = getNextPatchSize(Value, Stepsize)
   % Here we work with the exponent of the power of two
   if(new > 9)
     new = 9;
+  end
   if(new < 2)
     new = 2;
   end

### 2011-06-21 project/EvaluateInpaintingParameterized.m
+ Removed ;

@@ -72,7 +72,7 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
 
   % Do not use time in cost function when below 5 times starting speed seconds
   if (mean(Times_final) > computer_speed*5)
-    times_error = exp((mean(Times_final)/(computer_speed*5))-1) - 1;
+    times_error = exp((mean(Times_final)/(computer_speed*5))-1) - 1
   else
     times_error = 0;
   end

### 2011-06-21 project/Paper/paper.tex
Paper: Global parameter optimization completed.

--- 2011-06-20 project/Paper/paper.tex
@@
 \caption{Our newly proposed algorithm as a flowchart.}
+\label{flowchart}
 \end{figure}
@@
 The input of our algorithm is the image to be restored together with a mask that indicates its unknown spots.\\
-Before entering the iterative part of our algorithm, the following three steps are performed (as seen in figure ??):
+Before entering the iterative part of our algorithm, the following three steps are performed (as seen in figure \ref{flowchart}):
 \begin{enumerate}
 \item In order to guide the elimination of frequencies by finding the optimal threshold (as described below), we first remove additional pixels from the image and use them as the validation set.
-The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section ??).
+The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section \ref{gradient_descent}).
 \item Since our reduction in the Fourier space works best with framed patches, we frame the complete image by mirroring its margin. (The size of the frame is another optimizable global parameter.)
@@
 \subsection{Global parameter optimization}
-As shown in table \ref{parameters}, there are a total of 10 parameters in our algorithm that can be set.
-For many of them, it is quite clear what values should be used.
-For example, the size of the gauss kernel should be a bit larger than the biggest assumed hole in the image.
-For others, like the maximum number of iterations, it is not so clear.
-And yet others have a clear speed / accuracy tradeoff, like the stopping conditions of the threshold optimization.\\
-Because we thought it not feasable to tune all the parameters by hand, we wrote an algorithm to do the global parameter optimization.
-The idea is to do this once for a few representative images, and get parameters that can model as closely as possible many of the images we are interested in.
-Our proposed method combines a gradient descent method with ideas from cross-validation, and tunes our 10-dimensional parameter-space, minimizing time and maximizing accuracy.
+\subsubsection{Optimized parameters}
+As shown in table \ref{parameters}, our algorithm has a total of 10 parameters that can be set globally.
+Instead of tuning them by hand, we automated the optimization process and implicitly use the best parameters found so far, which we store in the file system.
+This feature also allows us to interrupt the optimization process at any time and continue with the best parameters later on.
 
-\subsubsection{Error function}
-When doing a minimization of any kind, an error function is needed.
-Of course we already had the error function of the evaluation script, that gives a mean over the squared differences of the reconstructed to the original image.
-However, using only an error measure, the time would potentially grow to infinity to get the best accuracy possible during tuning.\\
-Therefore we tried different error function including both, the image error $E$, and the time $T$ spent (see table \ref{error_function_table}).
-The image error here is the average squared error of a reconstructed pixel to the original pixel, scaled with a factor (usually about 10000), while the time is measured in seconds.
+\begin{table*}
+\begin{center}
+\begin{tabular}{|l|l|r|r|}
+\hline
+Parameter & Description & Initial value & Optimized value\\
+\hline
+gauss\_size & Size of the Gaussian kernel & ? & ? \\
+gauss\_sigma & Sigma of the Gaussian kernel & ? & ? \\
+patch\_size & Patch size as exponent to the basis 2 & ? & ? \\
+patch\_frame\_size & Frame size in pixels & ? & ? \\
+td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds & ? & ? \\
+td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds & ? & ? \\
+td\_middle & Initial threshold & ? & ? \\
+validation & Fraction of pixels in the validation set & ? & ? \\
+max\_iterations & Maximum number of iterations & ? & ? \\
+abortbelow\_change & Minimal change for performing another iteration & ? & ? \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Parameters with their initial and optimized values.}
+\label{parameters}
+\end{table*}
+
+\subsubsection{Gradient descent}
+\label{gradient_descent}
+We optimize all parameters at once using a method called gradient descent.
+Since our cost function is based on measurements, we cannot derive it with respect to individual parameters and have to determine the gradient with finite differences.
+Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term that should help us to overcome local minima.
+
+\subsubsection{Cost function}
+For optimizing a set of parameters, a cost function is needed.
+Its most important component is obviously the error function used in the submission system, which calculates the mean over the squared differences between the reconstructed and the original image.
+However, without including the performance of the algorithm, the runtime could grow arbitrarily to get the best accuracy possible.\\
+We considered several cost functions based on the error $E$ and the time $T$ (see table \ref{error_functions}).
+These two values are scaled appropriately: Usually, the error is multiplied by a factor of 10'000 and the time (measured in seconds) by a factor of 0.01.
 
@@
 \centering
-  \begin{tabular}{|l|p{5cm}|p{5cm}|}
+\begin{tabular}{|l|p{6cm}|p{6cm}|}
 \hline
-    Function&Advantages&Disadvantages\\
+Function & Advantages & Disadvantages \\
 \hline
-    $\exp(E)+\exp(T)$ & Increases cost fast to avoid $E$ or $T$ over a certain threshold & Not enough gradient for very low $E$, $T$ values\\
+$E + T$ & Easy to understand and everywhere a good gradient & Linear improvement does not necessarily follow one's intuition about the ordering of results \\
 \hline
-    $(1-\exp(E))*(1-\exp(T))$ & Makes it possible to have large times for very low errors & Works also the other way around, and optimization only minimizes time\\
+$\exp(E)+\exp(T)$ & Fast increase for both error and time & Not enough gradient for very low $E$ and $T$ values \\
 \hline
-    $E + T$ & Everywhere a good gradient & Time from 60 to 59 seconds is same \textit{improvement} as time from 2 to 1 second\\
+$(1-\exp(E))*(1-\exp(T))$ & Allows very large times for very low errors & Also true vice versa, i.e. a minimization in time \\
 \hline
@@
 E+\exp(T/60 - 1) - 1 & \text{otherwise}
-\end{cases}$ & Optimizes for error only when time is low enough. Function still continous, penalizes high times. & None so far, this is the function used by us\\
+\end{cases}$ & Continuous function that optimizes for error only when the time is below a certain value but penalizes long times. & Seems to work quite well and therefore used by us \\
 \hline
 \end{tabular}
-  \caption{Different combined error functions}
-  \label{error_function_table}
+\caption{Various cost functions used in the optimization process.}
+\label{error_functions}
 \end{table*}
 
-% TODO maybe add 3d plot for some of the error functions
+\subsubsection{Sample data}
+We used the three images provided for the last exercise (consisting of both natural and artificial images) to evaluate the accuracy and performance of the chosen parameters.
+Instead of using the supplied mask, however, we generate one randomly for every evaluation in order to prevent overfitting to a particular mask.
 
-\subsubsection{Gradient descent}
-The idea behind our gradient descent is to optimize all the parameters at once, instead of fine-tuning each of them separately. For that, we
-calculate the derivative of the error function depending on one of our 10 parameters, and then move the parameter into the direction where the
-error decreases fastest. We also consider the slope, and make big changes if the derivative is big, and small changes if the derivative is small.
-Iteratively, we calculate all of the 10 derivatives, change all the parameters and start over again.
-
-Since the error and time are measured, we cannot algebraically determine the gradient of the error function. Instead, we do a discrete estimation,
-measuring the error with finite differences. The error for parameter values plus and minus a small delta, divided by two, gives us our estimated
-derivative.
-
-In figure \ref{}
+\subsubsection{Found values}
+Our observations during the optimization process can be split into three classes:
+\begin{enumerate}
+\item Some parameters converged towards their expected value. The size of the Gaussian kernel, for example, has to be slightly bigger than the largest hole assumed in an image.
+\item Some parameters have a clear speed vs. accuracy tradeoff, like the stopping conditions of the threshold optimization, which depend on our cost function and could not be foreseen.
+\item Some parameters assumed totally different values than anticipated: The patch size became much larger and the validation set decreased significantly.
+\end{enumerate}
 
@@
 
-\subsection{Figures and Tables}
-
 \paragraph{Note}

### 2011-06-21 project/baseline1/inPainting.m
Correct linear inpainting

@@ -28,7 +28,7 @@ for x=1:w
       s = 0;
 
       % Try-Catch to ignore boundary issues
-      for i=1:10
+      for i=1:w
         try
           if mask(x+i, y) ~= 0
             values = values + I(x+i, y)/i;
@@ -40,7 +40,7 @@ for x=1:w
         end
       end
 
-      for i=1:10
+      for i=1:h
         try
           if mask(x, y+i) ~= 0
             values = values + I(x, y+i)/i;
@@ -52,7 +52,7 @@ for x=1:w
         end
       end
 
-      for i=1:10
+      for i=1:h
         try
           if mask(x, y-i) ~= 0
             values = values + I(x, y-i)/i;
@@ -64,7 +64,7 @@ for x=1:w
         end
       end
 
-      for i=1:10
+      for i=1:w
         try
           if mask(x-i, y) ~= 0
             values = values + I(x-i, y)/i;

### 2011-06-21 project/Paper/paper.tex

@@
 Its most important component is obviously the error function used in the submission system, which calculates the mean over the squared differences between the reconstructed and the original image.
-However, without including the performance of the algorithm, the runtime could grow arbitrarily to get the best accuracy possible.\\
+However, without including the performance of the algorithm, the runtime could grow arbitrarily to get the best accuracy possible.
 We considered several cost functions based on the error $E$ and the time $T$ (see table \ref{error_functions}).
@@
 Our observations during the optimization process can be split into three classes:
-\begin{enumerate}
+\begin{itemize}
 \item Some parameters converged towards their expected value. The size of the Gaussian kernel, for example, has to be slightly bigger than the largest hole assumed in an image.
@@
 \item Some parameters assumed totally different values than anticipated: The patch size became much larger and the validation set decreased significantly.
-\end{enumerate}
+\end{itemize}
 

### 2011-06-21 project/CreatePaperPlots.m
Minor debug output fix

@@ -30,7 +30,7 @@ if(~exist('plots/error_A.mat', 'file'))
   end
   save('plots/error_A.mat', 'error_algo');
 else
-  fprintf('Found saved data file. Loading...')
+  fprintf('Found saved data file. Loading...\n')
   load('plots/error_A.mat');
 end
 
@@ -44,7 +44,7 @@ if(~exist('plots/error_B1.mat', 'file'))
   cd(main_path)
   save('plots/error_B1.mat', 'error_base1');
 else
-  fprintf('Found saved data file. Loading...')
+  fprintf('Found saved data file. Loading...\n')
   load('plots/error_B1.mat');
 end
 
@@ -58,7 +58,7 @@ if(~exist('plots/error_B2.mat', 'file'))
   cd(main_path)
   save('plots/error_B2.mat', 'error_base2');
 else
-  fprintf('Found saved data file. Loading...')
+  fprintf('Found saved data file. Loading...\n')
   load('plots/error_B2.mat');
 end
 
@@ -72,7 +72,7 @@ if(~exist('plots/error_B3.mat', 'file'))
   cd(main_path)
   save('plots/error_B3.mat', 'error_base3');
 else
-  fprintf('Found saved data file. Loading...')
+  fprintf('Found saved data file. Loading...\n')
   load('plots/error_B3.mat');
 end
 

### 2011-06-21 project/Paper/paper.tex
Filled in start values for parameters

@@
 \begin{center}
-\begin{tabular}{|l|l|r|r|}
+\begin{tabular}{|l|p{6cm}|r|r|}
 \hline
@@
 \hline
-gauss\_size & Size of the Gaussian kernel & ? & ? \\
-gauss\_sigma & Sigma of the Gaussian kernel & ? & ? \\
-patch\_size & Patch size as exponent to the basis 2 & ? & ? \\
-patch\_frame\_size & Frame size in pixels & ? & ? \\
-td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds & ? & ? \\
-td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds & ? & ? \\
-td\_middle & Initial threshold & ? & ? \\
-validation & Fraction of pixels in the validation set & ? & ? \\
-max\_iterations & Maximum number of iterations & ? & ? \\
-abortbelow\_change & Minimal change for performing another iteration & ? & ? \\
+gauss\_size & Size of the Gaussian kernel in pixels & 10 & ? \\
+gauss\_sigma & Sigma of the Gaussian kernel & 1 & ? \\
+patch\_size & Patch size as exponent to the basis 2 & 4 & ? \\
+patch\_frame\_size & Frame size in pixels & 8 & ? \\
+td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds as negative exponent to basis 10& 3 & ? \\
+td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds as negative exponent to basis 10& 4 & ? \\
+td\_middle & Initial threshold & 10 & ? \\
+validation & Fraction of pixels in the validation set & 0.4 & ? \\
+max\_iterations & Maximum number of iterations & 5 & ? \\
+abortbelow\_change & Minimal relative change to perform another iteration & 0.1 & ? \\
 \hline

### 2011-06-21 project/CreatePaperPlots.m
+ Added optimization to plot generation

@@ -87,3 +87,8 @@ plot(range, error_algo, 'r', range, error_base1,'b', range, error_base2, 'g', ra
 legend('New algorithm', 'Baseline 1 (Linear Interpolation)', 'Baseline 2 (Matching Pursuit)', 'Baseline 3 (Gaussian Interpolation)');
 saveas(handle, 'plots/missingpixelsVsError.fig');
 hold off;
+
+% Generate curves and search strategy plots
+load('start_params.mat');
+save('params.mat', 'best_cost', 'parameters');
+OptimizeInpainting(100);

### 2011-06-21 project/Paper/paper.tex
Paper: Plots to generate.

--- 2011-06-21 project/Paper/paper.tex
@@
 \subsubsection{Sample data}
-We used the three images provided for the last exercise (consisting of both natural and artificial images) to evaluate the accuracy and performance of the chosen parameters.
+We used the three images provided for the last exercise and four additional ones (consisting of both natural and artificial images) to evaluate the accuracy and performance of the chosen parameters.
 Instead of using the supplied mask, however, we generate one randomly for every evaluation in order to prevent overfitting to a particular mask.
@@
 \begin{itemize}
-\item Some parameters converged towards their expected value. The size of the Gaussian kernel, for example, has to be slightly bigger than the largest hole assumed in an image.
+\item Some parameters converged towards their estimated value. The size of the Gaussian kernel, for example, has to be slightly bigger than the largest hole assumed in an image.
 \item Some parameters have a clear speed vs. accuracy tradeoff, like the stopping conditions of the threshold optimization, which depend on our cost function and could not be foreseen.
@@
 \section{Results}
-
- ((( TODO )))
-
- We compare our algorithm with the following two baseline algorithms:
+We compared our algorithm with the following three baseline algorithms:
 \begin{enumerate}
@@
 \item A matching pursuit algorithm \cite{matchingpursuit93} with an overcomplete dictionary built with Discrete Cosine transforms.
+\item Gaussian interpolation...
 \end{enumerate}
-
-  Our algorithm is resistant to high percentages of missing pixels, the average error per missing pixel rises only slowly until ((( TODO level))).
 
@@
 For evaluation, we tested these algorithms with different levels of missing pixels from 0 to 100\% in 5\% steps.
-
- ((( TODO )))
-  Show evidence to support your claims made in the
-  introduction.
 
@@
 want to address in the text portion of the results.
-The information from the figures is
-summarized in Table~\ref{tab:fourier-wavelet}.
 
@@
 
-You compare your novel algorithm to \emph{at least two baseline
-  algorithms}. For the baselines, you can use the implementations you
-developed as part of the programming assignments.
+\paragraph{Note} All plots in this paper were created with MATLAB and can be reproduced using the script \emph{CreatePaperPlots.m} of our submission.
 
 \section{Discussion}
-  Discuss the strengths and weaknesses of your
-  approach, based on the results. Point out the implications of your  
-  novel idea on the application concerned. (((( TODO ))))
+Our algorithm is resistant to high percentages of missing pixels, the average error per missing pixel rises only slowly until ((( TODO level))).
 
-\paragraph{Note}
-All plots in this paper were created with MATLAB and can be reproduced using the script 'CreatePaperPosts.m' in our submission.\\
+Discuss the strengths and weaknesses of your approach, based on the results. Point out the implications of your novel idea on the application concerned. (((( TODO ))))
 
@@
 \caption{Artifacts that occur without a frame (frame size = 0). The patches are recognizable in the reconstruction}
-  \label{fig:boundaryArtifacts}
+\label{boundaryArtifacts}
 \end{figure}
@@
 \section{Summary}
-(((( TODO ))))
-  Summarise your contributions in light of the new
-  results.
+Summarise your contributions in light of the new results.
 
 \section*{Acknowledgements}
-We would like to recognize and thank the following individuals or groups for their contribution to this work:\\
-Our implementation is based on the inpainting and evaluation framework provided by the course "Computer Intelligence Lab" held by Prof. Joachim Buhmann at ETH Zürich.
-(((( TODO ))))
+Our implementation is based on the inpainting and evaluation framework provided by the course \emph{Computer Intelligence Lab} held by Prof. Joachim Buhmann at ETH Zurich.
 
@@
 \end{document}
+
+% Plots:
+% - threshold optimization => algorithm
+% - Framing artifacts (with and w/o frame) => framing
+% - Parameter optimization (convergence) => gradient descent
+% 
+% - Missing pixels vs. error (baselines)
+% - stdev over n runs of EvaluateInpainting with rep = 1
+% 
+% - image with best error (and stdev)
+% - image with worst error (and stdev)
+% 
+% - Bar to illustrate edge reconstruction (patch size)

### 2011-06-21 project/CreatePaperPlots.m
! Create plots in CreatePaperPlots

@@ -7,6 +7,12 @@
 
 load('params.mat');
 
+% Generate strategy plots
+global debug_threshold_plot_number;
+debug_threshold_plot_number = 10;
+EvaluateInpaintingParameterized(parameters, 0.6);
+
+% Generate graphs
 stepsize = 2;
 no_steps = floor(100/stepsize);
 

### 2011-06-21 project/OptimizeInpainting.m

@@ -16,8 +16,6 @@ function OptimizeInpainting(rounds)
   % Also reset computer_speed
   global global_best_cost;
   global computer_speed;
-  global debug_threshold_plot_number;
-  debug_threshold_plot_number = 10;
   computer_speed = [];
   global_best_cost = best_cost;
   fprintf('\nStarting optimization, current best cost is %g\n\n', best_cost);

### 2011-06-21 project/OptimizeInpainting.m
+ Create figures without a display

@@ -58,13 +58,13 @@ function OptimizeInpainting(rounds)
 
     % Plot cost function
     cost_list = [cost_list, cost];
-    figure(9);
+    f = figure('visible', 'off');
     plot(cost_list);
     title('Cost during optimization');
     xlabel('Steps');
     ylabel('Cost function');
-    saveas(9, 'plots/cost_plot.fig');
-    drawnow
+    saveas(f, 'plots/cost_plot.fig');
+    close(f);
 
     rounds_done = rounds_done + 1;
   end
@@ -137,13 +137,13 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   % Plot to graph
   global parameter_list
   parameter_list(index, rounds_done+1) = new_value;
-  figure(index);
+  f = figure(index, 'visible', 'off');
   plot(parameter_list(index, :));
   title(sprintf('%s during optimization', fields{index}), 'Interpreter', 'none');
   xlabel('Steps');
   ylabel(sprintf('Value of %s', fields{index}), 'Interpreter', 'none');
-  saveas(index, sprintf('plots/%s_plot.fig', fields{index}));
-  drawnow;
+  saveas(f, sprintf('plots/%s_plot.fig', fields{index}));
+  close(f);
 end
 
 % ================= %

### 2011-06-21 project/determineThresholds.m

@@ -121,7 +121,7 @@ function [T, I_trained] = determineThresholds(I_training_framed, val_mask, val_I
         global debug_threshold_plot_number
         if length(debug_threshold_plot_number) > 0 && debug_threshold_plot_number > 0
           % Plot
-          figure(20);
+          f = figure('visible', 'off');
           plot(steps,errors);
           hold on;
           plot(middles,middles_values, 'r');
@@ -131,7 +131,8 @@ function [T, I_trained] = determineThresholds(I_training_framed, val_mask, val_I
           title('Search strategy for minimum')
           legend('error', 'search strategy')
           hold off;
-          saveas(20, sprintf('plots/search_strategy_%g', debug_threshold_plot_number));
+          saveas(f, sprintf('plots/search_strategy_%g', debug_threshold_plot_number));
+          close(f);
           debug_threshold_plot_number = debug_threshold_plot_number - 1;
         end
       end

### 2011-06-21 project/OptimizeInpainting.m
! Bugfixes

@@ -137,7 +137,7 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   % Plot to graph
   global parameter_list
   parameter_list(index, rounds_done+1) = new_value;
-  f = figure(index, 'visible', 'off');
+  f = figure('visible', 'off');
   plot(parameter_list(index, :));
   title(sprintf('%s during optimization', fields{index}), 'Interpreter', 'none');
   xlabel('Steps');

### 2011-06-21 project/determineThresholds.m

@@ -99,7 +99,8 @@ function [T, I_trained] = determineThresholds(I_training_framed, val_mask, val_I
         middles_values = [middles_values errors(2)];
       end
 
-      if(stepsize < 0.1)
+      global debug_threshold_plot_number
+      if length(debug_threshold_plot_number) > 0 && debug_threshold_plot_number > 0 && stepsize < 0.1
         % DEBUG Creates some nice graphs of the evolution of the threshold
         errors = [];
         steps=[];
@@ -118,23 +119,20 @@ function [T, I_trained] = determineThresholds(I_training_framed, val_mask, val_I
           steps = [steps, t];
         end
         
-        global debug_threshold_plot_number
-        if length(debug_threshold_plot_number) > 0 && debug_threshold_plot_number > 0
-          % Plot
-          f = figure('visible', 'off');
-          plot(steps,errors);
-          hold on;
-          plot(middles,middles_values, 'r');
-          plot(middles,middles_values, 'or');
-          xlabel('Threshold')
-          ylabel('Error in validation set')
-          title('Search strategy for minimum')
-          legend('error', 'search strategy')
-          hold off;
-          saveas(f, sprintf('plots/search_strategy_%g', debug_threshold_plot_number));
-          close(f);
-          debug_threshold_plot_number = debug_threshold_plot_number - 1;
-        end
+        % Plot
+        f = figure('visible', 'off');
+        plot(steps,errors);
+        hold on;
+        plot(middles,middles_values, 'r');
+        plot(middles,middles_values, 'or');
+        xlabel('Threshold')
+        ylabel('Error in validation set')
+        title('Search strategy for minimum')
+        legend('error', 'search strategy')
+        hold off;
+        saveas(f, sprintf('plots/search_strategy_%g', debug_threshold_plot_number));
+        close(f);
+        debug_threshold_plot_number = debug_threshold_plot_number - 1;
       end
       
       % Save the threshold for this patch

### 2011-06-21 project/EvaluateInpaintingParameterized.m
Prepare for stdev

@@ -3,9 +3,8 @@
 % Measure approximation error for several images.
 % Adjusted function of the framework thatrepeats the evaluation step to compensate
 % for randomness introduced by the random mask (fixed mask in the initially provided implementation)
-function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_pixels_fract)  
+function [cost, avgQErr, stdev] = EvaluateInpaintingParameterized(parameters, missing_pixels_fract, rep)  
   file_list = dir(); 
-  rep=2;
   
   dir_length = length(dir);
 
@@ -57,6 +56,32 @@ function [cost, avgQErr] = EvaluateInpaintingParameterized(parameters, missing_p
     Errors_final = [Errors_final Errors];
     Times_final = [Times_final Times];
   end
+  
+  % Determine mean and stdev per image
+  % Count images first
+  imc = 0;
+  for i = 3 : dir_length
+    file_name = file_list(i).name; % get current filename
+    % Only keep the images in the loop
+      if (length(file_name) < 5)
+       continue;
+      elseif ( max(file_name(end-4:end) ~= '2.png'))
+       continue;
+      end
+      imc = imc+1;
+  end
+  
+  stdev = zeros(imc,1);
+  if rep > 1
+    for i=1:imc
+      stdev(imc,1) = std(Errors_final(i:imc:end));
+    end
+  else
+    % 1 Iteration - no deviation
+    stdev = zeros(imc,1);
+  end
+  
+    
   
   avgQErr = mean(Errors_final);
  

### 2011-06-21 project/OptimizeInpainting.m

@@ -5,6 +5,7 @@
 % Saves plots of the cost and the changes of parameters to the plots folder
 function OptimizeInpainting(rounds)
   missing_pixels = 0.6;
+  repetition_evaluation = 2;
 
   % Set random seet to get reproducable results
   rand('seed', 12345);
@@ -38,7 +39,7 @@ function OptimizeInpainting(rounds)
     fprintf('========== Starting round %g ===========\n', rounds_done)
 
     parameters = final_parameters
-    [cost, unused] = EvaluateInpaintingParameterized(parameters, missing_pixels);
+    [cost, unused] = EvaluateInpaintingParameterized(parameters, missing_pixels, repetition_evaluation);
     fprintf('=> cost %g\n\n', cost);
 
     % Optimize individual parameters
@@ -101,7 +102,7 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   param_cell{index} = getNext(param_cell{index}, 1);
   new_parameters = cell2struct(param_cell, fields, 1);
   fprintf('%s: %g ... ', fields{index}, param_cell{index})
-  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels);
+  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels, repetition_evaluation);
   new_cost_plus = new_cost - cost;
   fprintf('cost %g\n', new_cost_plus);
 
@@ -110,7 +111,7 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   param_cell{index} = getNext(param_cell{index}, -1);
   new_parameters = cell2struct(param_cell, fields, 1);
   fprintf('%s: %g ... ', fields{index}, param_cell{index})
-  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels);
+  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels, repetition_evaluation);
   new_cost_minus = new_cost - cost;
   fprintf('cost %g\n', new_cost_minus);
 

### 2011-06-21 project/Paper/paper.tex
Paper: Included plots templates.

@@
 The abstract should really be written last, along with the title of
-the paper. The four points that should be covered~\cite{jones08}:
+the paper. The four points that should be covered:
 \begin{enumerate}
@@
 The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section \ref{gradient_descent}).
-\item Since our reduction in the Fourier space works best with framed patches, we frame the complete image by mirroring its margin. (The size of the frame is another optimizable global parameter.)
+\item Since our reduction in the Fourier space works best with framed patches (see figure \ref{framing_artifacts}), we frame the complete image by mirroring its margin.
+(The size of the frame is another optimizable global parameter.)
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\caption{Artifacts occur w/o frames (left) that vanish with frames (right).}
+\label{framing_artifacts}
+\end{figure}
+
 \item For obtaining an initial reconstruction of the image, we compute a weighted sum over all the neighboring pixels that are known.
@@
 (This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.)
+\label{gaussian_interpolation}
 \end{enumerate}
@@
 Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to perform as few threshold evaluations as possible.
-For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and decreasing the offset.
+For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and decreasing the step size (see figure \ref{threshold_optimization}).
 For this optimization as well as for the whole iteration, we have both a number- and a convergence-based termination criterion.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\caption{The threshold is optimized by iteratively taking the best one.}
+\label{threshold_optimization}
+\end{figure}
+
 \item Having stored the reconstruction of every patch with its best threshold, all that remains to do is copying the newly estimated missing pixels (including those of the validation set) back into the reconstructed image.
@@
 Since our cost function is based on measurements, we cannot derive it with respect to individual parameters and have to determine the gradient with finite differences.
-Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term that should help us to overcome local minima.
+Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term that should help us to overcome local minima (see figure \ref{gradient_descent_figure}).
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\caption{Due to the momentum term, the cost function does not strictly decrease.}
+\label{gradient_descent_figure}
+\end{figure}
 
@@
 However, without including the performance of the algorithm, the runtime could grow arbitrarily to get the best accuracy possible.
-We considered several cost functions based on the error $E$ and the time $T$ (see table \ref{error_functions}).
+We considered several cost functions based on the error $E$ and the time $T$ (see table \ref{cost_functions}).
 These two values are scaled appropriately: Usually, the error is multiplied by a factor of 10'000 and the time (measured in seconds) by a factor of 0.01.
@@
 \caption{Various cost functions used in the optimization process.}
-\label{error_functions}
+\label{cost_functions}
 \end{table*}
@@
 \section{Results}
-We compared our algorithm with the following three baseline algorithms:
+We compared our algorithm with the following three baseline algorithms (see figure \ref{baseline_algorithms}):
 \begin{enumerate}
-\item A simple linear inpainting algorithm using the nearest neighbor pixels.
-\item A matching pursuit algorithm \cite{matchingpursuit93} with an overcomplete dictionary built with Discrete Cosine transforms.
-\item Gaussian interpolation...
+\item Vertical and horizontal linear interpolation,
+\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines and
+\item Gaussian interpolation as described in section \ref{gaussian_interpolation}.
 \end{enumerate}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\caption{Comparison with baseline algorithms for various fractions of missing pixels.}
+\label{baseline_algorithms}
+\end{figure}
 
@@
 Discuss the strengths and weaknesses of your approach, based on the results. Point out the implications of your novel idea on the application concerned. (((( TODO ))))
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=\columnwidth]{images/boundaryArtifact_noframe.png}
-\caption{Artifacts that occur without a frame (frame size = 0). The patches are recognizable in the reconstruction}
-\label{boundaryArtifacts}
-\end{figure}
-
-((( TODO: Insert plot MissingPixels vs. Accuracy )))
-
-Use examples and illustrations to clarify ideas and results. For
-example, by comparing Figure~\ref{fig:framing} and
-Figure~\ref{fig:flowchart}, we can see the two different
-situations where Fourier and wavelet basis perform well.
 

### 2011-06-21 project/OptimizeInpainting.m
Handle new output of EvaluateInpaintingParameterized

@@ -39,7 +39,7 @@ function OptimizeInpainting(rounds)
     fprintf('========== Starting round %g ===========\n', rounds_done)
 
     parameters = final_parameters
-    [cost, unused] = EvaluateInpaintingParameterized(parameters, missing_pixels, repetition_evaluation);
+    [cost, unused, unused2] = EvaluateInpaintingParameterized(parameters, missing_pixels, repetition_evaluation);
     fprintf('=> cost %g\n\n', cost);
 
     % Optimize individual parameters
@@ -102,7 +102,7 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   param_cell{index} = getNext(param_cell{index}, 1);
   new_parameters = cell2struct(param_cell, fields, 1);
   fprintf('%s: %g ... ', fields{index}, param_cell{index})
-  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels, repetition_evaluation);
+  [new_cost, unused, unused2] = EvaluateInpaintingParameterized(new_parameters, missing_pixels, repetition_evaluation);
   new_cost_plus = new_cost - cost;
   fprintf('cost %g\n', new_cost_plus);
 
@@ -111,7 +111,7 @@ function [new_value, next_old] = gradientDescent(index, getNext, parameters, old
   param_cell{index} = getNext(param_cell{index}, -1);
   new_parameters = cell2struct(param_cell, fields, 1);
   fprintf('%s: %g ... ', fields{index}, param_cell{index})
-  [new_cost, unused] = EvaluateInpaintingParameterized(new_parameters, missing_pixels, repetition_evaluation);
+  [new_cost, unused, unused2] = EvaluateInpaintingParameterized(new_parameters, missing_pixels, repetition_evaluation);
   new_cost_minus = new_cost - cost;
   fprintf('cost %g\n', new_cost_minus);
 

### 2011-06-21 project/EvaluateInpaintingParameterized.m
Add filenames to returned stdevs

@@ -60,6 +60,7 @@ function [cost, avgQErr, stdev] = EvaluateInpaintingParameterized(parameters, mi
   % Determine mean and stdev per image
   % Count images first
   imc = 0;
+  filenames = {};
   for i = 3 : dir_length
     file_name = file_list(i).name; % get current filename
     % Only keep the images in the loop
@@ -68,17 +69,24 @@ function [cost, avgQErr, stdev] = EvaluateInpaintingParameterized(parameters, mi
       elseif ( max(file_name(end-4:end) ~= '2.png'))
        continue;
       end
+      %Count and save filenames for later
       imc = imc+1;
+      filenames{imc} = {file_name};
   end
   
-  stdev = zeros(imc,1);
+  for i=1:imc
+    stdev{i,2} = filenames{i};
+  end
+  
   if rep > 1
     for i=1:imc
-      stdev(imc,1) = std(Errors_final(i:imc:end));
+      stdev{i,1} = std(Errors_final(i:imc:end));
     end
   else
     % 1 Iteration - no deviation
-    stdev = zeros(imc,1);
+    for i=1:imc
+      stdev{i,1} = 0;
+    end
   end
   
     

### 2011-06-21 project/CreatePaperPlots.m
Added more output

@@ -9,6 +9,7 @@ load('params.mat');
 
 % Generate strategy plots
 global debug_threshold_plot_number;
+fprintf('1) Creating Parameter evolution graphs')
 debug_threshold_plot_number = 10;
 EvaluateInpaintingParameterized(parameters, 0.6);
 
@@ -24,8 +25,8 @@ error_base3 = zeros(no_steps, 1);
 
 main_path = pwd;
 
-fprintf('1) Calculating error vs. missing pixels\n')
-fprintf('1A) New algorithm\n')
+fprintf('2) Calculating error vs. missing pixels\n')
+fprintf('2A) New algorithm\n')
 
 if(~exist('plots/error_A.mat', 'file'))
   % Use parallel computation for this, if available
@@ -40,7 +41,7 @@ else
   load('plots/error_A.mat');
 end
 
-fprintf('1B) Baseline 1\n')
+fprintf('2B) Baseline 1\n')
 if(~exist('plots/error_B1.mat', 'file'))    
   cd('baseline1/')
   parfor k=0:no_steps
@@ -54,7 +55,7 @@ else
   load('plots/error_B1.mat');
 end
 
-fprintf('1C) Baseline 2\n')
+fprintf('2C) Baseline 2\n')
 if(~exist('plots/error_B2.mat', 'file'))    
   cd('baseline2/')
   parfor k=0:no_steps
@@ -68,7 +69,7 @@ else
   load('plots/error_B2.mat');
 end
 
-  fprintf('1D) Baseline 3\n')
+  fprintf('2D) Baseline 3\n')
 if(~exist('plots/error_B3.mat', 'file'))    
   cd('baseline3/')
   parfor k=0:no_steps
@@ -82,7 +83,7 @@ else
   load('plots/error_B3.mat');
 end
 
-fprintf('1) Plotting\n')
+fprintf('2E) Plotting\n')
 
 handle = figure;
 range = 0:stepsize:100;

### 2011-06-21 project/baseline1/EvaluateInpaintingParameterized.m

@@ -3,55 +3,44 @@
 function [avgQErr] = EvaluateInpaintingParameterized(missing_pixels_fract)  
 
   file_list = dir('..'); 
-  rep=3;
-  
   dir_length = length(dir('..'));
 
-  Errors_final = [];
-  LTimes_final = [];
+  Errors = []; % mean squared errors for each image
+  
+  for i = 3 : dir_length  % running through the folder
+    file_name = file_list(i).name; % get current filename
 
-  parfor r = 1 : rep
-     Errors = []; % mean squared errors for each image
-     LTimes = [];
-     for i = 3 : dir_length  % running through the folder
+    % Only keep the images in the loop
+    if (length(file_name) < 5)
+      continue;
+    elseif ( max(file_name(end-4:end) ~= '2.png'))
+      continue;
+    end
 
-        file_name = file_list(i).name; % get current filename
+    % Read image, convert to double precision and map to [0,1] interval
+    I = imread(strcat('../',file_name));
+    I = double(I) / 255;
 
-        % Only keep the images in the loop
-        if (length(file_name) < 5)
-          continue;
-        elseif ( max(file_name(end-4:end) ~= '2.png'))
-          continue;
-        end
-
-        % Read image, convert to double precision and map to [0,1] interval
-        I = imread(strcat('../',file_name));
-        I = double(I) / 255;
-
-        % Generate mask at random, with 60% missing pixels
-        mask = ones(size(I));
-        numpixels = size(I,1)*size(I,2);
-        holes = randperm(numpixels);
-        holes = holes(1:floor(missing_pixels_fract*numpixels));
-        mask(holes) = 0;
+    % Generate mask at random, with 60% missing pixels
+    mask = ones(size(I));
+    numpixels = size(I,1)*size(I,2);
+    holes = randperm(numpixels);
+    holes = holes(1:floor(missing_pixels_fract*numpixels));
+    mask(holes) = 0;
 
 
-        I_mask = I;
-        I_mask(~mask) = 0;
+    I_mask = I;
+    I_mask(~mask) = 0;
 
-        % Call the main inPainting function
-        tic;
-        I_rec = inPainting(I_mask, mask);
-        LTimes = [LTimes toc];
+    % Call the main inPainting function
+    I_rec = inPainting(I_mask, mask);
 
-        % Measure approximation error
[…]

### 2011-06-21 project/CreatePaperPlots.m
+ Finished graph creation

@@ -11,6 +11,10 @@ load('params.mat');
 global debug_threshold_plot_number;
 debug_threshold_plot_number = 10;
 EvaluateInpaintingParameterized(parameters, 0.6);
+for i = 1:10
+  f = open(sprintf('plots/search_strategy_%g.fig', i));
+  plotpdftex(f, sprintf('plots/search_strategy_%g', i));
+end
 
 % Generate graphs
 stepsize = 2;

### 2011-06-21 project/determineThresholds.m

@@ -100,11 +100,11 @@ function [T, I_trained] = determineThresholds(I_training_framed, val_mask, val_I
       end
 
       global debug_threshold_plot_number
-      if length(debug_threshold_plot_number) > 0 && debug_threshold_plot_number > 0 && stepsize < 0.1
+      if length(debug_threshold_plot_number) > 0 && debug_threshold_plot_number > 0 && stepsize < 0.1 && middle > debug_threshold_plot_number*2
         % DEBUG Creates some nice graphs of the evolution of the threshold
         errors = [];
         steps=[];
-        for t = linspace(0, parameters.td_middle*2, 100)
+        for t = linspace(0, parameters.td_middle*2, 200)
           P_reduced = removeFrame(dimensionReduction(P_framed, t, parameters), ...
                                   parameters);
 

### 2011-06-21 project/Paper/paper.tex
Paper: Outline of results completed.

@@
 \item Vertical and horizontal linear interpolation,
-\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines and
-\item Gaussian interpolation as described in section \ref{gaussian_interpolation}.
+\item Gaussian interpolation as described in section \ref{gaussian_interpolation} and
+\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines.
 \end{enumerate}
@@
 \includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Comparison with baseline algorithms for various fractions of missing pixels.}
+\caption{Baseline algorithms for various fractions of missing pixels.}
 \label{baseline_algorithms}
@@
 
-We tested our algorithm and the two baselines on a set of natural images and further on a set of artificially created ones. 
-For evaluation, we tested these algorithms with different levels of missing pixels from 0 to 100\% in 5\% steps.
+We evaluated the baseline algorithms with the same sample images as optimization. With 60\% missing pixels, the algorithms had the following errors and standard deviations:
 
-Organize the results section based on the sequence of table and
-figures you include. Prepare the tables and figures as soon as all
-the data are analyzed and arrange them in the sequence that best
-presents your findings in a logical way. A good strategy is to note,
-on a draft of each table or figure, the one or two key results you
-want to address in the text portion of the results.
+\begin{table}[h]
+\centering
+\begin{tabular}{l|r|r}
+Algorithm & Error & Deviation \\
+\hline
+Linear interpolation & ? & ? \\
+Gaussian interpolation & ? & ? \\
+Matching pursuit & ? & ? \\
+\end{tabular}
+\end{table}
 
-When reporting computational or measurement results, always
-report the mean (average value) along with a measure of variablility
-(standard deviation(s) or standard error of the mean). ((( TODO !!! )))
+Since our algorithm did not perform equally well for all sample images, we print those for which it worked best and worst as figures \ref{best_image} and \ref{worst_image}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\caption{Best image with error ? and standard deviation ?.}
+\label{best_image}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\caption{Worst image with error ? and standard deviation ?.}
+\label{worst_image}
+\end{figure}
+
+The strength of our method is definitely in reconstructing edges, which can be seen in figure \ref{edge_reconstruction}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\caption{Original image with mask (left) and its reconstruciton (right).}
+\label{edge_reconstruction}
+\end{figure}
 
@@
 \section{Discussion}
-Our algorithm is resistant to high percentages of missing pixels, the average error per missing pixel rises only slowly until ((( TODO level))).
-
-Discuss the strengths and weaknesses of your approach, based on the results. Point out the implications of your novel idea on the application concerned. (((( TODO ))))
+Discuss the strengths and weaknesses of your approach, based on the results. Point out the implications of your novel idea on the application concerned.
 

### 2011-06-21 project/Paper/paper.tex
+ Added plots to paper

--- 2011-06-21 project/Paper/paper.tex
@@
 \begin{enumerate}
-\item Using the abovementioned validation set, we evaluate several thresholds for every patch and keep for each the best one: After calculating a fast Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
-Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to perform as few threshold evaluations as possible.
-For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and decreasing the step size (see figure \ref{threshold_optimization}).
+\item Using the above mentioned validation set, we evaluate several thresholds for every patch and keep for each the best one:
+After calculating a fast Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
+Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to
+perform as few threshold evaluations as possible.
+For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and
+decreasing the step size by half (see figure \ref{threshold_optimization}). While this heuristic does not always reach the minimum,
+it works quite well for our error function, that appears to be very roughly convex.
 For this optimization as well as for the whole iteration, we have both a number- and a convergence-based termination criterion.
@@
 \centering
-\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\includegraphics[width=\columnwidth]{../plots/search_strategy_1.pdf}
+\includegraphics[width=\columnwidth]{../plots/search_strategy_6.pdf}
 \caption{The threshold is optimized by iteratively taking the best one.}

### 2011-06-21 project/CreatePaperPlots.m
Revert to parfor loop, > 1 iteration Adjust CreatePaperPlots

@@ -11,7 +11,7 @@ load('params.mat');
 global debug_threshold_plot_number;
 fprintf('1) Creating Parameter evolution graphs')
 debug_threshold_plot_number = 10;
-EvaluateInpaintingParameterized(parameters, 0.6);
+EvaluateInpaintingParameterized(parameters, 0.6, 1);
 
 % Generate graphs
 stepsize = 2;
@@ -32,7 +32,7 @@ if(~exist('plots/error_A.mat', 'file'))
   % Use parallel computation for this, if available
   parfor k=0:no_steps
     % Our algorithm
-    [unused, e] = EvaluateInpaintingParameterized(parameters,stepsize*k/100);
+    [unused, e, unused2] = EvaluateInpaintingParameterized(parameters,stepsize*k/100, 3);
     error_algo(k+1) = e/(k*stepsize/100);% Normalized
   end
   save('plots/error_A.mat', 'error_algo');
@@ -46,7 +46,7 @@ if(~exist('plots/error_B1.mat', 'file'))
   cd('baseline1/')
   parfor k=0:no_steps
     % Baseline 1
-    error_base1(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
+    [error_base1(k+1), unused] = feval('EvaluateInpaintingParameterized', stepsize*k/100);
   end
   cd(main_path)
   save('plots/error_B1.mat', 'error_base1');

### 2011-06-21 project/baseline1/EvaluateInpaintingParameterized.m

@@ -1,46 +1,84 @@
 % Measure approximation error for several images.
 
-function [avgQErr] = EvaluateInpaintingParameterized(missing_pixels_fract)  
+function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract)  
 
   file_list = dir('..'); 
+  rep=3;
+  
   dir_length = length(dir('..'));
 
-  Errors = []; % mean squared errors for each image
-  
-  for i = 3 : dir_length  % running through the folder
-    file_name = file_list(i).name; % get current filename
+  Errors_final = [];
 
-    % Only keep the images in the loop
-    if (length(file_name) < 5)
-      continue;
-    elseif ( max(file_name(end-4:end) ~= '2.png'))
-      continue;
-    end
+  parfor r = 1 : rep
+     Errors = []; % mean squared errors for each image
+     for i = 3 : dir_length  % running through the folder
 
-    % Read image, convert to double precision and map to [0,1] interval
-    I = imread(strcat('../',file_name));
-    I = double(I) / 255;
+        file_name = file_list(i).name; % get current filename
 
-    % Generate mask at random, with 60% missing pixels
-    mask = ones(size(I));
-    numpixels = size(I,1)*size(I,2);
-    holes = randperm(numpixels);
-    holes = holes(1:floor(missing_pixels_fract*numpixels));
-    mask(holes) = 0;
+        % Only keep the images in the loop
+        if (length(file_name) < 5)
+          continue;
+        elseif ( max(file_name(end-4:end) ~= '2.png'))
+          continue;
+        end
+
+        % Read image, convert to double precision and map to [0,1] interval
+        I = imread(strcat('../',file_name));
+        I = double(I) / 255;
+
+        % Generate mask at random, with 60% missing pixels
+        mask = ones(size(I));
+        numpixels = size(I,1)*size(I,2);
+        holes = randperm(numpixels);
+        holes = holes(1:floor(missing_pixels_fract*numpixels));
+        mask(holes) = 0;
 
 
-    I_mask = I;
-    I_mask(~mask) = 0;
+        I_mask = I;
+        I_mask(~mask) = 0;
 
-    % Call the main inPainting function
-    I_rec = inPainting(I_mask, mask);
+        % Call the main inPainting function
+        I_rec = inPainting(I_mask, mask);
 
-    % Measure approximation error
[…]

### 2011-06-21 project/baseline1/inPainting.m

@@ -79,7 +79,6 @@ for x=1:w
       I(x, y) = values / s;
     end
   end
-  x
 end
     
 

### 2011-06-21 project/OptimizeInpainting.m
New Parameters added

@@ -43,18 +43,18 @@ function OptimizeInpainting(rounds)
     fprintf('=> cost %g\n\n', cost);
 
     % Optimize individual parameters
-    [final_parameters.gauss_size, old]             = gradientDescent(1, @getNextGaussSize, parameters, old, cost, missing_pixels, rounds_done);
-    [final_parameters.gauss_sigma, old]            = gradientDescent(2, @getNextGaussSigma, parameters, old, cost, missing_pixels, rounds_done);
-    [final_parameters.patch_size, old]             = gradientDescent(3, @getNextPatchSize, parameters, old, cost, missing_pixels, rounds_done);
-    [final_parameters.patch_frame_size, old]       = gradientDescent(4, @getNextFrameSize, parameters, old, cost, missing_pixels, rounds_done);
-    [final_parameters.td_abortbelow_stdev, old]    = gradientDescent(5, @getNextTDAbortBelowStdev, parameters, old, cost, missing_pixels, rounds_done);
-    [final_parameters.td_abortbelow_stepsize, old] = gradientDescent(6, @getNextTDAbortBelowStep, parameters, old, cost, missing_pixels, rounds_done);
-    [final_parameters.td_middle, old]              = gradientDescent(7, @getNextTDMiddle, parameters, old, cost, missing_pixels, rounds_done);
-    [final_parameters.validation, old]             = gradientDescent(8, @getNextValidation, parameters, old, cost, missing_pixels, rounds_done);
+    [final_parameters.gauss_size, old]             = gradientDescent(1, @getNextGaussSize, parameters, old, cost, missing_pixels, rounds_done, repetition_evaluation);
+    [final_parameters.gauss_sigma, old]            = gradientDescent(2, @getNextGaussSigma, parameters, old, cost, missing_pixels, rounds_done, repetition_evaluation);
+    [final_parameters.patch_size, old]             = gradientDescent(3, @getNextPatchSize, parameters, old, cost, missing_pixels, rounds_done, repetition_evaluation);
+    [final_parameters.patch_frame_size, old]       = gradientDescent(4, @getNextFrameSize, parameters, old, cost, missing_pixels, rounds_done, repetition_evaluation);
+    [final_parameters.td_abortbelow_stdev, old]    = gradientDescent(5, @getNextTDAbortBelowStdev, parameters, old, cost, missing_pixels, rounds_done, repetition_evaluation);
[…]

### 2011-06-21 project/CreatePaperPlots.m
Determine stdev for baseline 2

@@ -66,7 +66,7 @@ if(~exist('plots/error_B2.mat', 'file'))
   cd('baseline2/')
   parfor k=0:no_steps
     % Baseline 2
-    error_base2(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
+    [error_base2(k+1), unused] = feval('EvaluateInpaintingParameterized', stepsize*k/100);
   end
   cd(main_path)
   save('plots/error_B2.mat', 'error_base2');

### 2011-06-21 project/baseline1/EvaluateInpaintingParameterized.m

@@ -48,7 +48,7 @@ function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract
   
   avgQErr = mean(Errors_final);
   
-   % Determine mean and stdev per image
+  % Determine mean and stdev per image
   % Count images first
   imc = 0;
   filenames = {};

### 2011-06-21 project/baseline2/EvaluateInpaintingParameterized.m

@@ -1,6 +1,6 @@
 % Measure approximation error for several images.
 
-function [avgQErr] = EvaluateInpaintingParameterized(missing_pixels_fract)  
+function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract)  
 
   file_list = dir('..'); 
   rep=3;
@@ -8,11 +8,9 @@ function [avgQErr] = EvaluateInpaintingParameterized(missing_pixels_fract)
   dir_length = length(dir('..'));
 
   Errors_final = [];
-  LTimes_final = [];
 
   parfor r = 1 : rep
      Errors = []; % mean squared errors for each image
-     LTimes = [];
      for i = 3 : dir_length  % running through the folder
 
         file_name = file_list(i).name; % get current filename
@@ -40,18 +38,47 @@ function [avgQErr] = EvaluateInpaintingParameterized(missing_pixels_fract)
         I_mask(~mask) = 0;
 
         % Call the main inPainting function
-        tic;
         I_rec = inPainting(I_mask, mask);
-        LTimes = [LTimes toc];
 
         % Measure approximation error
         Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
      end
      Errors_final = [Errors_final Errors];
-     LTimes_final = [LTimes_final LTimes];
   end
   
   avgQErr = mean(Errors_final);
+  
+  % Determine mean and stdev per image
+  % Count images first
+  imc = 0;
+  filenames = {};
+  for i = 3 : dir_length
+    file_name = file_list(i).name; % get current filename
+    % Only keep the images in the loop
+      if (length(file_name) < 5)
+       continue;
+      elseif ( max(file_name(end-4:end) ~= '2.png'))
+       continue;
+      end
+      %Count and save filenames for later
+      imc = imc+1;
+      filenames{imc} = {file_name};
+  end
+  
+  for i=1:imc
+    stdev{i,2} = filenames{i};
+  end
+  
+  if rep > 1
+    for i=1:imc
+      stdev{i,1} = std(Errors_final(i:imc:end));
+    end
+  else
+    % 1 Iteration - no deviation
+    for i=1:imc
+      stdev{i,1} = 0;
+    end
+  end
 
 
 end

### 2011-06-21 project/CreatePaperPlots.m
Added stdev to baseline 3

@@ -80,7 +80,7 @@ if(~exist('plots/error_B3.mat', 'file'))
   cd('baseline3/')
   parfor k=0:no_steps
     % Baseline 3
-    error_base3(k+1) = feval('EvaluateInpaintingParameterized', stepsize*k/100);
+    [error_base3(k+1), unused] = feval('EvaluateInpaintingParameterized', stepsize*k/100);
   end
   cd(main_path)
   save('plots/error_B3.mat', 'error_base3');

### 2011-06-21 project/CreatePaperPlots.m
Adapt return value

@@ -6,18 +6,36 @@
 % Requires parameters saved in a parameters.mat file
 
 load('params.mat');
+main_path = pwd;
 
 % Generate strategy plots
 global debug_threshold_plot_number;
-fprintf('1) Creating Parameter evolution graphs')
+fprintf('1) Creating Parameter evolution graphs\n')
 debug_threshold_plot_number = 10;
 
-EvaluateInpaintingParameterized(parameters, 0.6, 1);
+[unused, error_a, stdev_a] = EvaluateInpaintingParameterized(parameters, 0.6, 1);
 
 for i = 1:10
   f = open(sprintf('plots/search_strategy_%g.fig', i));
   plotpdftex(f, sprintf('plots/search_strategy_%g', i));
 end
+
+fprintf('2) Evaluating performance at 60%% missing pixels\n');
+cd('baseline1/')
+    % Baseline 1
+    [error_b1, stdev_b1] = feval('EvaluateInpaintingParameterized', 0.6);
+cd(main_path)
+
+cd('baseline2/')
+    % Baseline 2
+    [error_b2, stdev_b2] = feval('EvaluateInpaintingParameterized', 0.6);
+cd(main_path)
+
+cd('baseline3/')
+    % Baseline 3
+    [error_b3, stdev_b3] = feval('EvaluateInpaintingParameterized', 0.6);
+cd(main_path)
+
 
 % Generate graphs
 stepsize = 2;
@@ -29,10 +47,9 @@ error_base1 = zeros(no_steps, 1);
 error_base2 = zeros(no_steps, 1);
 error_base3 = zeros(no_steps, 1);
 
-main_path = pwd;
 
-fprintf('2) Calculating error vs. missing pixels\n')
-fprintf('2A) New algorithm\n')
+fprintf('3) Calculating error vs. missing pixels\n')
+fprintf('3A) New algorithm\n')
 
 if(~exist('plots/error_A.mat', 'file'))
   % Use parallel computation for this, if available
@@ -47,7 +64,7 @@ else
   load('plots/error_A.mat');
 end
 
-fprintf('2B) Baseline 1\n')
+fprintf('3B) Baseline 1\n')
 if(~exist('plots/error_B1.mat', 'file'))    
   cd('baseline1/')
   parfor k=0:no_steps
@@ -61,7 +78,7 @@ else
   load('plots/error_B1.mat');
 end
 
[…]

### 2011-06-21 project/CreatePaperPlots.m
+ Better createplots

@@ -13,29 +13,24 @@ global debug_threshold_plot_number;
 fprintf('1) Creating Parameter evolution graphs\n')
 debug_threshold_plot_number = 10;
 
-[unused, error_a, stdev_a] = EvaluateInpaintingParameterized(parameters, 0.6, 1);
+%[unused, error_a, stdev_a] = EvaluateInpaintingParameterized(parameters, 0.6, 1);
 
-for i = 1:10
-  f = open(sprintf('plots/search_strategy_%g.fig', i));
-  plotpdftex(f, sprintf('plots/search_strategy_%g', i));
-end
+%for i = 1:10
+%  f = open(sprintf('plots/search_strategy_%g.fig', i));
+%  plotpdftex(f, sprintf('plots/search_strategy_%g', i));
+%end
+
+addpath('baseline1', 'baseline2', 'baseline3');
 
 fprintf('2) Evaluating performance at 60%% missing pixels\n');
-cd('baseline1/')
-    % Baseline 1
-    [error_b1, stdev_b1] = feval('EvaluateInpaintingParameterized', 0.6);
-cd(main_path)
-
-cd('baseline2/')
-    % Baseline 2
-    [error_b2, stdev_b2] = feval('EvaluateInpaintingParameterized', 0.6);
-cd(main_path)
-
-cd('baseline3/')
-    % Baseline 3
-    [error_b3, stdev_b3] = feval('EvaluateInpaintingParameterized', 0.6);
-cd(main_path)
-
+fprintf('Working in inPainting ...\n')
+[avgQErr, stdev, stdev_runs, stdev_diff] = EvaluateInpaintingParameterizedStatistics(0.6, 10, @inPainting)
+fprintf('Working in baseline1 ...\n')
+[avgQErr, stdev, stdev_runs, stdev_diff] = EvaluateInpaintingParameterizedStatistics(0.6, 10, @baseline1)
+fprintf('Working in baseline2 ...\n')
+[avgQErr, stdev, stdev_runs, stdev_diff] = EvaluateInpaintingParameterizedStatistics(0.6, 10, @baseline2)
+fprintf('Working in baseline3 ...\n')
+[avgQErr, stdev, stdev_runs, stdev_diff] = EvaluateInpaintingParameterizedStatistics(0.6, 10, @baseline3)
 
 % Generate graphs
 stepsize = 2;
@@ -55,7 +50,7 @@ if(~exist('plots/error_A.mat', 'file'))
   % Use parallel computation for this, if available
[…]

### 2011-06-21 project/EvaluateInpaintingParameterizedStatistics.m

@@ -1,13 +1,13 @@
 % Measure approximation error for several images.
 
-function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract)  
-
-  file_list = dir('..'); 
-  rep=3;
+function [avgQErr, stdev, stdev_runs, stdev_diff] = EvaluateInpaintingParameterizedStatistics(missing_pixels_fract, rep, inPainting)
+  file_list = dir('.'); 
   
-  dir_length = length(dir('..'));
+  dir_length = length(dir('.'));
 
   Errors_final = [];
+  Error_means = [];
+  stdev_diffs = [];
 
   parfor r = 1 : rep
      Errors = []; % mean squared errors for each image
@@ -23,7 +23,7 @@ function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract
         end
 
         % Read image, convert to double precision and map to [0,1] interval
-        I = imread(strcat('../',file_name));
+        I = imread(strcat('./',file_name));
         I = double(I) / 255;
 
         % Generate mask at random, with 60% missing pixels
@@ -44,6 +44,8 @@ function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract
         Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
      end
      Errors_final = [Errors_final Errors];
+     Error_means = [Error_means mean(Errors)];
+     stdev_diffs = [stdev_diffs std(Errors)];
   end
   
   avgQErr = mean(Errors_final);
@@ -73,12 +75,13 @@ function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract
     for i=1:imc
       stdev{i,1} = std(Errors_final(i:imc:end));
     end
+    stdev_runs = std(Error_means);
   else
     % 1 Iteration - no deviation
     for i=1:imc
       stdev{i,1} = 0;
     end
+    stdev_runs = 0;
   end
-
-
+  stdev_diff = mean(stdev_diffs);
 end

### 2011-06-21 project/baseline1/baseline1.m

@@ -1,5 +1,4 @@
 function I_rec = inPainting(I, mask)
-
 % Baseline 1: Perform linear inpainting of the image 
 
 % INPUT

### 2011-06-21 project/baseline2/EvaluateInpaintingParameterized.m

@@ -1,84 +1,0 @@
-% Measure approximation error for several images.
-
-function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract)  
-
-  file_list = dir('..'); 
-  rep=3;
-  
-  dir_length = length(dir('..'));
-
-  Errors_final = [];
-
-  parfor r = 1 : rep
-     Errors = []; % mean squared errors for each image
-     for i = 3 : dir_length  % running through the folder
-
-        file_name = file_list(i).name; % get current filename
-
-        % Only keep the images in the loop
-        if (length(file_name) < 5)
-          continue;
-        elseif ( max(file_name(end-4:end) ~= '2.png'))
-          continue;
-        end
-
-        % Read image, convert to double precision and map to [0,1] interval
-        I = imread(strcat('../',file_name));
-        I = double(I) / 255;
-
-        % Generate mask at random, with 60% missing pixels
-        mask = ones(size(I));
-        numpixels = size(I,1)*size(I,2);
-        holes = randperm(numpixels);
-        holes = holes(1:floor(missing_pixels_fract*numpixels));
-        mask(holes) = 0;
-
-
-        I_mask = I;
-        I_mask(~mask) = 0;
-
-        % Call the main inPainting function
-        I_rec = inPainting(I_mask, mask);
-
-        % Measure approximation error
-        Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
-     end
-     Errors_final = [Errors_final Errors];
-  end
-  
-  avgQErr = mean(Errors_final);
-  
-  % Determine mean and stdev per image
-  % Count images first
-  imc = 0;
-  filenames = {};
-  for i = 3 : dir_length
-    file_name = file_list(i).name; % get current filename
-    % Only keep the images in the loop
-      if (length(file_name) < 5)
-       continue;
-      elseif ( max(file_name(end-4:end) ~= '2.png'))
-       continue;
-      end
-      %Count and save filenames for later
-      imc = imc+1;
-      filenames{imc} = {file_name};
-  end
-  
-  for i=1:imc
-    stdev{i,2} = filenames{i};
-  end
-  
-  if rep > 1
-    for i=1:imc
-      stdev{i,1} = std(Errors_final(i:imc:end));
-    end
-  else
-    % 1 Iteration - no deviation
-    for i=1:imc
-      stdev{i,1} = 0;
-    end
-  end
-
-
-end

### 2011-06-21 project/baseline3/EvaluateInpaintingParameterized.m

@@ -1,85 +1,0 @@
-% Measure approximation error for several images.
-
-function [avgQErr, stdev] = EvaluateInpaintingParameterized(missing_pixels_fract)  
-
-  file_list = dir('..'); 
-  rep=3;
-  
-  dir_length = length(dir('..'));
-
-  Errors_final = [];
-
-  parfor r = 1 : rep
-     Errors = []; % mean squared errors for each image
-     for i = 3 : dir_length  % running through the folder
-
-        file_name = file_list(i).name; % get current filename
-
-        % Only keep the images in the loop
-        if (length(file_name) < 5)
-          continue;
-        elseif ( max(file_name(end-4:end) ~= '2.png'))
-          continue;
-        end
-
-        % Read image, convert to double precision and map to [0,1] interval
-        I = imread(strcat('../',file_name));
-        I = double(I) / 255;
-
-        % Generate mask at random, with 60% missing pixels
-        mask = ones(size(I));
-        numpixels = size(I,1)*size(I,2);
-        holes = randperm(numpixels);
-        holes = holes(1:floor(missing_pixels_fract*numpixels));
-        mask(holes) = 0;
-
-
-        I_mask = I;
-        I_mask(~mask) = 0;
-
-        % Call the main inPainting function
-        I_rec = inPainting(I_mask, mask);
-
-        % Measure approximation error
-        Errors = [Errors mean(mean(mean( ((I - I_rec) ).^2)))];
-     end
-     Errors_final = [Errors_final Errors];
-  end
-  
-  avgQErr = mean(Errors_final);
-  
-  
-  % Determine mean and stdev per image
-  % Count images first
-  imc = 0;
-  filenames = {};
-  for i = 3 : dir_length
-    file_name = file_list(i).name; % get current filename
-    % Only keep the images in the loop
-      if (length(file_name) < 5)
-       continue;
-      elseif ( max(file_name(end-4:end) ~= '2.png'))
-       continue;
-      end
-      %Count and save filenames for later
-      imc = imc+1;
-      filenames{imc} = {file_name};
-  end
-  
-  for i=1:imc
-    stdev{i,2} = filenames{i};
-  end
-  
-  if rep > 1
-    for i=1:imc
-      stdev{i,1} = std(Errors_final(i:imc:end));
-    end
-  else
-    % 1 Iteration - no deviation
-    for i=1:imc
-      stdev{i,1} = 0;
-    end
-  end
-
-
-end

### 2011-06-21 project/Paper/paper.tex

@@
 \begin{document}
-\title{Inpainting with controlled dimensionality reduction in the Fourier domain}
+\title{Inpainting with controlled frequency suppression in the Fourier domain}
 
@@
 \begin{abstract}
-The abstract should really be written last, along with the title of
-the paper. The four points that should be covered:
-\begin{enumerate}
-\item State the problem.
-\item Say why it is an interesting problem.
-\item Say what your solution achieves. (CLAIM!)
-\item Say what follows from your solution.
-\end{enumerate}
+This paper presents a new approach to the problem of inpainting, which is about recovering (deliberately) lost image information.
+In comparison with the submissions of our peers, we have designed a very efficient algorithm that does not fall behind other solutions in terms of accuracy.
+By suppressing weak frequencies in the Fourier domain, we attempt to recover the intrinsic signal of the image.
+Since we choose the optimal threshold with a random validation set, our method performs particularly well for the reconstruction of edges.
 \end{abstract}
@@
 \begin{enumerate}
-\item Using the above mentioned validation set, we evaluate several thresholds for every patch and keep for each the best one:
-After calculating a fast Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
-Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to
-perform as few threshold evaluations as possible.
-For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and
-decreasing the step size by half (see figure \ref{threshold_optimization}). While this heuristic does not always reach the minimum,
-it works quite well for our error function, that appears to be very roughly convex.
+\item Using the above mentioned validation set, we evaluate several thresholds for every patch and keep for each the best one: After calculating a fast Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
+Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to perform as few threshold evaluations as possible.
+For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and halve the step size (see figure \ref{threshold_optimization}).
+While this heuristic does not always reach the minimum, it works quite well for our error function, which is often roughly convex.
 For this optimization as well as for the whole iteration, we have both a number- and a convergence-based termination criterion.
@@
 \includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Due to the momentum term, the cost function does not strictly decrease.}
+\caption{Due to the momentum, the cost function does not strictly decrease.}
 \label{gradient_descent_figure}
@@
 \section{Summary}
-Summarise your contributions in light of the new results.
+Making use of our domain knowledge about images, we have implemented an algorithm that performs well in terms of both accuracy and performance.
+The remarkable speed is achieved by relying on MATLAB's native functions for the most computationally intensive tasks.
+In our opinion, we have devised many nice methods: Gaussian interpolation, threshold optimization with validation set and frequency suppression in the Fourier domain to name just the most important ones.
+The key feature of our algorithm is certainly its ability to reconstruct the edges in an image.
 
@@
 \end{document}
-
-% Plots:
-% - threshold optimization => algorithm
-% - Framing artifacts (with and w/o frame) => framing
-% - Parameter optimization (convergence) => gradient descent
-% 
-% - Missing pixels vs. error (baselines)
-% - stdev over n runs of EvaluateInpainting with rep = 1
-% 
-% - image with best error (and stdev)
-% - image with worst error (and stdev)
-% 
-% - Bar to illustrate edge reconstruction (patch size)

### 2011-06-23 project/Paper/paper.tex
+ Reviewed paper up to global optimisation

@@
 \begin{abstract}
-This paper presents a new approach to the problem of inpainting, which is about recovering (deliberately) lost image information.
+This paper presents a new approach to the problem of inpainting, that is recovering of lost image information.
 In comparison with the submissions of our peers, we have designed a very efficient algorithm that does not fall behind other solutions in terms of accuracy.
 By suppressing weak frequencies in the Fourier domain, we attempt to recover the intrinsic signal of the image.
-Since we choose the optimal threshold with a random validation set, our method performs particularly well for the reconstruction of edges.
+Since we choose the optimal threshold using a random validation set, our method performs particularly well for the reconstruction of edges.
 \end{abstract}
@@
 Inpainting is about reconstructing the missing spots of an image by guessing the correct values.
-The quality of such a restoration method can be evaluated by masking an image and comparing the result with the original.\\
+The quality of such a restoration method can be evaluated by masking an image and comparing the reconstruction with the original.\\
 Our approach is based on the assumption that (natural) images have some underlying characteristic that can be distinguished from noise.
-The computed pixels can be viewed as a disturbed version of the genuine ones, hence the goal is to get rid of this noise.
-We do so by analyzing image patches in the Fourier space, removing the weak frequencies that (hopefully) capture the noise and thereby recovering the intrinsic signal.\\
+We initialize our reconstructed image with what we call Gaussian interpolation: Using MATLAB's filtering function, we compute a weighted sum over all the neighboring pixels that are known.
+This reconstruction introduces noise compared to the original image.
+The hope is that this noise is weak enough compared to the real signal, that we want to get rid of it.
+We do so by dividing the image into patches and analyzing these in the Fourier space, removing the weak frequencies that (hopefully) capture this noise and thereby recovering the intrinsic signal.\\
 We repeat this procedure with the newly determined values until the error of a randomly chosen validation set converges or the maximum number of iterations is reached.
-For a faster start, we initialize our reconstructed image with what we call Gaussian interpolation: Using MATLAB's filtering function, we compute a weighted sum over all the neighboring pixels that are known.
-In order to determine the best threshold for suppressing weaker frequencies, we randomly remove even more pixels from the image and use them as a validation set (for which we know the correct values).
+This validation set consists of some of known pixels given to the algorithm, that are removed before starting the reconstruction.
+In order to determine the  threshold for suppressing weaker frequencies, this validation set (for which we know the correct values) can be used as an error measure.
 
@@
 \subsection{Related work}
-Without actually knowing the broad field of research about inpainting (also known as image interpolation), we could not find any papers covering similar concepts as ours.
+Without knowing in depth the broad field of research about inpainting (also known as image interpolation), we were not able to find papers covering similar concepts as ours.
 Most methods are based on partial differential equations or some form of texture synthesis.
@@
 \tikzstyle{block} = [rectangle, draw, fill=blue!20, 
-    text width=7em, text centered, rounded corners, minimum height=4em]
+    text width=7em, text centered, rounded corners, minimum height=3em]
 
@@
 \begin{figure}
-\begin{tikzpicture}[node distance = 2cm, auto]
+\begin{tikzpicture}[node distance = 1.5cm, auto]
   % Place nodes
@@
   \node [block, below of=node3] (node4) {Gaussian interpolation};
-  \node [decision, below of=node4] (node5) {Iterative?};
+  \node [decision, below of=node4, node distance = 1.8cm] (node5) {Iterative?};
   \node [block, right of=node5, node distance = 3.1cm, fill=green!20] (node6) {Find best threshold};
-  \node [block, below of=node6, fill=green!20] (node7) {Reset known values except validation set};
+  \node [block, below of=node6, fill=green!20, node distance = 1.8cm] (node7) {Reset known values except validation set};
 
@@
 The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section \ref{gradient_descent}).
-\item Since our reduction in the Fourier space works best with framed patches (see figure \ref{framing_artifacts}), we frame the complete image by mirroring its margin.
-(The size of the frame is another optimizable global parameter.)
+\item Our reduction in the Fourier space works best with frames, meaning that we use large overlapping patches for the fourier transform, but only use the inner part for the reconstruction (see figure \ref{framing_artifacts}), we frame the complete image by mirroring its margin.
+The size of the frame is another optimizable global parameter.
 
@@
 \includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Artifacts occur w/o frames (left) that vanish with frames (right).}
+\caption{Artifacts occur w/o frames (left) that vanish with frames (right).} % TODO
 \label{framing_artifacts}
@@
 \item For obtaining an initial reconstruction of the image, we compute a weighted sum over all the neighboring pixels that are known.
-By setting all unknown pixels in the image to zero and encoding known values with 1s in the mask, elementwise division of the filtered image by the similarly filtered mask results in a normalized sum weighted by the filtering kernel.
-(This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.)
+By setting all unknown pixels in the image to zero and encoding known values with 1s in the mask, elementwise division of the filtered image by the similarly filtered mask results in a normalized sum weighted by the filtering kernel.\\
+This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.
 \label{gaussian_interpolation}
@@
 \begin{enumerate}
-\item Using the above mentioned validation set, we evaluate several thresholds for every patch and keep for each the best one: After calculating a fast Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
+\item Using the above mentioned validation set, we evaluate several thresholds for every patch and keep for each the best one: After calculating a Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
 Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to perform as few threshold evaluations as possible.
@@
 \centering
-\includegraphics[width=\columnwidth]{../plots/search_strategy_1.pdf}
-\includegraphics[width=\columnwidth]{../plots/search_strategy_6.pdf}
-\caption{The threshold is optimized by iteratively taking the best one.}
+\includegraphics[width=0.8\columnwidth]{../plots/search_strategy_6.pdf}
+\includegraphics[width=0.8\columnwidth]{../plots/search_strategy_2.pdf}
+\caption{The threshold is optimized by iteratively.}
 \label{threshold_optimization}

### 2011-06-23 project/CreatePaperPlots.m

@@ -5,7 +5,7 @@
 % Create Error vs. MissingPixels Plot
 % Requires parameters saved in a parameters.mat file
 
-load('params.mat');
+load('start_params.mat');
 main_path = pwd;
 
 % Generate strategy plots
@@ -20,6 +20,7 @@ for i = 1:10
   plotpdftex(f, sprintf('plots/search_strategy_%g', i));
 end
 
+load('params.mat');
 addpath('baseline1', 'baseline2', 'baseline3');
 
 fprintf('2) Evaluating performance at 60%% missing pixels\n');

### 2011-06-23 project/Paper/paper.tex
! Finished paper review

@@
 \usepackage{graphicx}
+\usepackage{subfig}
 \usepackage[pdftex]{hyperref}
@@
 \label{gradient_descent}
-We optimize all parameters at once using a method called gradient descent.
+We optimize all parameters at once using a gradient descent method.
 Since our cost function is based on measurements, we cannot derive it with respect to individual parameters and have to determine the gradient with finite differences.
-Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term that should help us to overcome local minima (see figure \ref{gradient_descent_figure}).
+Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term of 0.8 that should help us to overcome local minima and improve convergence speed.
+Figure \ref{gradient_descent_cost} shows the development of the cost for 100 optimisation steps.
 
@@
 \centering
-\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Due to the momentum, the cost function does not strictly decrease.}
-\label{gradient_descent_figure}
+\includegraphics[width=0.8\columnwidth]{../plots/cost_plot.pdf}
+\caption{Due to the momentum term, the cost function does not strictly decrease when optimising.}
+\label{gradient_descent_cost}
 \end{figure}
@@
 \centering
+{\renewcommand{\arraystretch}{1.5}
+\renewcommand{\tabcolsep}{0.2cm}
 \begin{tabular}{|l|p{6cm}|p{6cm}|}
@@
 \hline
-$\begin{cases}
-E & \text{if }T<60\\
-E+\exp(T/60 - 1) - 1 & \text{otherwise}
-\end{cases}$ & Continuous function that optimizes for error only when the time is below a certain value but penalizes long times. & Seems to work quite well and therefore used by us \\
+$ \begin{cases}
+E & \text{if }T<T_0\\
+E+\exp(T/T_0 - 1) - 1 & \text{otherwise}
+\end{cases} $ & Continuous function, optimizes for error when the time is below a target time $T_0$, penalizes long times. & Seems to work quite well, used by us \\
 \hline
-\end{tabular}
+\end{tabular}}
 \caption{Various cost functions used in the optimization process.}
@@
 \subsubsection{Sample data}
-We used the three images provided for the last exercise and four additional ones (consisting of both natural and artificial images) to evaluate the accuracy and performance of the chosen parameters.
+We used the three images provided for the last exercise and four additional ones (consisting of both natural and artificial images, see figure \ref{new_images}) to evaluate the accuracy and performance of the chosen parameters.
 Instead of using the supplied mask, however, we generate one randomly for every evaluation in order to prevent overfitting to a particular mask.
+
+\begin{figure*}
+  \centering
+  \subfloat{\includegraphics[width=0.22\textwidth]{../daisiesFromBing512x512}}~
+  \subfloat{\includegraphics[width=0.22\textwidth]{../hardedges_512x512}}~
+  \subfloat{\includegraphics[width=0.22\textwidth]{../penguinsFromBing512x512}}~
+  \subfloat{\includegraphics[width=0.22\textwidth]{../velos512x512}}~
+  \caption{The four additional images used during the optimisation.}
+  \label{new_images}
+\end{figure*}
 
@@
 \item Some parameters have a clear speed vs. accuracy tradeoff, like the stopping conditions of the threshold optimization, which depend on our cost function and could not be foreseen.
-\item Some parameters assumed totally different values than anticipated: The patch size became much larger and the validation set decreased significantly.
+\item Some parameters assumed totally different values than anticipated: The patch size became much larger and the validation set decreased significantly, as can be seen in figure \ref{gradient_descent_parameters}.
 \end{itemize}
+
+\begin{figure}
+  \centering
+  \subfloat{\label{fig:gull}\includegraphics[width=0.25\textwidth]{../plots/td_middle_plot.pdf}}                
+  \subfloat{\label{fig:tiger}\includegraphics[width=0.25\textwidth]{../plots/validation_plot.pdf}}
+  \caption{Two parameters that seem to converge quite well during the optimisation.}
+  \label{gradient_descent_parameters}
+\end{figure}
 
@@
 \item Gaussian interpolation as described in section \ref{gaussian_interpolation} and
-\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines.
+\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines. TODO this reference is wrong!
 \end{enumerate}
@@
 
-\paragraph{Note} All plots in this paper were created with MATLAB and can be reproduced using the script \emph{CreatePaperPlots.m} of our submission.
+\paragraph{Note} All plots in this paper were created with MATLAB and were produced using the script \emph{CreatePaperPlots.m} of our submission.
+However, due to the randomness of the algorithm and varying speed measurements for the cost function, results can be different for each run.
 
 \section{Discussion}
-Discuss the strengths and weaknesses of your approach, based on the results. Point out the implications of your novel idea on the application concerned.
+Discuss the strengths and weaknesses of your approach, based on the results. Point out the implications of your novel idea on the application concerned. TODO
 

### 2011-06-23 project/OptimizeInpainting.m

@@ -7,9 +7,6 @@ function OptimizeInpainting(rounds)
   missing_pixels = 0.6;
   repetition_evaluation = 2;
 
-  % Set random seet to get reproducable results
-  rand('seed', 12345);
-
   % Load parameters and best_cost from (binary) file
   load('params.mat');
 

### 2011-06-23 project/Paper/paper.tex
Paper: Removed and scaled images, minor corrections.

@@
 Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term of 0.8 that should help us to overcome local minima and improve convergence speed.
-Figure \ref{gradient_descent_cost} shows the development of the cost for 100 optimisation steps.
+Figure \ref{gradient_descent_cost} shows the development of the cost for 100 optimization steps.
 
@@
 \includegraphics[width=0.8\columnwidth]{../plots/cost_plot.pdf}
-\caption{Due to the momentum term, the cost function does not strictly decrease when optimising.}
+\caption{Due to the momentum term, the cost function does not strictly decrease when optimizing.}
 \label{gradient_descent_cost}
@@
   \centering
-  \subfloat{\includegraphics[width=0.22\textwidth]{../daisiesFromBing512x512}}~
-  \subfloat{\includegraphics[width=0.22\textwidth]{../hardedges_512x512}}~
-  \subfloat{\includegraphics[width=0.22\textwidth]{../penguinsFromBing512x512}}~
-  \subfloat{\includegraphics[width=0.22\textwidth]{../velos512x512}}~
-  \caption{The four additional images used during the optimisation.}
+  \subfloat{\includegraphics[width=0.24\textwidth]{../daisiesFromBing512x512}}~
+  \subfloat{\includegraphics[width=0.24\textwidth]{../penguinsFromBing512x512}}~
+  \subfloat{\includegraphics[width=0.24\textwidth]{../velos512x512}}~
+  \subfloat{\includegraphics[width=0.24\textwidth]{../hardedges_512x512}}~
+  \caption{The four additional images used during optimization.}
   \label{new_images}
@@
   \subfloat{\label{fig:tiger}\includegraphics[width=0.25\textwidth]{../plots/validation_plot.pdf}}
-  \caption{Two parameters that seem to converge quite well during the optimisation.}
+  \caption{Two parameters that seem to converge quite well during the optimization.}
   \label{gradient_descent_parameters}
@@
 \item Gaussian interpolation as described in section \ref{gaussian_interpolation} and
-\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines. TODO this reference is wrong!
+\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines.
 \end{enumerate}
@@
 
-Since our algorithm did not perform equally well for all sample images, we print those for which it worked best and worst as figures \ref{best_image} and \ref{worst_image}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Best image with error ? and standard deviation ?.}
-\label{best_image}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Worst image with error ? and standard deviation ?.}
-\label{worst_image}
-\end{figure}
+Our algorithm did not perform equally well for all sample images: It worked best for ? (error ? and standard deviation ?) and worst for ? (error ? and standard deviation ?).
 
@@
 \includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Original image with mask (left) and its reconstruciton (right).}
+\caption{Original image with mask (left) and its reconstruction (right).}
 \label{edge_reconstruction}

### 2011-06-23 project/CreatePaperPlots.m
Ignore step at 0% missing pixels - normalized error would be NaN

@@ -44,8 +44,9 @@ save('plots/errsAndStdevB3.mat', 'avgQErrB3', 'stdevB3', 'stdev_runsB3', 'stdev_
 
 % Generate graphs
 stepsize = 2;
-% Ignore the 100% - the runtime will be inacceptable
-no_steps = floor(100/stepsize)-1;
+% Ignore the 100% - the runtime will be inacceptable.
+% And ignore the 0% - normalizing would give NAN anyway
+no_steps = floor(100/stepsize)-2;
 
 % Init arrays
 error_algo = zeros(no_steps, 1);
@@ -61,8 +62,8 @@ if(~exist('plots/error_A.mat', 'file'))
   % Use parallel computation for this, if available
   parfor k=0:no_steps
     % Our algorithm
-    e = EvaluateInpaintingParameterizedStatistics(stepsize*k/100, 3, @inPainting);
-    error_algo(k+1) = e/(k*stepsize/100);% Normalized
+    e = EvaluateInpaintingParameterizedStatistics((stepsize+k*stepsize)/100, 3, @inPainting);
+    error_algo(k+1) = e/((stepsize+k*stepsize)/100);% Normalized
   end
   save('plots/error_A.mat', 'error_algo');
 else
@@ -74,8 +75,8 @@ fprintf('3B) Baseline 1\n')
 if(~exist('plots/error_B1.mat', 'file'))    
   parfor k=0:no_steps
     % Baseline 1
-    e = EvaluateInpaintingParameterizedStatistics(stepsize*k/100, 3, @baseline1);
-    error_base1(k+1) = e/(k*stepsize/100);% Normalized
+    e = EvaluateInpaintingParameterizedStatistics((stepsize+k*stepsize)/100, 3, @baseline1);
+    error_base1(k+1) = e/((stepsize+k*stepsize)/100);% Normalized
   end
   save('plots/error_B1.mat', 'error_base1');
 else
@@ -87,8 +88,8 @@ fprintf('3C) Baseline 2\n')
 if(~exist('plots/error_B2.mat', 'file'))    
   parfor k=0:no_steps
     % Baseline 2
-    e = EvaluateInpaintingParameterizedStatistics(stepsize*k/100, 3, @baseline2);
-    error_base2(k+1) = e/(k*stepsize/100);% Normalized
+    e = EvaluateInpaintingParameterizedStatistics((stepsize+k*stepsize)/100, 3, @baseline2);
[…]

### 2011-06-24 project/Paper/paper.tex
Added plot of missing pixel error
Added comparison view Added images to the 'other_images' folder
Paper: Smaller captions.
Paper: Minor changes.

@@
 
-%\usepackage{times}
-%\usepackage{balance}
+\usepackage[font=footnotesize]{caption}
+
 \usepackage{amsmath}
@@
 \usepackage[utf8]{inputenc}
-
 \usepackage{tikz}
 \usetikzlibrary{shapes,arrows}
+
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
 
@@
 \includegraphics[width=0.8\columnwidth]{../plots/search_strategy_2.pdf}
-\caption{The threshold is optimized by iteratively.}
+\caption{The threshold is iteratively optimized.}
 \label{threshold_optimization}
@@
 \centering
-\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
+\includegraphics[width=\columnwidth]{images/missingpixelsVsError.png}
 \caption{Baseline algorithms for various fractions of missing pixels.}
@@
 \centering
-\begin{tabular}{l|r|r}
+\begin{tabular}{l|c|c}
 Algorithm & Error & Deviation \\
 \hline
-Linear interpolation & ? & ? \\
-Gaussian interpolation & ? & ? \\
-Matching pursuit & ? & ? \\
+New algorithm & 5.73\e{-4} & 6.32\e{-6} \\
+Linear interpolation & 8.08\e{-4} & 7.10\e{-6} \\
+Gaussian interpolation & 5.74\e{-3} & 1.10\e{-5} \\
+Matching pursuit & 1.72\e{-3} & 1.71\e{-5} \\
 \end{tabular}
@@
 
-Our algorithm did not perform equally well for all sample images: It worked best for ? (error ? and standard deviation ?) and worst for ? (error ? and standard deviation ?).
+Our algorithm did not perform equally well for all sample images: It worked best for the spiral (error 1.25\e{-4} and standard deviation 1.92\e{-6}) and worst for the bicycles (error 3.05\e{-3} and standard deviation 4.09\e{-5}).
 
@@
 \centering
-\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Original image with mask (left) and its reconstruction (right).}
+\includegraphics[width=\columnwidth]{images/paper_reconstruction_bars32_mask8.png}
+\caption{Original image with mask (left, mask in blue) and its reconstruction (right). Enlarged for display in this paper.}
 \label{edge_reconstruction}

### 2011-06-24 project/Paper/paper.tex
Entered values of final submission

--- 2011-06-23 project/Paper/paper.tex
@@
 \hline
-gauss\_size & Size of the Gaussian kernel in pixels & 10 & ? \\
-gauss\_sigma & Sigma of the Gaussian kernel & 1 & ? \\
-patch\_size & Patch size as exponent to the basis 2 & 4 & ? \\
-patch\_frame\_size & Frame size in pixels & 8 & ? \\
-td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds as negative exponent to basis 10& 3 & ? \\
-td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds as negative exponent to basis 10& 4 & ? \\
-td\_middle & Initial threshold & 10 & ? \\
-validation & Fraction of pixels in the validation set & 0.4 & ? \\
-max\_iterations & Maximum number of iterations & 5 & ? \\
-abortbelow\_change & Minimal relative change to perform another iteration & 0.1 & ? \\
+gauss\_size & Size of the Gaussian kernel in pixels & 10 & 19 \\
+gauss\_sigma & Sigma of the Gaussian kernel & 1 & 0.6523 \\
+patch\_size & Patch size as exponent to the basis 2 & 4 & 9 \\
+patch\_frame\_size & Frame size in pixels & 8 & 0 \\
+td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds as negative exponent to basis 10& 3 & 1.9395 \\
+td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds as negative exponent to basis 10& 4 & 3.5226 \\
+td\_middle & Initial threshold & 10 & 16.1555 \\
+validation & Fraction of pixels in the validation set & 0.4 & 0.0151 \\
+max\_iterations & Maximum number of iterations & 5 & 7 \\
+abortbelow\_change & Minimal relative change to perform another iteration & 0.1 & 0.0465 \\
 \hline
@@
 \end{center}
-\caption{Parameters with their initial and optimized values.}
+\caption{Parameters with their initial and optimized values. Optimized values are rounded, if necessary (e.g. to whole pixels)}
 \label{parameters}

### 2011-06-24 project/Paper/paper.tex
Paper: Removed artifacts image.

--- 2011-06-24 project/Paper/paper.tex
@@
 The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section \ref{gradient_descent}).
-\item Our reduction in the Fourier space works best with frames, meaning that we use large overlapping patches for the fourier transform, but only use the inner part for the reconstruction (see figure \ref{framing_artifacts}), we frame the complete image by mirroring its margin.
+\item Our reduction in the Fourier space works best with frames, meaning that we use large overlapping patches for the fourier transform, but only use the inner part for the reconstruction, we frame the complete image by mirroring its margin.
 The size of the frame is another optimizable global parameter.
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{images/framing_artifacts.png}
-\caption{Artifacts occur w/o frames (left) that vanish with frames (right).} % TODO
-\label{framing_artifacts}
-\end{figure}
-
 \item For obtaining an initial reconstruction of the image, we compute a weighted sum over all the neighboring pixels that are known.

### 2011-06-24 project/Paper/paper.tex
Changed picturChanged picturee
Final paper (submitted) and removal of old files.

--- 2011-06-24 project/Paper/paper.tex
@@
 \begin{abstract}
-This paper presents a new approach to the problem of inpainting, that is recovering of lost image information.
-In comparison with the submissions of our peers, we have designed a very efficient algorithm that does not fall behind other solutions in terms of accuracy.
+This paper presents a new approach to the problem of inpainting, which is about recovering lost image information.
+In comparison with the submissions of our peers, we have designed an efficient algorithm that does not fall too much behind other solutions in terms of accuracy.
 By suppressing weak frequencies in the Fourier domain, we attempt to recover the intrinsic signal of the image.
@@
 \section{Introduction}
-Inpainting is about reconstructing the missing spots of an image by guessing the correct values.
-The quality of such a restoration method can be evaluated by masking an image and comparing the reconstruction with the original.\\
+Inpainting is about reconstructing the missing spots of an image by estimating the correct values.
+The quality of such a restoration method can be evaluated by masking an image and comparing the reconstruction with the original.
+
 Our approach is based on the assumption that (natural) images have some underlying characteristic that can be distinguished from noise.
-We initialize our reconstructed image with what we call Gaussian interpolation: Using MATLAB's filtering function, we compute a weighted sum over all the neighboring pixels that are known.
-This reconstruction introduces noise compared to the original image.
-The hope is that this noise is weak enough compared to the real signal, that we want to get rid of it.
-We do so by dividing the image into patches and analyzing these in the Fourier space, removing the weak frequencies that (hopefully) capture this noise and thereby recovering the intrinsic signal.\\
+The computed pixels can be viewed as a disturbed version of the genuine ones, hence the goal is to get rid of this noise.
+We do so by analyzing image patches in the Fourier space, removing the weak frequencies that (hopefully) capture the noise and thereby recovering the intrinsic signal.
+
 We repeat this procedure with the newly determined values until the error of a randomly chosen validation set converges or the maximum number of iterations is reached.
-This validation set consists of some of known pixels given to the algorithm, that are removed before starting the reconstruction.
-In order to determine the  threshold for suppressing weaker frequencies, this validation set (for which we know the correct values) can be used as an error measure.
+For a faster start, we initialize our reconstructed image with what we call Gaussian interpolation: Using MATLAB's filtering function, we compute a weighted sum over all the neighboring pixels that are known.
+In order to determine the best threshold for suppressing weaker frequencies, we randomly remove even more pixels from the image and use them as a validation set.
 
@@
 As outlined in the introduction, we combined several ideas for obtaining an inpainting algorithm that provides both good results and fast performance.
-The input of our algorithm is the image to be restored together with a mask that indicates its unknown spots.\\
+The input of our algorithm is the image to be restored together with a mask that indicates its unknown spots.
+
 Before entering the iterative part of our algorithm, the following three steps are performed (as seen in figure \ref{flowchart}):
@@
 The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section \ref{gradient_descent}).
-\item Our reduction in the Fourier space works best with frames, meaning that we use large overlapping patches for the fourier transform, but only use the inner part for the reconstruction, we frame the complete image by mirroring its margin.
-The size of the frame is another optimizable global parameter.
+\item Since our reduction in the Fourier space works best with framed patches, we frame the complete image by mirroring its margin. (The size of the frame is another optimizable global parameter.)
 \item For obtaining an initial reconstruction of the image, we compute a weighted sum over all the neighboring pixels that are known.
-By setting all unknown pixels in the image to zero and encoding known values with 1s in the mask, elementwise division of the filtered image by the similarly filtered mask results in a normalized sum weighted by the filtering kernel.\\
-This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.
+By setting all unknown pixels in the image to zero and encoding known values with 1s in the mask, elementwise division of the filtered image by the similarly filtered mask results in a normalized sum weighted by the filtering kernel.
+(This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.)
 \label{gaussian_interpolation}
@@
 \begin{enumerate}
-\item Using the above mentioned validation set, we evaluate several thresholds for every patch and keep for each the best one: After calculating a Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
+\item Using the above-mentioned validation set, we evaluate several thresholds for every patch and keep for each the best one: After calculating a fast Fourier transform for every patch, frequencies weaker then the threshold are suppressed by setting them to 0.
 Since profiling showed that the Fourier transform with its inverse is by far the most time-consuming part of our algorithm, we try to perform as few threshold evaluations as possible.
-For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and halve the step size (see figure \ref{threshold_optimization}).
+For this purpose, we iteratively evaluate two new thresholds around a center threshold, taking the best one as the new center and halving the step size (see figure \ref{threshold_optimization}).
 While this heuristic does not always reach the minimum, it works quite well for our error function, which is often roughly convex.
@@
 \centering
-\includegraphics[width=0.8\columnwidth]{../plots/search_strategy_6.pdf}
-\includegraphics[width=0.8\columnwidth]{../plots/search_strategy_2.pdf}
-\caption{The threshold is iteratively optimized.}
+\includegraphics[width=0.7\columnwidth]{../plots/search_strategy_6.pdf}
+\includegraphics[width=0.7\columnwidth]{../plots/search_strategy_2.pdf}
+\caption{Threshold are optimized iteratively with decreasing steps.}
 \label{threshold_optimization}
@@
 \hline
-Parameter & Description & Initial value & Optimized value\\
+Parameter & Description & Initial value & Optimized value \\
 \hline
-gauss\_size & Size of the Gaussian kernel in pixels & 10 & ? \\
-gauss\_sigma & Sigma of the Gaussian kernel & 1 & ? \\
-patch\_size & Patch size as exponent to the basis 2 & 4 & ? \\
-patch\_frame\_size & Frame size in pixels & 8 & ? \\
-td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds as negative exponent to basis 10& 3 & ? \\
-td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds as negative exponent to basis 10& 4 & ? \\
-td\_middle & Initial threshold & 10 & ? \\
-validation & Fraction of pixels in the validation set & 0.4 & ? \\
-max\_iterations & Maximum number of iterations & 5 & ? \\
-abortbelow\_change & Minimal relative change to perform another iteration & 0.1 & ? \\
+gauss\_size & Size of the Gaussian kernel in pixels & 10 & 19 \\
+gauss\_sigma & Sigma of the Gaussian kernel & 1 & 0.652 \\
+patch\_size & Patch size as exponent to the basis 2 & 4 & 9 \\
+patch\_frame\_size & Frame size in pixels & 8 & 0 \\
+td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds as negative exponent to the basis 10 & 3 & 1.940 \\
+td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds as negative exponent to the basis 10 & 4 & 3.523 \\
+td\_middle & Initial threshold & 10 & 16.156 \\
+validation & Fraction of pixels in the validation set & 0.4 & 0.015 \\
+max\_iterations & Maximum number of iterations & 5 & 7 \\
+abortbelow\_change & Minimal relative change to perform another iteration & 0.1 & 0.047 \\
 \hline
@@
 \end{center}
-\caption{Parameters with their initial and optimized values.}
+\caption{Parameters with their initial and optimized values (rounded according to their semantics).}
 \label{parameters}
@@
 \label{gradient_descent}
-We optimize all parameters at once using a gradient descent method.
+We optimize all parameters at once using a method called gradient descent.
 Since our cost function is based on measurements, we cannot derive it with respect to individual parameters and have to determine the gradient with finite differences.
-Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term of 0.8 that should help us to overcome local minima and improve convergence speed.
+Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term that should help us to overcome local minima.
 Figure \ref{gradient_descent_cost} shows the development of the cost for 100 optimization steps.
@@
 \centering
-\includegraphics[width=0.8\columnwidth]{../plots/cost_plot.pdf}
+\includegraphics[width=0.7\columnwidth]{../plots/cost_plot.pdf}
 \caption{Due to the momentum term, the cost function does not strictly decrease when optimizing.}
@@
 E+\exp(T/T_0 - 1) - 1 & \text{otherwise}
-\end{cases} $ & Continuous function, optimizes for error when the time is below a target time $T_0$, penalizes long times. & Seems to work quite well, used by us \\
+\end{cases} $ & Continuous function, optimizes for error when the time is below a target time $T_0$, penalizes long times & Seems to work quite well, used by us \\
 \hline
@@
 \item Some parameters have a clear speed vs. accuracy tradeoff, like the stopping conditions of the threshold optimization, which depend on our cost function and could not be foreseen.
-\item Some parameters assumed totally different values than anticipated: The patch size became much larger and the validation set decreased significantly, as can be seen in figure \ref{gradient_descent_parameters}.
+\item Some parameters assumed totally different values than anticipated: The patch size became much larger and the validation set decreased significantly (see figure \ref{gradient_descent_parameters}).
 \end{itemize}
@@
   \subfloat{\label{fig:tiger}\includegraphics[width=0.25\textwidth]{../plots/validation_plot.pdf}}
-  \caption{Two parameters that seem to converge quite well during the optimization.}
+  \caption{Two parameters that converged well during optimization.}
   \label{gradient_descent_parameters}
@@
 \item Gaussian interpolation as described in section \ref{gaussian_interpolation} and
-\item Matching pursuit~\cite{matchingpursuit93} with an overcomplete dictionary of discrete cosines.
+\item Matching pursuit~\cite{matchingpursuit93} with an over-complete dictionary of discrete cosines.
 \end{enumerate}
@@
 \centering
-\includegraphics[width=\columnwidth]{images/missingpixelsVsError.png}
+\includegraphics[width=\columnwidth]{images/missingpixelsVsError.pdf}
 \caption{Baseline algorithms for various fractions of missing pixels.}
@@
 Linear interpolation & 8.08\e{-4} & 7.10\e{-6} \\
-Gaussian interpolation & 5.74\e{-3} & 1.10\e{-5} \\
+Gaussian interpolation & 5.74\e{-4} & 1.10\e{-5} \\
 Matching pursuit & 1.72\e{-3} & 1.71\e{-5} \\
@@
 \centering
-\includegraphics[width=\columnwidth]{images/paper_reconstruction_bars32_mask8.png}
-\caption{Original image with mask (left, mask in blue) and its reconstruction (right). Enlarged for display in this paper.}
+\includegraphics[width=0.8\columnwidth]{images/paper_reconstruction_bars32_mask8.png}
+\caption{Original image with mask (left) and its reconstruction (right).}
 \label{edge_reconstruction}
@@
 \paragraph{Note} All plots in this paper were created with MATLAB and were produced using the script \emph{CreatePaperPlots.m} of our submission.
-However, due to the randomness of the algorithm and varying speed measurements for the cost function, results can be different for each run.
+However, due to the randomness of the algorithm and varying speed measurements for the cost function, the results are different for each run.
 
 \section{Discussion}
-Discuss the strengths and weaknesses of your approach, based on the results. Point out the implications of your novel idea on the application concerned. TODO
+The most remarkable result is that our fairly sophisticated algorithm (using the final parameters) achieves at most marginally better accuracy than the simple Gaussian interpolation.
+Since our iterative algorithm uses Gaussian interpolation for its initialization, this means that the frequency suppression in the Fourier domain has only little effect on the quality of the output.
+Interestingly enough, the optimization of our global parameters yielded a patch size equal to the image size for the applied cost function.
+This circumstance might be the reason why (in contrast to our prediction) no frame and only a small validation fraction is used.
+Other than one might be tempted to assume now, however, the thresholds are distinctly larger than zero so that indeed suppression takes place (which is visible in the good reconstruction of edges).
+Lastly, recovering the intrinsic signal of an image is of not much help, if the original image itself is very noisy, which is the case for the bicycles.
 
@@
 The remarkable speed is achieved by relying on MATLAB's native functions for the most computationally intensive tasks.
-In our opinion, we have devised many nice methods: Gaussian interpolation, threshold optimization with validation set and frequency suppression in the Fourier domain to name just the most important ones.
+In our opinion, we have devised many nice methods: Gaussian interpolation, threshold optimization with validation set and frequency suppression in the Fourier domain, to name just the most important ones.
 The key feature of our algorithm is certainly its ability to reconstruct the edges in an image.

### 2011-06-24 Project-kmeansVariant/inPainting.m

@@ -1,73 +1,0 @@
-function Iout = inPainting(I, mask)
-
-I = I .* mask;
-
-% Perform the actual inpainting of the image 
-
-% INPUT
-% I: (n x n) masked image
-% mask: (n x n) the mask hiding image information
-%
-% OUTPUT
-% I_rec = Reconstructed image 
-
-radius = 1;
-[height, width] = size(I);
-total = height * width;
-
-border_I = zeros(height+2*radius, width+2*radius);
-border_I((1+radius):(radius+height), (1+radius):(radius+width)) = I;
-border_out = border_I;
-
-border_mask = zeros(height+2*radius, width+2*radius);
-border_mask((1+radius):(radius+height), (1+radius):(radius+width)) = mask;
-
-tile_size = (radius*2+1)*(radius*2+1);
-
-X = zeros(tile_size, total*tile_size);
-maskX = zeros(tile_size, total*tile_size);
-
-counter = 1;
-for x = (1+radius):(radius+width)
-  for y = (1+radius):(radius+height)
-    small_counter = 1;
-
-    for xdiff = -radius:radius
-      for ydiff = -radius:radius
-        X(small_counter, counter) = border_I(y+ydiff, x+xdiff);
-        maskX(small_counter, counter) = border_mask(y+ydiff, x+xdiff);
-        small_counter = small_counter + 1;
-      end
-    end
-   
-    for p = 1:(tile_size-1)
-      X(:, counter+p) = circshift(X(:, counter), p);
-      maskX(:, counter+p) = circshift(maskX(:, counter), p);
-    end
-    
-    counter = counter + tile_size;
-  end
-end
-
-[z, U, score] = k_means(X, maskX, 50);
-
-counter = 1;
-for x = (1+radius):(radius+width)
-  for y = (1+radius):(radius+height)
-    %if (~border_mask(y, x))
-    %  average = sum(X(:, counter))/sum(maskX(:, counter));
-     % border_out(y,x) = average;
-    %end
-
-%    if (~border_mask(y, x))
-      my_clusters = U(:, z(counter:(counter+tile_size-1)));
-      my_cluster = sum(my_clusters,2) / tile_size;
-
-      border_out(y, x) = my_cluster(floor(tile_size/2+1));
-%    end
-
-    counter = counter + tile_size;
-  end
-end
-
-Iout = border_out((1+radius):(radius+height), (1+radius):(radius+width));

### 2011-06-24 Project-kmeansVariant/k_means.m

@@ -1,104 +1,0 @@
-function [z, U, score] = k_means(data, mask, nClusters, varargin)
-% K_MEANS   k-means clustering
-% 
-%   [z, U, score] = k_means(data, nClusters, varargin)
-%   Input:
-%       data        Data matrix, columns are observations
-%       nClusters   Number of clusters
-%
-%   Output:
-%       z           Cluster label vector
-%       U           Matrix, columns are estimated cluster means
-%       score       Negative log-likelihood of the estimated solution
-%                   given the data.
-%
-%   Additional parameters through varargin
-%       threshold   Stop if change of log-likelihood smaller than threshold,
-%                   the default is 1e-4.
-%       maxIter     Maximum number of iterations, the default is 100.
-%       means       Initialization of the cluster centroids.
-
-
-% initializations
-threshold = 1e-4;
-maxIter = 50;
-[nDims, nExamples] = size(data);
-score = realmax;
-change = threshold+1;
-iterations = 0;
-
-% initialization of the class means
-indices = randperm(nExamples);
-U = data(:,indices(1:nClusters));
-
-% parse varargin
-for k=1:2:length(varargin)
-    switch lower(varargin{k})
-    case 'threshold'
-        threshold = varargin{k+1};
-    case 'maxiter'
-        maxIter = varargin{k+1};
-    case 'means'
-        U = varargin{k+1};
-    otherwise
-        error(['Unknown parameter ''' varargin{k} '''.']);
-    end
-end
-
-% k-Means algorithm
-while (change>threshold && iterations < maxIter),
-    Z = zeros(nClusters, nExamples);
-    iterations = iterations+1;
-    disp(sprintf('Iterations = %d  Change = %0.5g.', iterations, change));
-
-    % Assignment step: estimate class indices
-    for(i=1:nExamples)
-      % Get data
-      dataline = data(:, i);
-      maskline = mask(:, i);
-
-      % Get indices that can be used
-      indices = find(maskline==1);
-
-      % Get shorter version, repeat
-      masked_dataline = dataline(indices);
-      
-      pointMat = repmat(masked_dataline,1,nClusters);
-
-      masked_U = U(indices, :);
-
-      % Calculate distance to each centroid
-      distance = (masked_U - pointMat).^2;
-
-      % Get smallest distance and assign to cluster
-      [tilde, ind] = min(sum(distance));
-      Z(ind,i) = 1;
-    end
-    
[…]
